---
title: "worksheetprep"
format: html
---

according to ethan the first part was uphill and then it was a little flat and then it was down and then swiggly and up and then down and then up and it went down and then down and out and then a little up and then down and then flat and then up and then flat and then steep uphill into sharp turn into finish.

```{r}
library(readr)
library(tidyverse)
library(modelr)
library(broom)
ski_df <- read_csv("~/Desktop/git/stat_289_score/awsmit22/skimodule/module/W10kTro_2257.csv")
ski_speeds <- ski_df |> mutate(Speed1 = 1.3/ Time1, Speed2 = 3/(Time2-Time1), Speed3 = 3.2/(Time3-Time2), Speed4 = 2.5/(Time4-Time3))
ski_speeds <- ski_speeds |> filter(!is.na(Rank4)) |> filter(!is.na(Speed1))
```

```{r}
write.csv(ski_speeds, "ski_speeds.csv", row.names = FALSE) 
```

```{r}
speeds_long <- ski_speeds |> pivot_longer(c("Speed1", "Speed2", "Speed3", "Speed4"), names_to = "speedinterval", values_to = "speed") |> pivot_longer(c("Rank1", "Rank2", "Rank3", "Rank4"), names_to = "rankinterval", values_to = "rank") |> relocate(speed, speedinterval, rankinterval, rank)
```

```{r}
ski_new <- ski_speeds |> pivot_longer(c("Speed1", "Speed2", "Speed3", "Speed4"), names_to = "speedinterval", values_to = "speed") |> relocate(speed, speedinterval)
```

```{r}
skimod_new<- aov(speed ~ Name + speedinterval, data = ski_new)
summary(skimod_new)
```

```{r}
tapply(ski_new$speed, ski_new$speedinterval, mean)
```

```{r}
ski_new2 <- na.omit(ski_new)
```

```{r}
tapply(ski_new2$speed, ski_new2$speedinterval, mean)
```

```{r}
ski_speeds_anova <- ski_new |> filter(!is.na(speed))
```

```{r}
tapply(ski_speeds_clean$speed, ski_speeds_clean$speedidinterval, mean)
```

```{r}
ski_df |> nrow()
```

```{r}
anovaski <- aov(speed ~ Name + speedinterval, data = ski_speeds_clean)
summary(anovaski)
```

```{r}
ski_speeds |> relocate(Speed1, Rank4)
```


# checking conditions for simple model

```{r}
mod_1 <- lm(Rank4~ Speed1, data = ski_speeds)
```


# linearity and variance
```{r}
#resid vs. fits plot
plot(mod_1$residuals~mod_1$fitted)
abline(0,0)

#scatter plot
plot(Rank4~Speed1, data = ski_speeds)
abline(mod_1)
```

The linearity and variance are both relatively good. I would say that the variance is stronger than the linearity.

```{r}
# Mod 1
mod_1 <- lm(Rank4 ~ Time1, data = ski_speeds)
summary(mod_1) # R^2: 77.7% 
plot(Rank4 ~ Time1, data = ski_speeds) # bad linearity
abline(h = 30)

# Mod 2 
mod_2 <- lm(Time4 ~ Time1, data = ski_speeds)
summary(mod_2) # R^2: 85.8%
plot(Time4 ~ Time1, data = ski_speeds) # better linearity
abline(h = 30.5)

# Mod 3
mod_3 <- lm(Rank4 ~ Rank1, data = ski_speeds)
summary(mod_3) # R^2: 83.7%
plot(Rank4 ~ Rank1, data = ski_speeds) # better linearity
abline(h = 30)

# Mod 4
mod_4 <- lm(Time4 ~ Rank1, data = ski_speeds)
summary(mod_4) # R^2: 81.9%
plot(Time4 ~ Rank1, data = ski_speeds) # bad linearity again
abline(h = 30.5)
```

***best mod is mod_2***

# possible follow-up questions

Interpret the slope coefficient (in context)  for the best of these models to predict Rank4.

```{r}
summary(mod_3)
# For every additional place in Rank1, Rank4 increases by 0.88 places. 
```

If a skier is ranked #14 after the first split, what rank would you predict for her finish?

```{r}
mod_3 <- lm(Rank4 ~ Rank1, data = ski_speeds)
summary(mod_3) 

Rank1 <- 14

Rank4 <- 2.95897 + 0.87807*Rank1

Rank4

# Predicted Rank4 is 15.25 which is about 15th place at the finish?
```

Interpret the slope coefficient (in context)  for the best of these models to predict Time4.

```{r}
summary(mod_2)

# For every additional minute Time1 increases, Time4 increases by 4.66 minutes. 
```

If a skier had a time of 5.00 minutes after the first split, what time would you predict for her finish?

```{r}
mod_2 <- lm(Time4 ~ Time1, data = ski_speeds)
summary(mod_2) 

Time1 <- 5

Time4 <- 6.7561 + 4.6618*Time1
Time4

# Predicted Time4 is 30.07 minutes. 
```



# residual normality
```{r}
#histogram of residuals
hist(mod_1$residuals)
# normal quantile plot
qqnorm(mod_1$residuals)
qqline(mod_1$residuals)
```
pretty good

# checking conditions for multi regression model

# linearity and variance
```{r}
#resid vs. fits plot
plot(modanova$residuals~modanova$fitted)
abline(0,0)
```

good


# residual normality
```{r}
#histogram of residuals
hist(modanova$residuals)
# normal quantile plot
qqnorm(modanova$residuals)
qqline(modanova$residuals)
```
not great

```{r}
cor(ski_speeds$Rank4, skispeeds$Speed1)
```


```{r}
lm(Rank4 ~ speed + Bib, data = ski_speeds_anova)
```

```{r}
write.csv(ski_speeds_anova, "ski_speeds_anova.csv", row.names = FALSE) 
```

