---
title: "pickleball_solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)

pickleBall <- read_csv("pickleBall.csv")
View(pickleBall)
```


Preform a Pearson's R correlation of `Overall` on `Overall Reliability`
```{r}
correlation <- pickleBall |>
  group_by(Sex) |>
  summarize(pearson_r = cor(Overall, `Overall Reliability`, method = "pearson"))
correlation
```

```{r}
cor_test_result <- cor.test(pickleBall$Overall, pickleBall$`Overall Reliability`, method = 'pearson')
cor_test_result

ggplot(data = pickleBall, aes(x = Overall, y = `Overall Reliability`, color = Sex)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  theme_minimal()
```

Model how Overall Reliability (reliable) is predicted by Overall (quality):
```{r}
model <- lm(`Overall Reliability` ~ Overall, data = pickleBall)
summary(model)
```

If you want to see if Sex moderates the relationship between Overall and Reliability:
```{r}
model2 <- lm(`Overall Reliability` ~ Overall + Sex, data = pickleBall)
summary(model2)
```

```{r}
model3 <- lm(`Overall Reliability` ~ Overall * Sex, data = pickleBall)
summary(model3)
```
This shows whether the relationship between quality and reliability is different for males vs females.

regression modeling quality versus reliability



Scatterplot of singles versus doubles rating
```{r}
ggplot(data = pickleBall, aes(x = Singles, y = Doubles, color = Sex)) +
  geom_point() +
  geom_smooth()
```

Scatterplot of doubles versus singles reliability
```{r}
ggplot(data = pickleBall, aes(x = `Singles Reliability`, y = `Doubles Reliability`, color = Sex)) +
  geom_point() +
  geom_smooth()
```

```{r}
ggplot(data = pickleBall, aes(x = Doubles, y = Mixed, color = Sex)) +
  geom_point() +
  geom_smooth()

model = lm(Mixed ~ Doubles, data = pickleBall)
plot(model)
```

```{r}
ggplot(data = pickleBall, aes(x = `Doubles Reliability`, y = `Mixed Reliability`, color = Sex)) +
  geom_point() +
  geom_smooth()
```

Histogram of Single rating with gender
```{r}
ggplot(data = pickleBall, aes(x = Singles, fill = Sex)) +
  geom_histogram()
```

Histogram of Single reliability with gender
```{r}
ggplot(data = pickleBall, aes(x = `Singles Reliability`, fill = Sex)) +
  geom_histogram()
```

Histogram of Double rating with gender
```{r}
ggplot(data = pickleBall, aes(x = Doubles, fill = Sex)) +
  geom_histogram()
```

Histogram of Double reliability with gender
```{r}
ggplot(data = pickleBall, aes(x = `Doubles Reliability`, fill = Sex)) +
  geom_histogram()
```


Boxplots:
```{r}
ggplot(data = pickleBall, aes(x = Singles, y = Doubles, fill = Sex)) +
  geom_boxplot()
```

```{r}
ggplot(data = pickleBall, aes(x = `Singles Reliability`, y = `Doubles Reliability`, fill = Sex)) +
  geom_boxplot()
```

```{r}
ggplot(data = pickleBall, aes(x = Singles, y = `Singles Reliability`, color = Sex)) +
  geom_point() +
  geom_smooth()

```


```{r}
pickle_long <- pivot_longer(pickleBall, cols = c(5, 7, 9, 11), names_to = "Category", values_to = "Rating")

ggplot(data = pickle_long, aes(x = Category, y = Rating, fill = Sex)) +
  geom_boxplot()
```

