---
title: "Scraping Ski to Sea Data"
format: html
editor: visual
---

# Scraping Ski to Sea Data 

Load Packages

```{r}
library(rvest)
library(dplyr)
library(lubridate)
library(tidyverse)
library(httr)
library(readr)
```

# Part 1) Creating Data Files for Each Leg (Sport)

Cross Country Ski

-   Note that the material in this chunk of code is repeated for each leg to gather data files

#### Question 1)

For the below code chunk, look through the results from [Ski to Sea Website](https://skitosea.pacificmultisports.com/events/){target="_blank"}. Find which leg number is associated with Cross Country Ski and which event number is associated with the first year (2009). Add in these numbers to the below code.

#### Question 2)

Load 2010 on the website and find the event and leg number for cross country skiing. Why does it make sense to use a for loop here?

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/" # go to URL and sift through results page

event <-  # notice how the event number changes for different years (2009 - 2019)
leg <-  # notice how the leg number changes for different sports 
isEvent <- TRUE # initializes the isEvent variable to track if the event exists

for (event in 101:111) { # starts a loop to iterate through events from 101 to 111
  url <- paste0(base_url, event, "/results/leg/", leg) # constructs the URL for the current event and leg
  response <- GET(url) # sends an HTTP GET request to the URL and stores the response
  if (response$status_code == 200) { # checks if the response status code is 200 (indicating a successful request)
    
    # scrape the data
    h <- read_html(url) # reads the HTML content of the URL
    tab <- h %>% html_nodes("table") # extracts all tables from the HTML content
    test <- tab %>% html_table() # converts the tables into data frames
    data <- test[[1]] # selects the first table and assigns it to the data variable
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/leg_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7 # increments the leg variable by 7 to move to the next leg for the next iteration.
    Sys.sleep(2) # pauses the script for 2 seconds to avoid overwhelming the server with requests
    
  } else {
    isEvent <- FALSE # if the event does not exist (response status code is not 200), sets isEvent to FALSE
  }
}
```

Downhill Skiing

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/"
event <- 101
leg <- 2
isEvent <- TRUE

for (event in 101:111) {
  url <- paste0(base_url, event, "/results/leg/", leg)
  response <- GET(url)
  if (response$status_code == 200) {
    
    # scrape the data
    h <- read_html(url)
    tab <- h %>% html_nodes("table")
    test <- tab %>% html_table()
    data <- test[[1]]
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/downhill_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7
    Sys.sleep(2)
    
  } else {
    isEvent <- FALSE
  }
}
```

#### Question 3)

Investigate the event and leg for the downhill ski leg in both 2009 and 2010, what is the main difference? Comment on any changes that occur to the event and leg when you shift to downhill skiing.


#### Question 4)

How do you anticipate this difference will impact the code for the rest of the legs? What is the maximum leg number for 2009-2019?

Running

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/"
event <- 101
leg <- 3
isEvent <- TRUE

for (event in 101:111) {
  url <- paste0(base_url, event, "/results/leg/", leg)
  response <- GET(url)
  if (response$status_code == 200) {
    
    # scrape the data
    h <- read_html(url)
    tab <- h %>% html_nodes("table")
    test <- tab %>% html_table()
    data <- test[[1]]
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/run_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7
    Sys.sleep(2)
    
  } else {
    isEvent <- FALSE
  }
}
```

Road Bike

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/"
event <- 101
leg <- 4
isEvent <- TRUE

for (event in 101:111) {
  url <- paste0(base_url, event, "/results/leg/", leg)
  response <- GET(url)
  if (response$status_code == 200) {
    
    # scrape the data
    h <- read_html(url)
    tab <- h %>% html_nodes("table")
    test <- tab %>% html_table()
    data <- test[[1]]
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/roadbike_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7
    Sys.sleep(2)
    
  } else {
    isEvent <- FALSE
  }
}
```

Canoe

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/"
event <- 101
leg <- 5
isEvent <- TRUE

for (event in 101:111) {
  url <- paste0(base_url, event, "/results/leg/", leg)
  response <- GET(url)
  if (response$status_code == 200) {
    
    # scrape the data
    h <- read_html(url)
    tab <- h %>% html_nodes("table")
    test <- tab %>% html_table()
    data <- test[[1]]
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/canoe_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7
    Sys.sleep(2)
    
  } else {
    isEvent <- FALSE
  }
}
```

Cross Country Bike

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/"
event <- 101
leg <- 6
isEvent <- TRUE

for (event in 101:111) {
  url <- paste0(base_url, event, "/results/leg/", leg)
  response <- GET(url)
  if (response$status_code == 200) {
    
    # scrape the data
    h <- read_html(url)
    tab <- h %>% html_nodes("table")
    test <- tab %>% html_table()
    data <- test[[1]]
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/xcbike_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7
    Sys.sleep(2)
    
  } else {
    isEvent <- FALSE
  }
}
```

Kayak

```{r}
base_url <- "https://skitosea.pacificmultisports.com/events/"
event <- 101
leg <- 7
isEvent <- TRUE

for (event in 101:111) {
  url <- paste0(base_url, event, "/results/leg/", leg)
  response <- GET(url)
  if (response$status_code == 200) {
    
    # scrape the data
    h <- read_html(url)
    tab <- h %>% html_nodes("table")
    test <- tab %>% html_table()
    data <- test[[1]]
    
    # save the data to a CSV file
    write.csv(data, file = paste0("sampeacock23/kayak_results/event_", event, "_leg_", leg, "_results.csv"))
    
    # append leg number
    leg <- leg + 7
    Sys.sleep(2)
    
  } else {
    isEvent <- FALSE
  }
}
```

# Part 2) Combining Data Files Into a Single Dataset for Each Leg

Combine Canoe Data

-   Note that the material in this chunk of code is repeated for each leg to combine files

#### Question 5)

Write a comment after each line of code to describe what that code is doing.

```{r}
file_names <- paste0("canoe_results/event_", 101:111, "_leg_", seq(5, 75, by = 7), "_results.csv") 

combined_canoe_data <- NULL =
nyears = length(file_names) 

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>% 
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>% 
  select(`Elapsed Time`, `Team Name`, `Year`) %>% 
  rename(canoe_time = `Elapsed Time`) 
  
  combined_canoe_data <- bind_rows(combined_canoe_data, data) 
}
```

Combine Downhill Data

```{r}
# List of file names
file_names <- paste0("downhill_results/event_", 101:111, "_leg_", seq(2, 72, by = 7), "_results.csv")

# Read and combine the files using a loop
combined_downhill_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(downhill_time = `Elapsed Time`)
  
  combined_downhill_data <- bind_rows(combined_downhill_data, data)
}
```

Combine Kayak Results

```{r}
# List of file names
file_names <- paste0("kayak_results/event_", 101:111, "_leg_", seq(7, 77, by = 7), "_results.csv")

# Read and combine the files using a loop
combined_kayak_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(kayak_time = `Elapsed Time`)
  
  combined_kayak_data <- bind_rows(combined_kayak_data, data)
}
```

Combine Roadbike Results

```{r}
# List of file names
file_names <- paste0("roadbike_results/event_", 101:111, "_leg_", seq(4, 74, by = 7), "_results.csv")

# Read and combine the files using a loop
combined_roadbike_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(roadbike_time = `Elapsed Time`)
  
  combined_roadbike_data <- bind_rows(combined_roadbike_data, data)
}
```

Combine Run Results

```{r}
# List of file names
file_names <- paste0("run_results/event_", 101:111, "_leg_", seq(3, 73, by = 7), "_results.csv")

# Read and combine the files using a loop
combined_run_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(run_time = `Elapsed Time`)
  
  combined_run_data <- bind_rows(combined_run_data, data)
}
```

Combine Cross Country Bike Results

```{r}
# List of file names
file_names <- paste0("xcbike_results/event_", 101:111, "_leg_", seq(6, 76, by = 7), "_results.csv")

# Read and combine the files using a loop
combined_xcbike_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(xcbike_time = `Elapsed Time`)
  
  combined_xcbike_data <- bind_rows(combined_xcbike_data, data)
}
```

Combine Cross Country Ski Results

```{r}
# List of file names
file_names <- paste0("xcski_results/event_", 101:111, "_leg_", seq(1, 71, by = 7), "_results.csv")

# Read and combine the files using a loop
combined_xcski_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(xcski_time = `Elapsed Time`)
  
  combined_xcski_data <- bind_rows(combined_xcski_data, data)
}
```

Combine Overall Results

```{r}
# List of file names
file_names <- paste0("overall_results/event_", 101:111,"_results.csv")

# Read and combine the files using a loop
combined_overall_data <- NULL
nyears = length(file_names)

for (i in 1:nyears) {
  data <- read_csv(file_names[i]) %>% 
  mutate(Year = i + 2008) %>%
  mutate(`Elapsed Time` = as.character(`Elapsed Time`)) %>%
  select(`Elapsed Time`, `Team Name`, `Year`) %>%
  rename(overall_time = `Elapsed Time`)
  
  combined_overall_data <- bind_rows(combined_overall_data, data)
}
```

# Part 3) Combine All Leg Datasets Into a Single Dataset

#### Question 6)
Join these data files

```{r}

```

#### Question 7)
Select relevant variables

```{r}

```

Mutate the elapsed time variables for each leg into minutes

```{r}

```

Save final data file to your project as a .csv

```{r}

```