---
title: "Welch's"
format: html
---


7) Split the dataset into two separate datasets. One for pre2018 and one for post2018 (including 2018)
```{r}
pre2018 <- menSkating |> filter(Year < 2018)

post2018 <- menSkating |> filter(Year >= 2018)
```

We will now preform a Welch's t-test which will compare the means of the two groups to determine if they are statistically significant.

8) 

a - Write down the Null Hypothesis and Alternative Hypothesis for this test

Null - There is no significant difference between the groups (mean difference is equal to 0)
Alternative - There is a significant difference between the groups (mean difference does not equal 0)

b - Preform a Welch's Two Sample t-test on the two groups. Interpret your results
```{r}
t.test(pre2018$TP, post2018$TP)
```

reject the null hypothesis

There is significant evidence (t = -8.4785, df = 104.53, p-value = 1.588e-13) to support the idea that there is a difference in the mean of Total Points between the years.

Notice how the Welch's t-test also preforms a confidence interval. If the confidence interval does not include 0 then we know that it supports the p-value.

c - Interpret the confidence interval

With 95% confidence, the true difference in mean total points between pre2018 and post2018 lies between −60.87 and −37.79. Since this interval does not include zero, it further supports the conclusion that there is a significant difference between the groups.


Residual and cook's d for outlier:
```{r}
ggplot(data = menSkating, aes(x = SP, y = FS, color = factor(Year), group = Year)) +
 geom_point() +
 geom_smooth(method = "lm", se = TRUE) +
 scale_color_brewer(palette = "Dark2") +
  labs(title = "Figure Skating Scatterplot",
       x = "Short Program",
       y = "Free Skate")

plot(model)
standardized_residuals <- rstandard(model)
standardized_residuals

cooks_d <- cooks.distance(model)
cooks_d
```

