---
title: "NonlinearLeastSquares_Solutions"
format: 
  html:
    embed-resources: true
format: html
---

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(broom)
library(readr)
powerlift <- read_csv(here::here("frhinehart/Powerlifting/module/powerlift_sample_data_updated.csv")) |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass) |>
  drop_na(Best3DeadliftKg)
```


1) Graph a scatter plot showing the relationship comparing Age and Best3DeadliftKg. Add a smoother. What do you notice about the curve?

```{r}
ggplot(data = powerlift, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(color = "blue", se = FALSE)
```

It is nonlinear.

2) Fit a nonlinear least squares model using the started gamma function provided below:
```{r}
# Defining a started gamma function
started_gamma <- function(x, a, b, x0) {
  x_start <- pmax(x - x0, 0)
  a * x_start * exp(-b * x_start)}
```

```{r}
nls_fit <- nls(Best3DeadliftKg ~ started_gamma(Age, a, b, x0),
 data = powerlift,
 start = list(a = 300, b = 0.05, x0 = 10))
nls_fit
```

a) Add predicted values to your dataset
```{r}
powerlift$pred <- predict(nls_fit)
```

b) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a prediction line. 
```{r}
ggplot(powerlift, aes(x = Age)) +
  geom_point(aes(y = Best3DeadliftKg), 
             alpha = 0.3, color = "gray40") +
  geom_smooth(aes(y = Best3DeadliftKg), color = "blue", fill = "blue") +
  geom_line(aes(y = pred), color = "red", linewidth = 1.2) +
 labs(title = "Fitted started Gamma Curve to Deadlift vs. Age",
   x = "Age",
   y = "Best 3 Deadlift (kg)") +
 theme_minimal()
```


3) Fit a nonlinear least squares model using the reparameterized double exponential with x_shift provided below
```{r}
double_exp_shift <- function(x, A, k1, k2, x_shift) {
  x_start <- pmax(x - x_shift, 0)
  A * exp(-k1 * x_start) * (1 - exp(-k2 * x_start))}
```

```{r}
nls_fit2 <- nls(Best3DeadliftKg ~ double_exp_shift(Age, A, k1, k2, x_shift),
  data = powerlift,
  start = list(A = 300, k1 = 0.02, k2 = 0.1, x_shift = 27))
nls_fit2
```

a) Add predicted values to your dataset
```{r}
powerlift$pred <- predict(nls_fit2)
```

b) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a line of best fit.
```{r}
ggplot(powerlift, aes(x = Age)) +
  geom_point(aes(y = Best3DeadliftKg), alpha = 0.3, color = "gray40") +
  geom_smooth(aes(y = Best3DeadliftKg), color = "blue", fill = "blue") +
  geom_line(aes(y = pred), color = "darkgreen", linewidth = 1.2) +
  geom_vline(xintercept = 8.5) +
  labs(title = "Double Exponential with Explicit shift Parameter",
    subtitle = "Estimated shift location is directly modeled",
    x = "Age",
    y = "Best 3 Deadlift (kg)") +
  theme_minimal()
```

c) Use a confidence interval for x-shift parameter and find the coefficients for A, k1, k2, and x_shift. Interpret your results, why can the lower bound not be calculated? Why does the value not it make sense?
```{r}
confint(nls_fit2, "x_shift")
coef(nls_fit2)
```

Based on the model fit with starting, we are 95% confident that the shift age of performance is less than or equal to 8.46 years. However, the lower bound of the confidence interval could not be estimated, suggesting the model may be unstable or the parameter is not well-identified.

*Delta method, boot strap



4) Compare both models you created. Which one is more useful in predicting dead lift using age? Report the residual sum-of-squares for both models

Gamma residual sum-of-squares: 9838215

Reparameterized residual sum-of-squares: 9448921

reparameterized double exponential with x_shift is more useful since it has a lower residual sum-of-squares

5) If you completed the other powerlift module, use the indicator variable code provided below to recreate a 5th order polynomial regression model and find the AIC
```{r}
powerlift_new$SexIND = ifelse(powerlift_new$Sex == "M", 1, 0)
``` 

```{r}
dl_finish <- lm(data = powerlift_new, Best3DeadliftKg ~ poly(Age, 5)*SexIND)
summary(dl_finish)

AIC(dl_finish)
```

a) Find the AIC for the model you answered in question 4
```{r}
AIC(nls_fit2)
```

b) Compare the polynomial model' AIC to your answer in question 4's AIC. Which one is better in predicting dead lift using age? Report each of the model's AIC. 

The polynomial model is better as it has a lower AIC

dl_finish AIC: 38328
nls_fit2 AIC: 40775.1

