---
title: "NonlinearLeastSqauresPowerliftingWorksheet"
format: 
  html:
    embed-resources: true
format: html
---

Powerlifting is a strength sport where participants compete to lift the heaviest weight. Competitors compete within weight classes, so their opponents are around the same weight as them. The sport consists of three different lifts: Squat, Bench, and Deadlift. Squat lifts require lifters to squat so that their hips are below their knees and lift the weight to a specific depth. Bench lifts are done while the lifter is lying on a bench. Bench lifters start with the bar weight being held in the air and must lower the bar down to their body. Once the bar touches their body, lifters wait for the command to lift the bar back into the air with their elbows fully extended. Deadlifting, like squats, start with the lifter standing up. The lifter must lift the bar off the ground and get into the position where their shoulders are back and knees are locked. Competitors are overseen by three judges who determine whether the lift follows the rules of the events. If a judge believes the lifter followed the rules, they will turn on a white light. In order for a lift to count, competitors must receive at least two white lights. 

The Powerlifting dataset features data from individual lifters during their most recent competitions. Variables include characteristics such as sex, age, birth year and age classes, and division. Other variables include 3 types of weightlifting: squats, benches, and deadlifts.The relevant variables for this specific module are Age and Best3DeadliftKg. 

For the sake of this module, the data set was tidied so that it only included the event "SBD" (Squat, Bench, Deadlift) and two types of equipment use: Raw, meaning no equipment except knee sleeves, and Single-ply, which is a single layer suit. The data was also tided to exclude any lifter that did not finish the competition or had missing information for key variables.

For this worksheet, we want to use Age as predictors for Best3DeadliftKg.

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(broom)
library(readr)
powerlift <- read_csv("frhinehart/Powerlifting/powerlift_sample_data_updated.csv") |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass) |>
  drop_na(Best3DeadliftKg)
```




1) Fit a nonlinear least squares model using the shifted gamma function provided below

```{r}
# Defining a shifted gamma function
shifted_gamma <- function(x, a, b, x0) {
  x_shift <- pmax(x - x0, 0)
  a * x_shift * exp(-b * x_shift)
}
```


a) Add predicted values to your dataset


b) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a line of best fit. 



2) Fit a nonlinear least squares model using the reparameterized double exponential with x_peak provided below
```{r}
double_exp_peak <- function(x, A, k1, k2, x_peak) {
  x_shift <- pmax(x - x_peak, 0)
  A * exp(-k1 * x_shift) * (1 - exp(-k2 * x_shift))
}
```


a) Add predicted values to your dataset


b) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a line of best fit.


c) Use a confidence interval for x-peak parameter and find the coefficients for A, k1, k2, and x_peak. Interpret your results, why can the lower bound not be calculated?




3) Compare both models you created. Which one is more useful in predicting dead lift using age? Report the residual sum-of-squares for both models



