---
title: "NonlinearLeastSquares_Solutions"
format: 
  html:
    embed-resources: true
format: html
---

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(broom)
library(readr)
powerlift <- read_csv("frhinehart/Powerlifting/powerlift_sample_data_updated.csv") |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass) |>
  drop_na(Best3DeadliftKg)
```

```{r}
# Defining a shifted gamma function
shifted_gamma <- function(x, a, b, x0) {
  x_shift <- pmax(x - x0, 0)
  a * x_shift * exp(-b * x_shift)
}
```

1) Fit a nonlinear least squares model using the shifted gamma function provided above

```{r}
nls_fit <- nls(
  Best3DeadliftKg ~ shifted_gamma(Age, a, b, x0),
 data = powerlift,
 start = list(a = 300, b = 0.05, x0 = 10)
)
nls_fit
```


a) Add predicted values to your dataset
```{r}
powerlift$pred <- predict(nls_fit)
```

b) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a line of best fit. 
```{r}
ggplot(powerlift, aes(x = Age)) +
  geom_point(aes(y = Best3DeadliftKg), 
             alpha = 0.3, color = "gray40") +
  geom_smooth(aes(y = Best3DeadliftKg), color = "blue", fill = "blue") +
  geom_line(aes(y = pred), color = "red", linewidth = 1.2) +
 labs(
   title = "Fitted Shifted Gamma Curve to Deadlift vs. Age",
   x = "Age",
   y = "Best 3 Deadlift (kg)") +
 theme_minimal()
```


2) Fit a nonlinear least squares model using the reparameterized double exponential with x_peak provided below
```{r}
double_exp_peak <- function(x, A, k1, k2, x_peak) {
  x_shift <- pmax(x - x_peak, 0)
  A * exp(-k1 * x_shift) * (1 - exp(-k2 * x_shift))
}
```

```{r}
nls_fit2 <- nls(
  Best3DeadliftKg ~ double_exp_peak(Age, A, k1, k2, x_peak),
  data = powerlift,
  start = list(A = 300, k1 = 0.02, k2 = 0.1, x_peak = 27)
)
nls_fit2
```

a) Add predicted values to your dataset
```{r}
powerlift$pred <- predict(nls_fit2)
```

b) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a line of best fit.
```{r}
ggplot(powerlift, aes(x = Age)) +
  geom_point(aes(y = Best3DeadliftKg), alpha = 0.3, color = "gray40") +
  geom_smooth(aes(y = Best3DeadliftKg), color = "blue", fill = "blue") +
  geom_line(aes(y = pred), color = "darkgreen", linewidth = 1.2) +
  labs(
    title = "Double Exponential with Explicit Peak Parameter",
    subtitle = "Estimated peak location is directly modeled",
    x = "Age",
    y = "Best 3 Deadlift (kg)"
  ) +
  theme_minimal()
```

c) Use a confidence interval for x-peak parameter and find the coeffients for A, k1, k2, and x_peak. Interpret your results, why can the lower bound not be calcualted?
```{r}
confint(nls_fit, "x_peak")
coef(nls_fit)
```

Based on the model fit with shifting, we are 95% confident that the peak age of performance is less than or equal to 8.46 years. However, the lower bound of the confidence interval could not be estimated, suggesting the model may be unstable or the parameter is not well-identified.

3) Compare both models you created. Which one is more useful in predicting dead lift using age? Report the residual sum-of-squares for both models

Gamma residual sum-of-squares: 9838215

Reparameterized residual sum-of-squares: 9448921

reparameterized double exponential with x_peak is more useful since it has a lower residual sum-of-squares

