---
title: "2025_Track_NLS_SOLUTIONS"
format: html
---

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)

track <- read.csv("WomensTrack2025.csv")
```

1. Make a scatterplot using distance (Event) to predict speed

```{r}
ggplot(track, aes(x = Event, y = Speed)) +
  geom_point()
```

The scatterplot above shows that at first, there is very quick decay in speed as distance increases, but then that decay slows and soon levels out. There are a few different ways we could account for this trend in a model. One would be to apply a transformation to the data and continue to use a linear model, another way (the one we will be exploring) is to switch to a non-linear model. 

There are many different kinds of non-linear models, so the first step in using one is to decide the most useful way to build it. In this case, there are 2 non-linear models that follow a similar trend to the data here. One is an exponential decay model and the other is a logarithmic decay mode. 

We will start by applying an exponential decay model to the data. An exponential decay model has the following form:

$$ y = ab^x$$
Where a is the initial value of y when x = 0 and b is the decay term (how fast the decay is).

Next we will apply a logarithmic decay model. Logarithmic decay models have the following form:

$$ y = a - b * \ln(x)$$
Where a is the value of y when x = 1-where the decay curve starts- and b is how fast the value decreases. As b is larger, the drop will be steeper at first. The ln(x) term allows the decay to slow over time.

--------
Report AIC, BIC, R-squared (for each model).

If you completed the transformation module with this data set, compare the model you created there with the models from this module.

Which model do you prefer? Why?