---
title: "outlierSoccer_solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
  docx:  default
  pdf:  default
---

Read in the Dataset:
```{r}
library(readr)
SoccerData <- read_csv("SoccerData.csv")
```

1) Create a scatterplot using games played (GP) to predict goals scored (G). Color by Sex and add a smoother. What do you notice from the plot? Are there any clear trends or outliers? 
```{r}
ggplot(data = SoccerData, aes(x = GP, y = G, color = Sex)) +
  geom_point() +
  geom_smooth()
```

Notice that the females tend to have more goals than the males and there is a distinct outlier that could be influential. 

2) Write out the population model.

$$G = \beta_0 + \beta_1GP $$

3) Fit a linear model using games played (GP) to predict goals (G). Write out the fitted model. What does the slope mean in the context of this model?

```{r}
model = lm(G ~ GP, data = SoccerData)
summary(model)
```

$$ \hat{G} = -0.14122 + 0.12946GP$$
The slope means that for every additional game played, a player is predicted to score approximately 0.129 more goals, on average, assuming all other factors remain constant.

4) Use the model summary plots to determine which observation has the highest residual and leverage. Who does this observation correspond to?
```{r}
plot(model)
```

Point 33 has the highest residual and leverage and it corresponding to Sophie Smith.

5) Based on the Residuals vs Leverage plot, is that point an influential outlier to the model? How do you know?

Yes because it is above the .5 cook's distance line meaning it has moderate influence on the regression model. 

6) Calculate the cook's distance of the influential point. 

```{r}
cooks_d <- cooks.distance(model)
cooks_d
```

7.565e-01

7) How might removing the point influence the model?

The model's fit might improve allowing for more precise estimates