---
title: "badminton_questions_prep"
format: html
---

# Libraries

```{r}
library(tidyverse)
library(readr)
```

# Data

```{r}
badminton_df <- read_csv("badminton.csv")

badminton_df <- badminton_df |> select(!...1)
```

# Question 1

Create a histogram of `Wins` and describe the distribution.

```{r}
hist(badminton_df$Wins)
```

*Right skewed, values clustered at the lower end of the x-axis.*

# Question 2

Obtain the summary statistics of `Wins` and fill them in below. 

```{r}
mean(badminton_df$Wins)
sd(badminton_df$Wins)
min(badminton_df$Wins)
max(badminton_df$Wins)
median(badminton_df$Wins)
```

*Mean: 10.76*
*Standard deviation: 15.84*
*Minimum: 0*
*Maximum: 107*
*Median: 4*

# Question 3

Calculate the range for Wins and for pts_for. 

```{r}
# Wins 
107-0 #107

# pts_for
max(badminton_df$pts_for)
min(badminton_df$pts_for)
6042-62 #5980
```

# Question 4

The ranges for Wins and pts_for are both quite large. What do you think is a possible reason for this flaw in the data?

*Some players have played for longer than others, for example, there could be a player who only played in the 2023 World Tour whereas there could also be a player who's played in all the World Tours since 2018.*

# Question 5

Create a side by side box plot of Wins for each Category. Describe the distribution.

```{r}
boxplot(Wins~Category, badminton_df)
```

*The range for Doubles appears bigger than that for Singles, both seem to have a lot of outliers.*

# Question 6

Perform a t.test to determine if there is a significant difference in the mean number of wins for doubles vs. singles. 

```{r}
t.test(Wins~Category, data = badminton_df)
```

*Null hypothesis: The mean wins for doubles = mean wins for singles*

*Alternative hypothesis: The mean wins for doubles != mean wins for singles*

*t-value: 0.17*

*p-value: 0.87*

*Conclusion: Do Not Reject Null Hypothesis*

*0.87 > 0.05*

*We do not have significant evidence that the mean wins for doubles is higher than the mean wins for Singles.* 

# Question 7

Fit a simple linear model of `pts_for` predicting `Wins`.

```{r}
lm(Wins~ pts_for, data = badminton_df)
```

*Wins = -0.6 + 0.017pts_for*

# Question 8

Interpret the slope of the model you just fit.

*For every additional point a player scores, their wins increase by 0.017.*

# Question 9

If a player has 216 points calculate their predicted wins. 

```{r}
-0.6 + (0.017*216)
```

*They would have about 3 wins.*

# Question 10

Assess the overall fit of the model. 

```{r}
mod <- lm(Wins~pts_for, data = badminton_df)
summary(mod)
```

*Null hypothesis: The slope = 0*

*Alternative hypothesis: The slope != 0*

*t-value: 124.73*

*p-value: <2e-16*

*Conclusion: Reject Null Hypothesis*

*<2e-16 < 0.05*

*We have significant evidence of a positive slope when using pts_for to predict Wins.* 

# Question 11

Find the 90% confidence interval of the slope of the model.

```{r}
mod <- lm(Wins~pts_for, data = badminton_df)
confint(mod, level=0.90)
```

*Lower: 0.01679255, Upper: 0.01724367*

*We are 90% confident that the slope of the model of pts_for predicting Wins is between 0.0168 and 0.0172*

# Question 12

Do you think that pts_for is really the most effective predictor of Wins? Why or why not? What could be a better predictor?

*pts_for is an obvious predictor of Wins as the more points a player has, the more wins they will have typically. However, this could make it ineffectvie as a predictor of Wins as they are so closely related to one another. A better predictor could be shot_pct or pts_agst.*




