---
title: "NBA Sampling Distribution"
author: "Cooper and Ivan"
format: 
  html:
    embed-resources: true
---

We're in need of some sampling distribution modules and I thought I would outline one that you could  make if you are interested.

Sport: NBA
Player: Shai Gilgeous-Alexander
Variables: Points (measured at the per game level)
Parameter of Interest: PPG (average) for post season run for different lengths 

Scenario: Use SGA's regular season points distribution as the population to "model" what if scenarios for his post-season scoring average. 


```{r}
# install.packages("hoopR") # run only once
```

```{r}
library(hoopR)
library(dplyr)
library(stringr)
library(lubridate)
```


```{r}
all_players <- load_nba_player_box(
  seasons = most_recent_nba_season()
  )

sga <- all_players |> filter(str_detect(athlete_display_name, "Shai Gilgeous-Alexander")) |>
  filter(team_name == "Thunder") |>
  filter(did_not_play == FALSE)

# This gives both post season (season_type == 3)
# and regular season (season_type == 2)
# Note that the NBA Cup Championship game is
# included in game-by-game data here. This
# means that SGA has 77 regular season games 
# to use instead of the 76 listed in his
# Stats page (https://www.espn.com/nba/player/stats/_/id/4278073/shai-gilgeous-alexander)


```


Here is the process.

1. Extract just regular season games. Use these as the "population".
2. Export the Points column only to a csv file for StatKey.

StatKey version 
Sample "n" games from reg season points. Calculate mean points and build the sampling distribution.

Pre-simulation:
- Look at a dotplot of his Points and describe is (shaoe, center, spread, etc)


Simulation 1:
- Figure out how many games SGA played in the first round and use that as his "n"
- Can compare his actual PPG for the first round to the dotplot that is created for this simulation

Simulation 2:
- Append the second round to this (i.e., n is now the total number of games played in the first two rounds)

Simulation 3:
- Add the conference finals to the sample size

Simulation 4: 
- Use all 23 playoff games
- Can again compare his playoff PPG to the resulting distribution

For each Simulation round, have students discuss shape, center, spread for that round. At the end students should come up with some general "rules" or "trends" that they see.


This then leads into the Central Limit Theory (Normal model) style.