---
title: "lax faceoffs review module"
format: html
editor: visual
---

```{r}
library(readr)
lax <- read_csv("aj_modules/college_lacrosse/single_prop_module/data/lax_2022_2023.csv")
```
 
1. Suppose Duke is playing against Maryland. If each team has an equal chance of winning any faceoff (i.e., the winner of the faceoff is completely random), what is the chance that Duke would win a single faceoff?

p=0.5

2. Duke won 330 out of 548 total faceoffs in 2022-2023. State the null and alternative hypotheses to determine if Duke’s faceoff win percentage is better than what we’d expect with random chance only?

Ho: p = 0.5
Ha: p > 0.5

3. Determine the sample size and sample proportion for Duke's faceoff performance.
n= 548
phat = 330/548 
```{r}
duke_phat <- 330/548
```

4. Under the assumption that the null hypothesis is true, check if the sample satisfies the success-failure condition.

5.  Calculate the test statistic for Duke's faceoff performance.

zstat = 4.776

6. Based on the significance level ⍺ = .05, determine if this sample proportion provides evidence that Duke has a higher faceoff win percentage than expected under random chance in the 2022-2023 season. Provide all necessary details and a conclusion in context.

Normal distribution pvalue = 0
We reject Ho, so we have significant evidence that Duke's faceoff percentage is better than what we expect with random chance (0.5)

7. Considering that Duke was the second-best team in the country last season, reflect on the hypothesis test results and its implications on faceoffs in Division I Lacrosse. How does this statement align with the findings from the hypothesis test? Consider the broader implications of faceoffs in Division I Lacrosse and how this aspect of the game may impact team rankings and outcomes.

Having a high faceoff percentage means that you will have a successful season. 
