---
title: "Test data"
format: html
editor: visual
---

```{r}
library(readr)
library(dplyr)
D3_2022_Postseason <- read_csv("hdono20/D3_2022_Postseason.csv")
```

```{r}
# clean up dataset
d3_wsoc2022 <- D3_2022_Postseason %>%
  rename("away_team" = Away) %>%
  rename("home_team" = Home) %>%
  rename("away_score" = 'Away Score') %>%
  rename("home_score" = 'Home Score') %>%
  rename("date" = Date)

write_csv(d3_wsoc2022, "d3_wsoc2022.csv")
```





```{r}
Standings_D3 <- read_csv("hdono20/Standings_D3.csv")
```


Test for correlation
```{r}
cor(d3_wsoc2022$away_score, d3_wsoc2022$home_score)
```

Negative correlation between away score and home score:
-0.3108331. We expect a negative correlation between away score and home score because if a team scores a lot of goals, then the other team most likely did not score that many. 
```{r}
#Test for correlation
cor.test(d3_wsoc2022$away_score, d3_wsoc2022$home_score)
```
r*sqrt(n-2)/(sqrt(1-r)) correlation value

not normally distributed based on scatterplot 


2-way table for distribution of scores: contigency table
```{r}

library(dplyr)
test <- D3_2022_Postseason %>%
  count(`Away Score`)

  
test1 <- D3_2022_Postseason %>%
   count(`Home Score`)
```

```{r}
test_split <- test[5:15,] %>%
  mutate(count_sum = sum(n)) %>%
  mutate(Away_Score = '4 or more') %>%
  select(Away_Score, count_sum) %>%
  slice(1) %>%
  rename("n" = "count_sum") 
```

```{r}
test_new <- test[1:4,] %>%
  rename("Away_Score" = 'Away Score') %>%
  mutate(Away_Score = as.character(Away_Score))
```

```{r}
Away_Score_join <- bind_rows(test_new, test_split)
```

```{r}
test_split_Home <- test1[5:19,] %>%
  mutate(count_sum = sum(n)) %>%
  mutate(Home_Score = '4 or more') %>%
  select(Home_Score, count_sum) %>%
  slice(1) %>%
  rename("n" = "count_sum") 
```


```{r}
test_new_Home <- test1[1:4,] %>%
rename("Home_Score" = 'Home Score') %>%
    mutate(Home_Score = as.character(Home_Score))

```

```{r}
Home_Score_join <- bind_rows(test_new_Home, test_split_Home)
```


```{r}
truncated_dataset1 <- truncated_dataset %>%
  filter(Home_Score_truncated == "3+" & Away_Score_truncated == "2")
 
```





```{r}
#Make values that have a score greater than 4, just 4 for both Away Score and Home Score
truncated_dataset <- D3_2022_Postseason %>%
  mutate(Away_Score_truncated = if_else(`Away Score` > 2, "3+", as.character(`Away Score`))) %>%
  mutate(Home_Score_truncated = if_else(`Home Score` > 2, "3+", as.character(`Home Score`)))
```


```{r}
two_way_table <- table(truncated_dataset$Home_Score_truncated, truncated_dataset$Away_Score_truncated)
two_way_table
```

Write a csv file for this
```{r}
write.csv(two_way_table, "two_way_table.csv", row.names = FALSE)
```

```{r}
library(readr)
two_way_table <- read_csv("hdono20/two_way_table.csv")
```



Chi-square test: tests for association between two categorical variables

```{r}
chi_sq_test <- chisq.test(truncated_dataset$Home_Score_truncated, truncated_dataset$Away_Score_truncated)
chi_sq_test
```

Get expected counts
```{r}
chi_sq_test$expected
```


```{r}
assocplot(t(two_way_table))
```


```{r}
chi_sq_test$residuals
```
```{r}
plot(d3_wsoc2022$away_score, d3_wsoc2022$home_score)
```

