---
title: "SportMovies"
format: html
editor: visual
---

-   Webscrapping and Joining data sets would be for STAT234
-   Analysis of the final joined product will be for STAT113

Websites to show: IMDb: https://developer.imdb.com/non-commercial-datasets/ Wikipedia Sports: https://en.wikipedia.org/wiki/List_of_sports_films

Load packages

```{r}
library(tidyverse)
library(rvest)

```

Load in IMDb movies data

```{r}
title_basics <- readr::read_tsv("https://datasets.imdbws.com/title.basics.tsv.gz")
ratings <- readr::read_tsv("https://datasets.imdbws.com/title.ratings.tsv.gz") 
```

Join the IMDb movie data sets together

```{r}
IMDb <- full_join(title_basics, ratings, by = "tconst") # do we want this? theres a lot of NAs
#write.csv(IMDb,'agbign20/data/IMDb.csv')
```

Webscrape Sport Films data from Wikipedia for every sport we want to use

```{r}
url <- "https://en.wikipedia.org/wiki/List_of_sports_films"
## convert the html code into something R can read
h <- read_html(url)
## grabs the tables
tab <- h %>% html_nodes("table")
objs <- tab %>% html_table()
lapply(objs,dim)
```

Create individual data frames from the wikipedia

```{r}
cricket <- objs[[5]] %>% mutate(sport = "cricket")
rowing <- objs[[19]] %>% mutate(sport = "rowing")
figureSkating <- objs[[26]] %>% mutate(sport = "figure skating")
golf <- objs[[27]] %>% mutate(sport = "golf")
martialArts <- objs[[30]] %>% mutate(sport = "martial arts")
rodeo <- objs[[50]] %>% mutate(sport = "rodeo")
```

Create data frames for the sports on Wikipedia with their own separate links

```{r}
#football
football_url <- "https://en.wikipedia.org/wiki/List_of_American_football_films"
## convert the html code into something R can read
football_h <- read_html(football_url)
## grabs the tables
football_tab <- football_h %>% html_nodes("table")
football_objs <- football_tab %>% html_table()
lapply(football_objs,dim)
football <- football_objs[[1]] %>% mutate(sport = "football")


#basketball
basketball_url <- "https://en.wikipedia.org/wiki/List_of_basketball_films"
basketball_h <- read_html(basketball_url)
basketball_tab <- basketball_h %>% html_nodes("table")
basketball_objs <- basketball_tab %>% html_table()
lapply(basketball_objs,dim)
basketball <- basketball_objs[[1]] %>% mutate(sport = "basketball") %>% mutate(Year = as.integer(Year))


#baseball
baseball_url <- "https://en.wikipedia.org/wiki/List_of_baseball_films"
baseball_h <- read_html(baseball_url)
baseball_tab <- baseball_h %>% html_nodes("table")
baseball_objs <- baseball_tab %>% html_table()
lapply(baseball_objs,dim)
baseball <- baseball_objs[[1]] %>% mutate(sport = "baseball") %>% rename("Genre"="Type")


#iceHockey
icehockey_url <- "https://en.wikipedia.org/wiki/List_of_films_about_ice_hockey"
icehockey_h <- read_html(icehockey_url)
icehockey_tab <- icehockey_h %>% html_nodes("table")
icehockey_objs <- icehockey_tab %>% html_table()
lapply(icehockey_objs,dim)
icehockey <- icehockey_objs[[1]] %>% mutate(sport = "ice hockey")


#autoRacing
autoracing_url <- "https://en.wikipedia.org/wiki/List_of_auto_racing_films"
autoracing_h <- read_html(autoracing_url)
autoracing_tab <- autoracing_h %>% html_nodes("table")
autoracing_objs <- autoracing_tab %>% html_table()
lapply(autoracing_objs,dim)
autoracing <- autoracing_objs[[2]] %>% mutate(sport = "autoracing")

#cycling 
cycling_url <- "https://en.wikipedia.org/wiki/List_of_films_about_bicycles_and_cycling"
cycling_h <- read_html(cycling_url)
cycling_tab <- cycling_h %>% html_nodes("table")
cycling_objs <- cycling_tab %>% html_table()
lapply(cycling_objs,dim)
cycling <- cycling_objs[[1]] %>% mutate(sport = "cycling")

# sport of athletics
athletics_url <- "https://en.wikipedia.org/wiki/List_of_films_about_the_sport_of_athletics"
athletics_h <- read_html(athletics_url)
athletics_tab <- athletics_h %>% html_nodes("table")
athletics_objs <- athletics_tab %>% html_table()
lapply(athletics_objs,dim)
athletics <- athletics_objs[[2]] %>% mutate(sport = "athletics")

#boxing
boxing_url <-"https://en.wikipedia.org/wiki/List_of_boxing_films"
boxing_h <- read_html(boxing_url)
boxing_tab <- boxing_h %>% html_nodes("table")
boxing_objs <- boxing_tab %>% html_table()
lapply(boxing_objs,dim)
boxing <- boxing_objs[[2]] %>% mutate(sport = "boxing")

#horseRacing
horse_url <- "https://en.wikipedia.org/wiki/List_of_films_about_horse_racing"
horse_h <- read_html(horse_url)
horse_tab <- horse_h %>% html_nodes("table")
horse_objs <- horse_tab %>% html_table()
lapply(horse_objs,dim)
horse <- horse_objs[[1]] %>% mutate(sport = "horse racing")
```

Join all the sport films

```{r}
sportFilms <- bind_rows(athletics, autoracing, baseball, basketball, boxing, cricket, cycling, figureSkating, football, golf, horse, icehockey, martialArts, rodeo) %>% unite(col = "Types", c("Event(s)", "Type", "Style", "Ref"), na.rm = TRUE) %>% unite(col = "Summary", c("Summary", "Notes"), na.rm=TRUE)
write.csv(sportFilms, 'agbign20/data/sportfilms.csv')
```
