---
title: "Clinton Canoe Regatta: Looking at Normality in Distributions"
format: html
---

```{r}
library(tidyverse)
library(readr)
library(ggplot2)
library(forcats)

clinton <- read.csv("phoebechar/canoeing/clinton/data_repo_submission/full_clinton.csv")

clinton <- clinton |> group_by(year, bib) |> slice(1)

clinton <- clinton |> filter(time != 0)

ggplot(clinton, aes(x = time)) +
  geom_histogram(binwidth = 600)

clinton_pro <- clinton |> mutate(pro = if_else(grepl("pro", classID), 1, 0)) |>
  filter(pro == 1)

clinton_canoe <- clinton |> mutate(pro = if_else(grepl("pro", classID), 1, 0)) |>
  filter(pro == 0) |> mutate(canoe = if_else(grepl("oc", classID), 1, 0)) |> filter(canoe == 1)
```

This histogram shows the distribution of times on the Clinton from 2014-2025 for only the pro boats. 
```{r}
ggplot(clinton_pro, aes(x = time)) +
  geom_histogram(binwidth = 600, colour = "royalblue4", fill = "royalblue3") +
  xlim(25000, 55000) +
  ylim(0, 70) +
  labs(x = "Time to Complete the Race (in sec)",
       y = "Count (number of boats)",
       title = "Time distribution for pro canoes at the Clinton") +
  theme_minimal()
```

Describe the shape, spread, and skew of the data.
Estimate the mode.


This histogram shows the distribution of times on the Clinton from 2014-2025 for all canoes that are not pro boats
```{r}
ggplot(clinton_canoe, aes(x = time)) +
  geom_histogram(binwidth = 600, colour = "royalblue4", fill = "royalblue3")+
  xlim(25000, 55000) +
  ylim(0, 70) +
  labs(x = "Time to Complete the Race (in sec)",
       y = "Count (number of boats)",
       title = "Time distribution for non-pro canoes at the Clinton") +
  theme_minimal()
```
Describe the shape, spread, and skew of the data.
Estimate the mode.

How are the distributions different? Provide a possible explanation for why this could be.

In 2025, Adam Blanchard and Rebecca Sutton placed first in the C2 Stock (not pro) Mixed class with a time of 8 hours, 32 minutes, and 21 seconds (30741 seconds). In the same year, Dean Brown & Chris Prater placed 40th in the pro class with a time of 8 hours, 3 minutes and 11 seconds (28991 seconds). Which of these times is better relatively?

Observe skew and compare. Comment on mode/mean. Compare times relatively. z-scores. 

```{r}
quantiles <- c(0, 0.25, 0.5, 0.75, 1)

canoe_sum <- tibble(value = c(mean(clinton_canoe$time), sd(clinton_canoe$time),
                              quantile(clinton_canoe$time, quantiles)
                              ), 
                    stat = c("Mean", "SD", "Min", "Q1", "Median", "Q3", "Max"))

canoe_sum <- canoe_sum |> tidyr::pivot_wider(names_from = stat, values_from = value) |>
  knitr::kable()


pro_sum <- tibble(value = c(mean(clinton_pro$time), sd(clinton_pro$time),
                              quantile(clinton_pro$time, quantiles)
                              ), 
                    stat = c("Mean", "SD", "Min", "Q1", "Median", "Q3", "Max"))

pro_sum <- pro_sum |> tidyr::pivot_wider(names_from = stat, values_from = value) |>
  knitr::kable()
```

