---
title: "Clinton Canoe Regatta: Looking at Normality in Distributions"
format: html
---

```{r}
library(tidyverse)
library(readr)
library(ggplot2)

clinton <- read.csv("phoebechar/canoeing/clinton/data_repo_submission/full_clinton.csv")

clinton <- clinton |> group_by(year, bib) |> slice(1)

clinton <- clinton |> filter(time != 0)

ggplot(clinton, aes(x = time)) +
  geom_histogram(binwidth = 600)

clinton_pro <- clinton |> mutate(pro = if_else(grepl("pro", classID), 1, 0)) |>
  filter(pro == 1)

clinton_canoe <- clinton |> mutate(pro = if_else(grepl("pro", classID), 1, 0)) |>
  filter(pro == 0) |> mutate(canoe = if_else(grepl("oc", classID), 1, 0)) |> filter(canoe == 1)


```

This histogram shows the distribution of times on the Clinton from 2014-2025 for only the pro boats. 
```{r}
ggplot(clinton_pro, aes(x = time)) +
  geom_histogram(binwidth = 600) +
  xlim(25000, 55000) +
  ylim(0, 70)
```
Describe the shape, spread, and skew of the data.
Estimate the mode.


This histogram shows the distribution of times on the Clinton from 2014-2025 for all canoes that are not pro boats
```{r}
ggplot(clinton_canoe, aes(x = time)) +
  geom_histogram(binwidth = 600)+
  xlim(25000, 55000) +
  ylim(0, 70)
```
Describe the shape, spread, and skew of the data.
Estimate the mode.

How are the distributions different? Provide a possible explanation for why this could be.

In 2025, Adam Blanchard and Rebecca Sutton placed first in the C2 Stock (not pro) Mixed class with a time of 8 hours, 32 minutes, and 21 seconds (30741 seconds). In the same year, Dean Brown & Chris Prater placed 40th in the pro class with a time of 8 hours, 3 minutes and 11 seconds (28991 seconds). Which of these times is better relatively?

Observe skew and compare. Comment on mode/mean. Compare times relatively. z-scores. 



