---
title: "Chi-Square_Module"
format: docx
editor: visual
---

```{r}
library(tidyverse)
cdSportMoviesCount <- read.csv("agbign20/data/cdSportMoviesCount.csv")
cdSportMoviesCount <- cdSportMoviesCount[,-1]
```

## Chi-Square STAT-113 Module

1. In your own words, what is a Chi-Square Test of Independence? What questions is the test trying to answer?

Chi-squre test of independence is a hypothesis test used to see if two categorical variables are likely to be associated or not.
The questions: Are the two variables independent? Are variable 1 and variable 2 related?


2.  Is there a connection between certain types of sports and whether a sport film is more likely to be a comedy or drama movie? Use the data located in cdSportsMoviesCount.csv to address this research question.

```{=html}
<!-- -->
```

(a) State the appropriate hypotheses and check the conditions, report the test statistic, degrees of freedom and p-value, reach a decision and give a conclusion within the context of the problem.

```{r}
chisq.test(cdSportMoviesCount$Comedy, cdSportMoviesCount$Drama, correct = FALSE)
```

H_o: Sport is not associated with Genre (Comedy, Drama) H_a: Sport is associated with Genre (Comedy, Drama) Checking the conditions: Reasonable sample size in each category


(b) Report the test statistic and degrees of freedom

 Test Statistic: X-squared = 30
 Degrees of freedom = 25
 

(c) What is the p-value? And what evidence does it show?

p-value = 0.2243 
Do not reject null hypothesis. 
There is no significant evidence that types of sports is associated with whether a sport film genre is a comedy or drama.


(d) Provide a brief summary of the analysis using the top contributors to the chi-square statistic to provide justification.

There is no difference in the type of sport and the genre of the sport films (comedy or drama in this case). The test statistic is not large meaning that the null hypothesis will not be rejected. The two categorical variables are independent


(e) Based on this example, what can you say about the relationship between the test statistic and the p-value?

When there is a large test statistic the null hypothesis will be rejected (depending on the df value), meaning the p-value is very low and significant. To determine how large the test statistic needs to be, the degrees of freedom is used.
