---
title: "sportfilms"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

```{r}
allSportFilms <- read.csv("agbign20/data/allSportsFilms.csv")
```

Manipulate how we want it to be

```{r}
allSportFilms <- allSportFilms[, -c(1,2,5,7,8)]
allSportFilms <- allSportFilms %>%
  mutate(runtimeMinutes = as.numeric(runtimeMinutes)) %>%
  filter(isAdult == "0")
allSportFilms <- allSportFilms[,-3]
```

```{r}
ggplot(allSportFilms, aes(x=averageRating)) + geom_bar()
ggplot(allSportFilms, aes(x=titleType)) + geom_bar() # mostly tv episodes
ggplot(allSportFilms, aes(x=runtimeMinutes)) + geom_bar() # lets look at this one closer
ggplot(allSportFilms, aes(x=sport)) + geom_bar() # mostly boxing movies
```

```{r}
allSportFilms %>% filter(runtimeMinutes <= 250) %>% ggplot(aes(x=runtimeMinutes)) + geom_bar()
# looks like about 40 and 60 minutes have the highest count, makes sense since tvepisode is the highest count of title type
```


Create a boxplot with sport on x axis (pick common sports), comedy and drama as different boxplot colors, and average rating on y axis

NOTE: Might need to change the genre variable to get different results

```{r}
#sports <- c("boxing", "football", "autoracing", "martialArts", "golf", "ice hockey")

comedyDramaSports <- allSportFilms %>% mutate(ComedyIND = ifelse(str_detect(genres,"Comedy"),"Comedy",NA),
         DramaIND = ifelse(str_detect(genres,"Drama"),"Drama",NA)) %>%
  pivot_longer(cols=ComedyIND:DramaIND,names_to=NULL,values_to="Type") %>%
  drop_na("Type") %>% select(primaryTitle, Year, Type, averageRating, sport) %>%
  filter(sport %in% sports)

ggplot(comedyDramaSports, aes(x = sport, y = averageRating, fill = Type)) + geom_boxplot()

```
explain not much is going on
two way anova


```{r}
newAllSportFilms <- allSportFilms %>% 
  filter(titleType == "movie") %>% 
  filter(Year >= 1980)

newAllSportFilms$Genre <- paste(newAllSportFilms$genre, newAllSportFilms$Genres, sep = ",")
newAllSportFilms <- newAllSportFilms[,-c(1,4,8)]

cdsports <- newAllSportFilms %>% mutate(ComedyIND = ifelse(str_detect(Genre, "Comedy"), "Comedy", NA),
                                     DramaIND = ifelse(str_detect(Genre, "Drama"), "Drama", NA)) %>%
  pivot_longer(cols=ComedyIND:DramaIND,names_to = NULL, values_to = "Type") %>%
  drop_na("Type") %>% select(primaryTitle,Year, Type, averageRating, sport)
```

1970s instead of 1980s would be 1700 obervations vs 1500 observations

Comedy vs Drama Films

Side by side boxplot of comedy and drama
```{r}
ggplot(cdsports, aes(x = Type, y = averageRating, fill = Type)) + geom_boxplot()
```

Two sample t test
```{r}
comedy <- cdsports %>% filter(Type == "Comedy")
drama <- cdsports %>% filter(Type == "Drama")
t.test(comedy$averageRating, drama$averageRating, var.equal = TRUE) # welchs t test (doesnt assume variance is equal)
```
t test statistic = -3.5366
p-value = 0.0004946
The 95% CI for the true difference in means between the two groups turns out to be [-0.6402101, -0.1820104]
The sample mean of group 1 was 5.88189 and the sample mean of group 2 was 6.293
Ho: u1 = u2 (mu) --> the two population means are equal
Ha: u1 does not equal u2 (mu) --> the two population means are not equal

Because the pvalue of our test is about 0, we reject the null hypothesis.
There is sufficient evidence that there is a difference in the mean average rating between comedy and drama films.

Break down by sport
```{r}
cdsports %>% count(sport, Type) %>% view() # look at the number of comedy and drama movies for each sport
cdsports <- cdsports %>% 
  mutate(sport = fct_other(cdsports$sport, keep = c("autoracing", "baseball", "basketball", "football", "martialArts"))) %>%
  mutate(Type = as.factor(Type))
ggplot(cdsports, aes(x = sport, y = averageRating, fill = Type)) + geom_boxplot()
```



Two Way Anova Analysis 
```{r}

cdsports %>% group_by(Type, sport) %>%
  summarize(mean = round(mean(averageRating, na.rm = TRUE)),
            sd = round(sd(averageRating, na.rm = TRUE))) %>% view()

mod <- aov(averageRating ~ Type * sport, data = cdsports)
summary(mod)
```
With sport held constant, average rating is significantly different between the two genres for type.
With type held constant, average rating is not significantly different between the sports.
The interaction between type and sport is not quite significant.


tukey HSD : type and sports
```{r}
library(multcomp)

mod1 = glm(averageRating ~ Type*sport, data=cdsports)


summary(glht(mod1, mcp(sport ="Tukey"))) # avg rating is not significantly different between sports
summary(glht(mod1, mcp(Type ="Tukey"))) # avg rating is not significantly different between the two genres
```



















Question:
- to get rid double counting movies if they are both comedy and drama, see if their indicator variables are both 1, make an indicator variable like bothIND and then make the other two indictors both 0 again


