
```{r}
library(tidyverse)
```

```{r}
data <- read_csv("timeout2.csv")
```
214/2718? 

```{r}
# the total number of sets = 3293
max(data$overall_set_number)
```

```{r}
# losing team won the set after the time out was called 
sum(data$losing_team_win)
```

in total: out of 3293 sets, in 1011 of them, the losing team at the time out won the set
1011/3293 = 0.307 = 30.7 % 

Construct a 95% confidence interval for the proportion of losing teams winning a set after the final timeout.

p^ = 0.307
Standard Error = sqrt((0.307*(1-0.307))/3293) = 0.008037851
z = 1.96
Margin of Error = 1.96*0.008037851 = 0.01575419

0.307 +- 0.01575419

95% CI = (0.291,0.323)

first time out or laast time out in the set 
after the first time out who wins the set 
after the last tme out who wins the set 


proportion of teams that win the set after losing before the last time out 

```{r}
data_last_timeout <- data %>% group_by(overall_set_number) %>%
  slice_max(home_after_timeout)

sum(data_last_timeout$losing_team_win)
```
168/2900 = 0.058 = 5.8% of teams that were losing win the game after the last time out


proportion of teams that win the set after losing before the first time out 
```{r}
data_first_timeout <- data %>% group_by(overall_set_number) %>%
  slice_min(home_before_timeout)

sum(data_first_timeout$losing_team_win)
```
513/2875 = 0.178 = 17.8% of teams that were losing before the first timeout win the set 
