
```{r}
library(tidyverse)
```

```{r}
data <- read_csv("timeout2.csv")
```
214/2718

first time out or last time out in the set 
after the first time out who wins the set 
after the last tme out who wins the set 

```{r}
# distinct number of overall sets there are
n_distinct(data$overall_set_number)
```


proportion of teams that win the set after being behind before the last time out 

```{r}
# getting the last time out of a set - putting into data set 
data_last_timeout <- data %>% group_by(overall_set_number) %>%
  slice_tail(n=1)

# getting how many losing teams win the set out of the last time outs 
sum(data_last_timeout$losing_team_win)

# the number of sets 
n_distinct(data_last_timeout$overall_set_number)
```

139/2781 = 0.04998 = 4.998% of teams that were losing win the game after the last time out

proportion of teams that win the set after losing before the first time out 
```{r}
# getting the first time out of a set - putting into data set 
data_first_timeout <- data %>% group_by(overall_set_number) %>%
  slice_head(n=1)

# getting how many losing teams win the set out of the first 
sum(data_first_timeout$losing_team_win)

# the number of sets 
n_distinct(data_first_timeout$overall_set_number)
```

497/2781 = 0.1787 = 17.87% of teams that were losing before the first timeout win the set 
