---
title: "ufc"
author: "Brendan Karadenes"
date: "2024-02-01"
output: html_document
---

Fighter Data


source: https://www.kaggle.com/datasets/rajeevw/ufcdata

Motivation:
 - Data contains information about UFC fighters from 1993-2021
 - Includes height, weight, DOB, Reach, Stance and Takedown and striking info.
 - Use data for intro stat classes
 - Possible correlations: Stance vs. Decision, Striking Accuracy vs. Reach (within weightclass groups)
 
 Data Definitions
 
 - fighter_name: First and Last name of fighters
 - Height: Height (in inches) of fighters
 - Weight: Weight in pounds of the fighters
 - Reach: Wingspan of fighters
 - Stance: Fighter postion
 - DOB: Birthdate of fighter
 - SLpM: Strikes landed per minute
 - Str_Acc: Striking Accuracy percentage
 - SApm: Strikes Attempted per minute
 - Str_Def: Striking defense percentage
 - TD_Avg: Takedown average per fight
 - TD_Acc: Takedown accuracy in percentage
 - TD_Def: Takedown defense in percentage
 - Sub_Avg: Submission average
 




Quick Sample question ideas/cleaning data

```{r}
library(tidyverse)
library(readr)
fight_data <- read.csv("bkaradenes/raw_fighter_details.csv")
```

Example Stat 113 Questions.

1.

```{r}
fight_data2 <- fight_data %>%
  filter(Reach != "")%>%
  mutate(Reach = parse_number(Reach))
```


```{r}
library(tidyverse)
ggplot(data = fight_data2) + geom_boxplot(mapping = aes(x = Reach)) + theme_minimal()
fight_data2 %>%
  summarise(sd(Str_Acc))


```

1. Based on the boxplot on fighter reaches, give an approximate 5 number summary.

2. Are there any outliers?

3. Is fighter reach approximately normal? Why or why not.

Example 2.

```{r}
fight_data2 <- fight_data %>%
  mutate(Str_Acc = parse_number(Str_Acc)) %>%
  filter(Str_Acc != 0)
```


```{r}
ggplot(data = fight_data2) + geom_histogram(mapping = aes(x = Str_Acc)) + labs(x = "Striking Accuracy Percentage", title = "Distribution of Striking Accuracy % for UFC fighters") + theme_minimal() + theme(plot.title = element_text(hjust = .4))
```

1. Describe the distribution for fighter striking accuracies.

2. Where is the approximate median?


```{r}
fight_data2<-fight_data2%>%
  mutate(Weight = parse_number(Weight)) %>%
  mutate(Reach = parse_number(Reach)) %>%
  filter(Reach != "")
  
```

```{r}
fight_data2 <- fight_data2 %>%
  mutate(Reach = parse_number(Reach)) %>%
  filter(Reach != "")
```

```{r}
fight_data2 <- fight_data2 %>%
  mutate(Str_Def = parse_number(Str_Def)) %>%
  mutate(TD_Acc = parse_number(TD_Acc)) %>%
  mutate(TD_Def = parse_number(TD_Def))
```


ScatterPlot

```{r}
fight_data2 %>%
  ggplot(mapping = aes(x = Str_Acc, y = Reach)) + geom_smooth(method = "lm") + geom_point()
```

Stat113 Striking Accuracy % histogram Potential Questions

- Z-scores, probabilities by the tails, since normal distributions, p-values

```{r}
fight_data2 %>%
  summary(mean(Str_Acc))
```

```{r}
orthodox_fight <- fight_data2 %>%
  filter(Stance == "Switch") # try different weight classes
```

```{r}
ggplot(data = orthodox_fight, mapping = aes(x = Str_Acc)) + geom_histogram()
```



# Questions
Use worksheet

The average UFC striking percentage for fighters from 1993-2021 sat at about 44.3%. Assume the distribution is approximnently normal.

1. Calculate the proportion 
 








