---
editor: 
  markdown: 
    wrap: 72
---

exploration

```{r}
library(tidyverse)
library(forcats)
```

Tidying the data to get it how i want

```{r}
asia_cup <- read_csv("~/Desktop/asiacup.csv", col_names = TRUE)
View(asia_cup)
```

Coin toss - 1 is won 0 is lost coin toss result- 1 is won 0 is lost

data cleaning

```{r}
asia_cup <- select(asia_cup, `Team`,
         `Opponent`,
         `Ground`,
         `Year`,
         `Toss`,
         `Selection`,
         `Run Scored`,
         `Fours`,
         `Sixes`,
         `Extras`,
         `Highest Score`,
         `Result`,
         `Given Extras`) %>%
  rename(Host = Ground,
         `Extras Scored` = Extras) %>%
  mutate(Result = ifelse(Result == "Win D/L", "Win", Result),
         Result = ifelse(Result == "Lose D/L", "Lose", Result),
         Result = ifelse(Result == "Win", 1, 0),
         Toss = ifelse(Toss == "Win", 1, 0),
         Selection = ifelse(Selection == "Batting", 0, 1)) %>%
  filter(Result != "No Result") %>%
  na.omit()
```

```{r}
write_csv(asia_cup,
          file = "~/GitHub/stat_289_score/vgjohn/cricket_module/Data_Repo_Submission/asia_cup.csv")
```

stat 213/234 level, r handout predicting win/loss based on fours sixes
and extra runs scored proportions difference in means for countries?
making a new variable ? diff in proportions

```{r}
n_distinct(asia_cup$Team)
```

There can be no result in a cricket match. it is the case when the match
cannot be finished in time, usually due to weather or lighting. It has
the same significance as a draw in most cases for the sake of this they
have been removed

Make a table that contains 5 new variables: `Total Games`, `Wins`,
`Losses`, and `Win Percentage` that calculates the corresponding win
percentage for each distinct country. Which country has the highest win
percentage?

```{r}
win_percentage <- asia_cup %>%
  group_by(Team) %>%
  summarize(
    `Total Games` = n(), 
    `Wins` = sum(Result == 1),
    `Losses` = sum(Result == 0),
    `Win Percentage` = Wins/`Total Games`
  )

win_percentage
```

```{r}
asia_cup %>% filter(Team == "Afghanistan")
```

looking at runs scored for the teams with the highest win percentage

what year did India score the most runs

```{r}
runs_year <- asia_cup %>% filter(Team == "India") %>%
  group_by(Year) %>%
  summarize(Runs = sum(`Run Scored`))

max_runs <- slice_max(runs_year, Runs)

runs_year
max_runs
```

most extras scored

```{r}
asia_cup$Team
group_by(asia_cup, Team) %>% slice_max(`Extras Scored`) %>%
  arrange(desc(`Extras Scored`))
```

density plot of the wins over time for each team

```{r}

```

difference in means

Conduct the appropriate test if want to figure out if there is a
significant difference in the number of runs scored between a team that
lost the coin toss and a team that won the coin toss. what are the null
and alternative hypothesis in context.

null - the difference in average runs scored between a team that won the
coin toss and a team that lost the coin toss is zero alternative - the
difference in average runs scored between a team that won the coin toss
and that lost the coin toss is not equal to zero.

At the 0.05 level, is there a significant difference in number of runs
scored for teams that won the toss and teams that lost the toss? why?
interpret it

```{r}
t.test(`Run Scored` ~ Toss, data = asia_cup)
```

We cannot conclude that there is a significant difference in runs scored
in a cricket match between a team that won the coin toss and a team that
lost the coin toss. We know this because the p-value of the two sample t
test is 0.0582, which is above 0.05 which we would need to reject the
null hypothesis that the difference in means between the two groups is
zero.

diff means for 4s between won and lost match

```{r}
t.test(`Fours` ~ Result, data = asia_cup)
```

logistic regression

fit a logistic regression for result using run scored, extras scored,
given extras, and highest score as predictors

```{r}
logit <- glm(Result ~ `Run Scored` + `Sixes` + `Selection` + `Given Extras` + Selection:`Run Scored`,
    family = binomial,
    data = asia_cup)

summary(logit)
```

```{r}
logit.result2 = glm(Result ~  `Fours` + `Highest Score` + `Given Extras`,
                   family = binomial,
                   data = asia_cup)

summary(logit.result2)
```

What would the estimated model say about the chance that a team with
\_\_\_\_\_ extra runs \_\_\_\_\_ given extras and \_\_\_\_\_ runs scored
wins the game

interpret the coefficients

```{r}
logit3 <- glm(Result ~ `Highest Score` + Selection + Selection:`Highest Score`+ `Toss`, 
family = binomial,
data = asia_cup)

summary(logit3)
```


```{r}
-4.814+0.018(265)+0.147(8)+4.440(0)+0.007(5)-0.018(0)
```
-4.814+0.018(265)+0.147(8)+4.440(0)+0.007(5)-0.018(0)

e^(-4.814+0.018(265)+0.147(8)+4.440(0)+0.007(5)-0.018(0))/(1+ e^(-4.814+0.018(265)+0.147(8)+4.440(0)+0.007(5)-0.018(0)) )