---
title: "horse_exploration"
author: "eric seltzer"
format: html
---

```{r}
library(tidyverse)
```

Variables and Descriptions

| Variable  | Description                                              |
|-----------|----------------------------------------------------------|
| rid       | Race ID                                                  |
| horseName | Name of horse                                            |
| age       | Age of horse                                             |
| saddle    | Saddle # where horse starts                              |
| isFav     | Was horse a favorite before race? (Can be more than one) |
| position  | Finishing position, 40 if horse didn't finish            |
| RPR       | RP rating                                                |
| TR        | Topspeed                                                 |
| OR        | Official rating                                          |
| Weight    | Horse weight (kg)                                        |
| res_win   | Did the horse win or not                                 |
| res_place | Did the horse place or not                               |

```{r}
horse <-
  read_csv(here::here("esselt21/module_exploration/data/horses.csv"))
```

```{r}
horseClean <-
  horse |>
  select(c(rid, horseName, age, saddle, isFav, position, RPR, TR, OR, weight, res_win, res_place))
```

```{r}

```

