---
title: "nba_wingspan_exploration"
format: html
editor: visual
---


# From crafted NBA, wingspans dataset
```{r}
library(tidyverse)
library(rvest)
library(janitor)
```

```{r}
url <- "https://craftednba.com/player-traits/length"
objs <-
  url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

wingspans <- objs[[1]]
```


## clean up data
```{r}
tidy_wingspans <- 
  wingspans %>%
  select(-1) %>%
  # tidy height
  mutate(
    height = Height
  ) %>%
  separate_wider_delim(
      cols = Height, 
      names = c("feet", "inches"), 
      delim = "'") %>%
  mutate(
    feet = parse_number(feet),
    inches = parse_number(inches),
    height_inches = feet * 12 + inches
  ) %>%
  select(-feet, -inches) %>%
  # tidy wingspan
  mutate(wingspan = Wingspan) %>%
  separate_wider_delim(
    cols = Wingspan,
    names = c("feetW", "inchesW"),
    delim = "'") %>%
  mutate(
    feetW = parse_number(feetW),
    inchesW = parse_number(inchesW),
    wingspan_inches = feetW * 12 + inchesW
  ) %>%
  select(-feetW, -inchesW) %>%
  rename(wingspan_advantage = Length,
         name = Name) %>%
  relocate(wingspan_advantage, .after = last_col()) %>%
  # tidy name
  separate_wider_delim(
    cols = name,
    names = c("name", "position"),
    delim = " | ") %>%
  arrange(name) %>%
  # remove free agents
  filter(str_detect(name, pattern = "[A-Z]{3}$")) %>%
  mutate(
    team = str_extract(name, pattern = "[A-Z]{3}$"),
    name = str_remove(name, pattern = "[A-Z]{3}$")
  ) %>%
  relocate(team, .after = name)

```


```{r}
# fixing duplicate players
tidy_wingspans %>%
  group_by(team, position, height, wingspan) %>%
  filter(n() > 1) %>%
  summarise(name)

# only dupes are Alex Sarr and Bub Carrington, so remove them
# fix name errors
nba_wingspans_2025 <- tidy_wingspans %>%
  filter(name != "Alexandre Sarr" & name != "Carlton Carrington") #%>%
  # mutate(name = if_else(name == "AJ Johnson", "A.J Johnson", name))

# remove name that shouldn't be there
nba_wingspans_2025 <- nba_wingspans_2025 %>%
  slice(-149)

```

## write csv
```{r}
#write_csv()
```



# From basketball reference, shooting and adjusted shooting
```{r}
# shooting table
url <- "https://www.basketball-reference.com/leagues/NBA_2025_shooting.html"
objs2 <-
  url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

shooting <- objs2[[1]]
```

```{r}
# adjusted shooting table
url <- "https://www.basketball-reference.com/leagues/NBA_2025_adj_shooting.html"
objs3 <-
  url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

adjusted_shooting <- objs3[[1]]
```

## general tidying and selecting rows
```{r}
# clean up shooting table
tidy_shooting <- shooting %>% 
  row_to_names(row_number = 1) %>%
  clean_names() %>%
  select(2, 6, 8, 9, 10, 11, 16, 17, 22) %>%
  rename(games = g,
         minutes = mp,
         avg_dist_of_fg = dist,
         `2pt_rate` = x2p,
         `3pt_rate` = x3p,
         `2pt_percent` = x2p_2,
         `3pt_percent` = x3p_2) %>%
  mutate(games = as.numeric(games))
```

```{r}
# clean up adjusted shooting
tidy_adjusted_shooting <- adjusted_shooting %>%
  row_to_names(row_number = 1) %>%
  clean_names() %>%
  select(2, 6, 12, 13, 14) %>%
  mutate(g = as.numeric(g))
```


## remove certain rows
When a player has played on two teams they will show up three times, the stats for each team and the total stats thorughout the season.
```{r}
tidy_shooting <- tidy_shooting %>%
  group_by(player) %>%
  slice_max(order_by = games, n = 1, with_ties = FALSE) %>%
  ungroup()
```

```{r}
#check code above
tidy_shooting %>%
  count(player) %>%
  filter(n > 1)
```


```{r}
tidy_adjusted_shooting <- tidy_adjusted_shooting %>%
  group_by(player) %>%
  slice_max(order_by = g, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-g)
```


## combine shooting and adjusted shooting
```{r}
nba_shooting_2025 <-
  tidy_shooting %>%
  left_join(tidy_adjusted_shooting, by = "player") %>%
  slice(-361) %>%
  mutate(
    across(3:12, as.numeric)
  )
```


# Check: combine wingspan data with shooting data
```{r}
nba_wingspans_2025 %>%
  left_join(nba_shooting_2025, by = c("name" = "player"))
```

```{r}
# Check with fuzzyjoin
library(fuzzyjoin)

stringdist_join(
nba_wingspans_2025, nba_shooting_2025,
by = c("name" = "player"),
method = "osa",
max_dist = 2,
mode = "left"
) %>%
  select(-player) %>%
  # show players that it doubles up
  count(name) %>%
  filter(n > 1)

# don't think fuzzy join is possible for this dataset.
```



# Combine wingspan and shooting data
About 15 names don't match between datasets, so change them in the wingspan dataset. Also a couple of players didn't play at all during the 2024-2025 season so remove them from the wingspan dataset.
```{r}
nba_wingspans_shooting_2025 <- nba_wingspans_2025 %>%
  # names that dont match between datasets
  mutate(
    name = if_else(name == "Alperen Sengun", "Alperen Şengün", name),
    name = if_else(name == "Danté Exum", "Dante Exum", name),
    name = if_else(name == "GG Jackson", "GG Jackson II", name),
    name = if_else(name == "Jimmy Butler III", "Jimmy Butler", name),
    name = if_else(name == "Monté Morris", "Monte Morris", name),
    name = if_else(name == "Robert Williams III", "Robert Williams", name),
    name = if_else(name == "Ronald Holland II", "Ron Holland", name),
    name = if_else(name == "Tristan da Silva", "Tristan Da Silva", name),
    name = if_else(name == "Xavier Tillman", "Xavier Tillman Sr.", name)
  ) %>%
  # players that didn't play in 2024-2025 season
  # 85, 320, 360
  slice(-85,-320,-360) %>%
  # check combining
  left_join(nba_shooting_2025, by = c("name" = "player"))
  
```

```{r}
write_csv(nba_wingspans_shooting_2025, "partial_wingspan_shooting.csv")
```




