---
title: "SoccerModuleSolutions"
format: html
---

```{r Read data, echo= FALSE, output = FALSE}
## Read in Data
library(tidyverse)
library(ggiraphExtra)

players_df <- read.csv("https://query.data.world/s/kgjdpxxovikkuxbduexafg6fbrbjav?dws=00000",
                       header=TRUE, stringsAsFactors=FALSE)


```




```{r Height Distribution, echo=FALSE, output=FALSE}

## Exploratory Data Analysis


## Height Distribution
library(ggplot2)

players_df_height <- 
  players_df |> 
  filter(height_in_cm > 0)


ggplot(data = players_df_height, 
       aes(x= height_in_cm)) +
  geom_bar()

```

```{r Boxplot of Height across Diff pos, echo=FALSE, output=FALSE, results='asis'}

## Distribution of Height across different positions

ggplot(data = players_df_height, 
       aes(y= height_in_cm,
           color = position)) +
  geom_boxplot() + 
  theme(legend.position = 'none') +
  labs(y = "Height (cm)")

```

A researcher wants to investigate whether there are differences among the mean height for different positions (Goalkeeper, Defender, Midfield, or Attack) of soccer players. Goalkeepers are typically thought to be taller than other positions to make up in wingspan to cover the goal better. The ‘players_df’ dataset contains information on each player including their name, current club, position, and height. For this analysis, only the position and height of the player will be used. A sample of 20 players was taken from each position and will be used for this analysis.



```{r echo=FALSE, output = TRUE, results='asis'}

set.seed(10294119)

height_sample_df <- players_df_height |> group_by(position) |>
  slice_sample(n = 20)

```


```{r echo=FALSE, output = TRUE, results='asis'}
height_sample_df$position <- factor(height_sample_df$position)

ggplot(data = height_sample_df, 
       aes(y= height_in_cm,
           x = position)) +
  geom_boxplot(aes(color = position)) +
  theme(legend.position = 'none') +
  labs(y = "Height (cm)")
```


#### Question 1.

a. Calculate the mean height for each position.

```{r}
height_sample_df |> group_by(position) |>
  summarise(Mean_height = mean(height_in_cm))
```


b. Are there any issues with the data that could affect the results?

The data seems to be relatively symmetric with no visible outliers from the boxplots. However, there seems to be an issue with variability. Spread in boxplot differed greatly.


c. Write of the null and alternative hypotheses that would be consistent with the research question.

$H_o : \mu_1 = \mu_2 = \mu_3 = \mu_4 = 0  $

$H_a : at \: least\: one \: \mu_i \neq 0 $


d. Complete the ANOVA F-test.

```{r echo= FALSE, output = TRUE}
height_sample_df$position <- recode(height_sample_df$position, 
                                    Goalkeeper = 'GK', 
                                       Defender = 'Def',
                                      Midfield = 'Mid',
                                       Attack = 'Att')

a1<- aov(data=height_sample_df, height_in_cm ~ position)


summary(a1)

```

e. Appropriately interpret results of analysis of variance F-test. 

$H_o : \mu_1 = \mu_2 = \mu_3 = \mu_4 = 0  $

$H_a : at \: least\: one \: \mu_i \neq 0 $


F-value = 11.18 with df = 3 and 76

p-value $\approx 0.000$ 

General Conclusion:

Reject $H_o$

Conclusion in context:

We have found strong enough evidence to support that there is a a clear relationship between the position and height of a soccer player.


#### Question 2. 

We will now perform a follow-up analysis to determine which pairs of positions have significantly different height means.

```{r echo= FALSE}

TukeyHSD(a1, conf.level=.95)

```


#### Tukey Pair-wise comparison Plot


```{r echo=FALSE}

ggHSD(TukeyHSD(a1, conf.level=.95), no = 1)

```

a.	Write a few sentences summarizing the interesting findings between the groups

Based on the Tukey HSD at the 5% level goalkeeper and attack positions, goalkeeper and defender positions ,and goalkeepers and midfielder positions are significantly different from each other. On average, goalkeepers seem to be the tallest, followed by defenders and then attackers and midfielders. From this, we can conclude that the natural assumption that goalkeepers are taller to account for their wingspan to better cover the goal is correct.

