---
title: "nba_exploration"
format: html
editor: visual
---

# 1. From crafted NBA, wingspans dataset

```{r}
library(tidyverse)
library(rvest)
library(janitor)
```

```{r}
url <- "https://craftednba.com/player-traits/length"
objs <-
  url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

wingspans <- objs[[1]]
```

## clean up data

```{r}
tidy_wingspans <- 
  wingspans %>%
  select(-1) %>% # in preclean
  
  # tidy height
  mutate(
    height = Height # worksheet, copies Height var, to use later
  ) %>%
  separate_wider_delim( # worksheet
      cols = Height, 
      names = c("feet", "inches"), # removes Height and splits into ft and in
      delim = "'") %>%
  mutate( # worksheet
    feet = parse_number(feet), 
    inches = parse_number(inches),
    height_inches = feet * 12 + inches # new var, in numerical(inches) format
  ) %>%
  select(-feet, -inches) %>% # removing helping vars
  
  # tidy wingspan
  mutate(wingspan = Wingspan) %>% # worksheet, copies Wingspan var, to use later
  separate_wider_delim( # worksheet
    cols = Wingspan,
    names = c("feetW", "inchesW"), # removes wingspan, splits into ftW and inW
    delim = "'") %>%
  mutate( # worksheet
    feetW = parse_number(feetW),
    inchesW = parse_number(inchesW),
    wingspan_inches = feetW * 12 + inchesW # new var, in numerical(inches) format
  ) %>%
  select(-feetW, -inchesW) %>% # remove helping var
  
  rename(wingspan_advantage = Length, # in preclean
         name = Name) %>%
  relocate(wingspan_advantage, .after = last_col()) %>%
  
  # above will be part a
  # and below will be part b
  
  # tidy name
  separate_wider_delim( # worksheet
    cols = name,
    names = c("name", "position"),
    delim = " | ") %>%
  arrange(name) %>%
  
  
  # remove free agents
  filter(str_detect(name, pattern = "[A-Z]{3}$")) %>%
  mutate(
    team = str_extract(name, pattern = "[A-Z]{3}$"),
    name = str_remove(name, pattern = "[A-Z]{3}$")
  ) %>%
  relocate(team, .after = name)

```

```{r}
# fixing duplicate players
tidy_wingspans %>%
  group_by(team, position, height, wingspan) %>%
  filter(n() > 1) %>%
  summarise(name)

# only dupes are Alex Sarr and Bub Carrington, so remove them
# fix name errors
nba_wingspans_2025 <- tidy_wingspans %>%
  filter(name != "Alexandre Sarr" & name != "Carlton Carrington") #%>%
  # mutate(name = if_else(name == "AJ Johnson", "A.J Johnson", name))

# remove name that shouldn't be there
nba_wingspans_2025 <- nba_wingspans_2025 %>%
  slice(-149)

```

## write csv

```{r}
#write_csv()
```

# 2. From basketball reference, shooting and adjusted shooting

```{r}
# shooting table
url <- "https://www.basketball-reference.com/leagues/NBA_2025_shooting.html"
objs2 <-
  url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

shooting <- objs2[[1]]
```

```{r}
# adjusted shooting table
url <- "https://www.basketball-reference.com/leagues/NBA_2025_adj_shooting.html"
objs3 <-
  url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

adjusted_shooting <- objs3[[1]]
```

## general tidying and selecting rows

```{r}
# clean up shooting table
tidy_shooting <- shooting %>% 
  row_to_names(row_number = 1) %>%
  clean_names() %>%
  select(2, 6, 8, 9, 10, 11, 16, 17, 22) %>%
  rename(games = g,
         minutes = mp,
         avg_dist_of_fg = dist,
         `2pt_rate` = x2p,
         `3pt_rate` = x3p,
         `2pt_percent` = x2p_2,
         `3pt_percent` = x3p_2) %>%
  mutate(games = as.numeric(games))
```

```{r}
# clean up adjusted shooting
tidy_adjusted_shooting <- adjusted_shooting %>%
  row_to_names(row_number = 1) %>%
  clean_names() %>%
  select(2, 6, 12, 13, 14) %>%
  mutate(g = as.numeric(g))
```

## remove certain rows

When a player has played on two teams they will show up three times, the stats for each team and the total stats throughout the season.

```{r}
tidy_shooting <- tidy_shooting %>%
  group_by(player) %>%
  slice_max(order_by = games, n = 1, with_ties = FALSE) %>%
  ungroup()
```

```{r}
#check code above
tidy_shooting %>%
  count(player) %>%
  filter(n > 1)
```

```{r}
tidy_adjusted_shooting <- tidy_adjusted_shooting %>%
  group_by(player) %>%
  slice_max(order_by = g, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-g)
```

## combine shooting and adjusted shooting

```{r}
nba_shooting_2025 <-
  tidy_shooting %>%
  left_join(tidy_adjusted_shooting, by = "player") %>%
  slice(-361) %>%
  mutate(
    across(3:12, as.numeric)
  )
```

## Check: combine wingspan data with shooting data

```{r}
nba_wingspans_2025 %>%
  left_join(nba_shooting_2025, by = c("name" = "player"))
```

```{r}
# Check with fuzzyjoin (don't need this)
library(fuzzyjoin)

stringdist_join(
nba_wingspans_2025, nba_shooting_2025,
by = c("name" = "player"),
method = "osa",
max_dist = 2,
mode = "left"
) %>%
  select(-player) %>%
  # show players that it doubles up
  count(name) %>%
  filter(n > 1)

# don't think fuzzy join is possible for this dataset.
```

# Combine wingspan and shooting data

About 15 names don't match between datasets, so change them in the wingspan dataset. Also a couple of players didn't play at all during the 2024-2025 season so remove them from the wingspan dataset.

```{r}
#nba_wingspans_shooting_2025 <- 
  nba_wingspans_2025 %>%
  # names that dont match between datasets
  mutate(
    name = if_else(name == "Alperen Sengun", "Alperen Şengün", name),
    name = if_else(name == "Danté Exum", "Dante Exum", name),
    name = if_else(name == "GG Jackson", "GG Jackson II", name),
    name = if_else(name == "Jimmy Butler III", "Jimmy Butler", name),
    name = if_else(name == "Monté Morris", "Monte Morris", name),
    name = if_else(name == "Robert Williams III", "Robert Williams", name),
    name = if_else(name == "Ronald Holland II", "Ron Holland", name),
    name = if_else(name == "Tristan da Silva", "Tristan Da Silva", name),
    name = if_else(name == "Xavier Tillman", "Xavier Tillman Sr.", name)
  ) %>%
  # players that didn't play in 2024-2025 season
  # 85, 320, 360
  #slice(-85,-320,-360) %>%
  # check combining
  left_join(nba_shooting_2025, by = c("name" = "player"))
  
```

```{r}
write_csv(nba_wingspans_shooting_2025, "partial_wingspan_shooting.csv")
```

# 3. From basketball reference, per game, per 100 poss and advanced tables

```{r}
# per 36 min table
url4 <- "https://www.basketball-reference.com/leagues/NBA_2025_per_minute.html"
objs4 <-
  url4 %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

per_min <- objs4[[1]]
```

```{r}
# per 100 poss table
url5 <- "https://www.basketball-reference.com/leagues/NBA_2025_per_poss.html"
objs5 <-
  url5 %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

per_100_poss <- objs5[[1]]
```

```{r}
# advanced table
url6 <- "https://www.basketball-reference.com/leagues/NBA_2025_advanced.html"
objs6 <-
  url6 %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

advanced <- objs6[[1]]
```

## general tidying and selecting rows

```{r}
# clean up per 36 min
tidy_per_min <- 
  per_min %>%
  clean_names() %>%
  select(2, 6, 24, 26, 27, 30) %>%
  group_by(player) %>%
  slice_max(order_by = g, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-g)
```

```{r}
# clean up per 100 poss
tidy_per_100_poss <- per_100_poss %>%
  clean_names() %>%
  select(2, 6, 32) %>%
  group_by(player) %>%
  slice_max(order_by = g, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-g)
```

```{r}
# clean up advanced
tidy_advanced <- advanced %>%
  clean_names() %>%
  select(2, 6, 22, 26) %>%
  group_by(player) %>%
  slice_max(order_by = g, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(-g)
```

## combine per game, per 100 poss and advanced

```{r}
nba_defense <- tidy_per_min %>%
  left_join(tidy_per_100_poss, by = "player") %>%
  left_join(tidy_advanced, by = "player") %>%
  slice(-361)
```

# 4. Combine all data, wingspan and shooting data with defensive data

```{r}
full_data <- 
  nba_wingspans_shooting_2025 %>%
  left_join(nba_defense, by = c("name" = "player")) %>%
  drop_na() %>%
  filter(minutes >= 440) %>%
  relocate(fg_percent, .after = `3pt_rate`) %>%
  relocate(ft_percent, .after = `3pt_percent`) %>%
  relocate(pts, .after = ts_percent)
```

## csv

```{r}
write_csv(full_data, "nba_wingspan_stats_2025.csv")
```

```{r}
 completeData <- read_csv(here::here("colney/nba_tidying/exploration/nba_full_stats_2025.csv"))
```


# 5. From NBA, deflections data

```{r}
# hustle table
url7 <- "https://www.nba.com/stats/players/hustle?SeasonType=Regular+Season"
objs7 <-
  url7 %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table()

deflections <- objs7[[2]]
```

# 6. Look for relationships in the data

```{r}
full_data %>%
  ggplot(., aes(x = wingspan_advantage)) +
  geom_density()
```

```{r}
full_data %>%
  filter (minutes > 1500) %>%
  filter(position == "C") %>%
  ggplot(., aes(x = wingspan_advantage, y = trb)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE)
```
