---
title: "ump_question_prep"
format: html
---
# Libraries

```{r}
library(tidyverse)
library(readr)
library(car)
```

# Data

```{r}
umpires_df <- read_csv("~/Desktop/git/score_new/awsmit22/mlb_umpires_module/stat113/exploration/umpires.csv")
```

# Question 1

Describe the distribution of `k_pct` based on this histogram.

```{r}
ggplot(data = umpires_df, aes(x = k_pct)) + geom_histogram(bins = 20, colour = "black", fill = "lightblue") + theme_minimal() + labs(x = "Strikeout Percentage", title = "Histogram of Strikeout Percentage") + theme(plot.title = element_text(hjust = 0.5))
```

**Strikeout percentage is left skewed, meaning that is more common for umpires to call a high number of strikeouts.**

# Question 2

What is the mean `k_pct` for all umpires? 

```{r}
mean(umpires_df$k_pct)
```

**The mean strike out percentage for all umpires is 20.8%**

# Question 3

What is the standard deviation of `k_pct` for all umpires? 

```{r}
sd(umpires_df$k_pct)
```

**The standard deviation of strike out percentage for all umpires is 2.03**

# Question 4

Calculate the IQR for `k_pct` using the summary statistics provided below. 

```{r}
summary(umpires_df$k_pct)

iqrk <- 22.2 - 19.9
iqrk
```

**The IQR for strike out percentage is 2.3**

# Question 5

The below table shows the 25 umpires with the lowest strikeout percentages in the dataset. Use the IQR method to determine if there are any outliers.

```{r}
umpires_df |> group_by(Umpire) |> summarise(k_pct = mean(k_pct)) |> arrange(desc(k_pct)) |> tail(n = 25) |> print(n = 25)
```

```{r}
1.5*iqrk 
lower_outk <- 22.2-3.45
lower_outk
```

**All of the umpires in this list except for the top 11 are outliers as they are all less than `Q1-1.5IQR`.**

# Question 6

Describe the distribution of `bb_pct` based on this histogram.

```{r}
ggplot(data = umpires_df, aes(x = bb_pct)) + geom_histogram(bins = 20, colour = "black", fill = "lightblue") + theme_minimal() + labs(x = "Walk Percentage", title = "Histogram of Walk Percentage") + theme(plot.title = element_text(hjust = 0.5))
```

**Walk percentage is right skewed, meaning that is more common for umpires to call a low number of walks.**

# Question 7

What is the mean `bb_pct` for all umpires? 

```{r}
mean(umpires_df$bb_pct)
```

**The mean walk percentage for all umpires is 8.48%.**

# Question 8

What is the standard deviation of `bb_pct` for all umpires? 

```{r}
sd(umpires_df$bb_pct)
```

**The standard deviation of walk percentage for all umpires is 0.68.**

# Question 9

Calculate the IQR for `bb_pct` using the summary statistics provided below. 

```{r}
summary(umpires_df$bb_pct)

iqrw <- 8.7 - 8.1
iqrw
```

**The IQR for walk percentage is 0.6**

# Question 10

The below table shows the 10 umpires with the highest walk percentages in the dataset. Use the IQR method to determine if there are any outliers.

```{r}
umpires_df |> group_by(Umpire) |> summarise(bb_pct = mean(bb_pct)) |> arrange(desc(bb_pct)) |> head(n = 10)
```

```{r}
1.5*iqrw 
upper_outw <- 8.7+0.9
upper_outw
```

**All of the umpires in this list except for Chris Tiller are outliers as they are all more than `Q3+1.5IQR`.**




