---
title: "Positions in the United Rugby Championships: Understanding Indicator Variables | SOLUTIONS"
author: "Phoebe Jones"
format: html
---

```{r}
library(tidyverse)
library(ggplot) ## for graphing
library(dplyr) ##

players <- read.csv("phoebechar/rugby/module/URCpositions.csv")
```

### The following scatterplot shows height of players being used as a predictor for weight. The graph has a separate line for each position.
```{r}
ggplot(players, aes(x = Height, y = Weight, colour = Position)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
### In the graph, we can see that some positions have very different lines. To make a model that accounts for this, we will have to use indicator variables.

### There are 6 positions present in the data set, so we need 5 indicator variables. The code block below defines 4 of them: the indicators for prop, hooker, lock, and scrum-half. For each of them, the observation will be assigned a `1` if that is the player's position, and a `0` otherwise.

```{r}
players <- players |> mutate(prop_ind = if_else(Position == "Prop", 1, 0),
                                     hook_ind = if_else(Position == "Hooker", 1, 0), 
                                     lock_ind = if_else(Position == "Lock", 1, 0), 
                                     scrum_ind = if_else(Position == "Scrum-half", 1, 0))
```

### Define the 5th indicator variable (for `Fullback`).
```{r}
players <- players |> mutate(full_ind = if_else(Position == "Fullback", 1, 0))
```

### Fit a model that uses `Height` as a predictor for `Weight` and includes all of the indicator variables. Print a summary of the model.
```{r}
modPos = lm(Weight ~ Height + prop_ind + hook_ind + lock_ind + scrum_ind + full_ind, data = players)

summary(modPos)
```

### What does the summary tell you? Is it consistent with the scatterplot?

The summary shows

### Is it helpful to have indicator variables for ALL of the different positions in this dataset? Why or why not?

Answers may vary. Students should point out that according to both the scatterplot and the model summary, the scrum-hal, fly-half, and fullback positions have very similar lines and most likely do not need separate indicator variables.

### Regardless of your answer, refit the model without the indicator variables for scrum-half and fullback. Instead we will leave them grouped in with the fly-halves. We will call this group the Backs.
```{r}
modPos2 = lm(Weight ~ Height + prop_ind + hook_ind + lock_ind, data = players)

summary(modPos2)
```

```{r}
players <- players |> mutate(Position2 = if_else(Position %in% c("Fullback", "Scrum-half", "Fly-half"), "Backs", Position))

ggplot(players, aes(x = Height, y = Weight)) +
  geom_point(aes(colour = Position)) +
  geom_smooth(aes(colour = Position2), method = "lm", se = FALSE)
```

INTERPRETATIONS