---
title: "nba_handout_key"
author: "eric seltzer"
format: html
---

# Introduction
You are an NBA GM for the Boston Celtics, one of the most historic teams of all time. You have acquired access to a data set that includes statistics and salary information for players ranging from 1990-2023. You're goal is to successfully visualize this data set, and determine a way to evaluate which type of players you want on your team. Ideally, you want players that play above what they are being paid and therefore are undervalued. 

# Question 1
Create visualizations to look at the distribution of minutes played and salary.
```{r}
library(tidyverse)
```

```{r}
nba <- read_csv(here::here("esselt21/nba_salary_module/module/initial_nba.csv"))
```

```{r}
ggplot(data = nba,
       aes(x = MP)) +
  geom_histogram(color = "black",
                 fill = "ivory")
```

```{r}
ggplot(data = nba,
       aes(x = salary)) +
  geom_histogram(color = "black",
                 fill = "ivory")
```

# Question 2
```{r}
nba <-
  nba |>
  mutate(salary = salary / 1000)
```

```{r}
ggplot(data = nba,
       aes(x = salary)) +
  geom_histogram(color = "black",
                 fill = "ivory")
```

```{r}
nba <-
  nba |>
  mutate(inflationAdjSalary = inflationAdjSalary / 1000)
```

```{r}
nba <-
  nba |>
  mutate(loginflationsalary = log(inflationAdjSalary))
```

```{r}
ggplot(data = nba,
       aes(x = loginflationsalary)) +
  geom_histogram(color = "black",
                 fill = "ivory")
```

