---
title: "kenduskeag_graphs"
format: html
---

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)

kenduskeag_full <- read_csv("phoebechar/multirace/full_kenduskeag.csv")
```

#### 1. Make a bar graph comparing counts of the different boat types.

```{r}
ggplot(kenduskeag_full, aes(x = boatType)) +
  geom_bar()
```

##### a. Change the color and fill of the bars to any combination you like

```{r}
ggplot(kenduskeag_full, aes(x = boatType)) +
  geom_bar(colour = "darkslategrey", fill = "darkseagreen")
```

##### b. Make a horizontal version of the same graph

```{r}
ggplot(kenduskeag_full, aes(y = boatType)) +
  geom_bar(colour = "darkslategrey", fill = "darkseagreen")
```

#### 2. Make a histogram showing the distribution of times.

```{r}
ggplot(kenduskeag_full, aes(x = time)) +
  geom_histogram()
```

##### a. Change the number of bins (test out a few options!)

```{r}
ggplot(kenduskeag_full, aes(x = time)) +
  geom_histogram(bins = 17)
```

##### b. Add a frequency line

```{r}
ggplot(kenduskeag_full, aes(x = time)) +
  geom_histogram() +
  geom_freqpoly()
```

```{r}
ggplot(kenduskeag_full, aes(x = courseID)) +
  geom_bar(position = "fill", aes(fill = boatType))
```

```{r}
kenduskeag_attend <- kenduskeag_full |> group_by(year, bib) |> slice(1) |> ungroup() |> group_by(year) |> summarise(n_boats = n())
```

```{r}
ggplot(kenduskeag_attend, aes(x = year, y = n_boats)) +
  geom_col()
```

#### Make a pink violin plot showing time differences based on the boat type.

```{r}
ggplot(kenduskeag_full, aes(x = boatType, y = time)) +
  geom_violin(color = "hotpink", fill = "pink")
```
