---
title: "MLB Pitching and Batting Styles Solutions"
format: docx
---

Major League Baseball (MLB) is often described as a sport of matchups. Every plate appearance is a one-on-one battle between a pitcher and a batter, and no two players approach that battle in exactly the same way. Some pitchers attack hitters with velocity and strikeouts, while others rely on control and limiting walks. On the other side of the ball, some batters swing for power, others prioritize making contact, and only a few manage to do both. Modern baseball statistics make it easier than ever to quantify these differences. With hundreds of pitchers and hitters competing across a full season, we can use performance data to identify trends and patterns that define different types of players. 

In this activity, we'll use data from the 2024 MLB season and apply K-means clustering, an unsupervised learning method that groups players based on similar statistical profiles. Rather than focusing on individual players, our goal is to understand how players vary, and whether they fall into distinguishable groups based on their hitting or pitching performance. For example, can we statistically separate power hitters from contact hitters? Are there clear differences between high-strikeout pitchers and control specialists?

By the end of this activity, you'll use K-means clustering to uncover common player archetypes, providing insight into the many different roles and skill sets found across the MLB.

# 0. Load the Following Packages

```{r}
library(tidyverse)
library(cluster)
library(factoextra)
```

# 1. Load in Pitching Dataset

Load in the `mlb_pitching_2024.csv` file from the data folder. Each row represents a pitcher from the 2024 regular season who faced at least 502 batters. Call the new object `pitching`.

```{r}
pitching <- read_csv(here::here("colney/mlb_k-means/data_repo_submission/mlb_pitching_2024.csv"))
```

# 2. Exploring Pitching Dataset

Pitchers in this dataset are described by their strikeout rate (`k_rate`) and walk rate (`bb_rate`).

- `k_rate`: The proportion of batters a pitcher struck out (e.g., 0.27 = 27%)
- `bb_rate`: The proportion of batters a pitcher walked (e.g, 0.08 = 8%)

These two stats give us a sense of a pitcher's general approach: do they rely on overpowering hitters with strikeouts, or do they focus on control and limiting walks?

Create a scatterplot of `k_rate` vs. `bb_rate` to look for patterns or trends. Do you see any groups forming in the plot?
```{r}
ggplot(pitching, aes(x = k_rate, y = bb_rate)) + 
  geom_point()
```

Solution: In the scatterplot, there appears to be a distinct group with high strikeout rates and low walk rates, as well as another group with high strikeout and high walk rates. Pitchers with low strikeout rates tend to be more mixed together.

# 3. Introduction to K-mean Clustering
So far, you've explored how MLB pitchers differ in terms of strikeout rate and walk rate. You may have noticed natural groupings in the scatterplot. K-means clustering is a method that helps us formally identify those kinds of groups in the data. 

K-means is an unsupervised learning method, which means it doesn't use any labels or outcomes to guide the analysis. Instead, it groups observations, in this case its pitchers, based purely on how similar they are across the variables we choose (like `k_rate`, `bb_rate`). 

Here's how K-means clustering works using the MLB pitching dataset:

1. You choose the number of groups/clusters (K) you want the algorithm to find. For example, K = 4 means you're asking for 4 groups of similar pitchers.

2. The algorithm randomly picks K starting points in the plot. These are called centers (or centroids) and these are the initial guesses for where the "middle" for each group might be.

3. Each pitcher is assigned to closest center based on distance, pitchers with similar strikeout and walk rates will end up near the same center.

4. The center of each group is then recalculated as the average position of all pitchers in that group.

5. Steps 3-4 repeat until the groups stop changing. 

At the end, each pitcher is assigned to a cluster with others who have similar pitching profiles. For example, one group might contain high strikeout, high walk pitchers, while another contains low strikeout, low walk pitchers. 

K-means clustering is useful when:

- You want to group observations based on similarities, without needing labels. 

- You're interested in exploring variation and structure in the data.

- You want to check whether natural groupings support your intuition.

- You're working with numerical data and want a simple, interpretable method to identify patterns. 

a. Based on the scatterplot from question 2, how many clusters do you think might make sense for this pitching data?

Solution: Based on the scatterplot, two clusters are clearly separated, and there's a larger group that could possibly be split further. So, choosing 3 or 4 clusters seems reasonable.

b. Why might it be useful to group pitchers using a method like K-means instead of relying on just one stat like strikeout rate?

Solution: K-means allows us to consider multiple variables at the same time, in this case, both strikeout and walk rate. So we get a more complete picture of a pitcher's style. It helps uncover different combinations of strikeout and walk rate, to group similar pitchers together. 

# 4. Prepare Data for Clustering
Before running the K-means algorithm, we need to prepare the data. This involves selecting the numeric variables we want to use and scaling them so that both variables are on the same scale.

a. Select Numerical Variables

K-means clustering requires only numeric input. In this case, we'll use just two variables, `k_rate` and `bb_rate`.Use `select()` to create a new data frame called `pitching_data` with just these two columns.
```{r}
pitching_data <- pitching %>%
  select(3, 4)
```

b. Scale the Data

K-means clustering is sensitive to the scale of the variables, so it's important to standardize them before clustering. The process is called scaling, and in R, the `scale()` function does this automatically by converting each value to a z-score

This ensures that both `k_rate` and `bb_rate` are measured on the same scale, so one doesn't unfairly influence the clustering results just because its original values were larger or more spread out. 

Use the `scale()` function to create a new data frame called `pitching_scaled` of standardized values.
```{r}
pitching_scaled <- scale(pitching_data)
```

c. Choosing the Number of Clusters

K-means requires you to specify the number of clusters (K) ahead of time. To decide what value to use, there are two common methods to select the optimal number of clusters.

- The Elbow Method: This looks at the total within-cluster variation. As the number of clusters increases, the variation within each group naturally decreases, but eventually the improvement slows down.

You're looking for a bend or "elbow" in the plot, this is the point where adding more clusters doesn't significantly reduce the variation. Here is the code to create the plot.
```{r}
fviz_nbclust(pitching_scaled, kmeans, method = "wss")
```

- The Silhouette Method: This measures how well each point fits within its assigned cluster. It calculates a silhouette score for each observation, which reflects how similar it is to points in its own cluster versus other clusters. 

Higher scores mean better defined clusters. You're looking for the number of clusters that has the highest average silhouette width. Here is the code to create the plot.
```{r}
fviz_nbclust(pitching_scaled, kmeans, method = "silhouette")
```

Based on the two plots, a reasonable number of clusters for the pitching data appears to be 4 or 5. To keep things simple, we'll move forward using 4 clusters.

# 5. Applying K-means
Now that you've selected a number of clusters (K=4), it's time to run the K-means algorithm and assign each pitcher to a cluster. 

a. Run the K-means Algorithm
The code below applies K-means clustering to the scaled pitching data.

*Note:* The `set.seed()` function is used to make the results reproducible. K-means starts with random cluster centers, so setting a seed ensures that the randomization produces the same result each time you run the code.
```{r}
set.seed(123)

k_pitching <- kmeans(pitching_scaled, centers = 4, nstart = 25)
```

What's Happening in the `kmeans` function:
- centers = 4, tells the algorithm to find 4 clusters

- nstart = 25, tells R to run the algorithm 25 times with different random starting points and keep the best result and keep the best result.

b. Add Cluster Labels to the Original Data
Now that the algorithm has assigned each pitcher to a cluster, we'll add this information to the original dataset, so we can visualize and interpret the clusters. The code below does this.
```{r}
pitching_clustered <- pitching %>%
  mutate(cluster = factor(k_pitching$cluster))
```

c. Visualize the Clusters
Use the `pitching_clustered` data table to plot `k_rate` vs `bb_rate`, and color each point by the assigned cluster.
```{r}
pitching_clustered %>%
  ggplot(., aes(x = k_rate, y = bb_rate, color = cluster)) +
  geom_point()
```

d. What patterns do you notice in how the clusters are separated in the scatterplot? Are any clusters more distinct or overlapping than others?

Solution: There appears to be two clusters that are clearly separated, cluster 3 and 4. Clusters 1 and 2 overlap a bit, where pitchers have moderate values for both stats. Overall, the clusters show distinct pitching styles.

# 6. Interpreting the Pitching Clusters  
Based on the strikeout and walk rates, each cluster represents a different type of pitcher. Here's one way to interpret the four clusters.

- Cluster 1 (Red): Struggling Pitchers
  - low `k_rate`, high `bb_rate`
  - These pitchers don't strike out many batters and give up a lot of walks. This often leads to poor results because they're not getting outs efficiently and are allowing many free base runners. This group often bounce between the majors and minors leagues or move to bullpen roles, as they struggle to consistently compete against top-level hitters.

- Cluster 2 (Green): Control Specialists
  - low `k_rate`, low `bb_rate`
  - These pitchers don't rely on strikeouts but rarely walk hitters either. They succeed by throwing strikes and letting the defense make plays, also known as "pitch to contact". You'll often find veterans or crafty pitchers in this group who rely on movement, timing, precision rather than velocity.

- Cluster 3 (Blue): Elite Pitchers
  - high `k_rate`, low `bb_rate`
  - These are the top performers. They strike out a lot of batters and rarely issue walks, meaning they control the game by keeping hitters off balance and off base. This is the group that earns $100+ million contracts and become All-Stars. Teams build around these kinds of pitchers.
  
- Cluster 4 (Purple): Fireballers
  - high `k_rate` high `bb_rate`
  - These pitchers rack up strikeouts but also walk more batters than average. They're aggressive and hard to hit, but their lack of control can lead to risky innings. Some can develop into elite pitchers, but others become high-risk, high-reward pitchers. Capable of blowing hitters away or losing the strike zone entirely. 
  
# 7. Clustering MLB Batters
Now that you've practiced K-means clustering with the pitching dataset, it's time to apply the same process to the batting dataset. 

The main difference this time is that you're working with four numeric variables instead of just two: `batting_avg`, `slg_percent`, `hr_per_hit` and `k_rate`.

This means you'll be clustering in 4-dimensional space, which makes it harder to visualize directly, but you can still explore relationships using scatterplots of two variables at a time.

a. Load in the `mlb_batting_2024.csv` file
```{r}
batting <- read_csv(here::here("colney/mlb_k-means/data_repo_submission/mlb_batting_2024.csv"))
```

b. Select numeric variables
```{r}
batting_data <- batting %>%
  select(3, 4, 5, 6)
```

c. Scale the data
```{r}
batting_scaled <- scale(batting_data)
```

d. Use the Elbow and Silhouette methods to choose the number of clusters
```{r}
fviz_nbclust(batting_scaled, kmeans, method = "wss")

fviz_nbclust(batting_scaled, kmeans, method = "silhouette")
```

e. Run K-means
```{r}
k_batting <- kmeans(batting_scaled, centers = 4, nstart = 25)
```

f. Add cluster labels
```{r}
batting_clustered <- batting %>%
  mutate(cluster = factor(k_batting$cluster))
```

g. Visualize the clusters

Since this K-means analysis used four variables, we can't fully visualize the clusters in one plot. However, we can still explore how the groups differ by looking at scatterplots of two variables at a time. 

Create a scatterplot of `k_rate` vs. `slg_percent`, coloring the points by cluster. This shows the trade-off between power and contact, highlighting whether hitters with more extra-base power tend to strike out more.
```{r}
batting_clustered %>%
  ggplot(., aes(x = k_rate, y = slg_percent, color = cluster)) +
  geom_point()
```

Create a scatterplot of `batting_avg` vs. `hr_per_hit`, coloring the points by cluster. This helps compare overall contact ability and home run power, revealing if hitters with high batting averages also generate a lot of power. 
```{r}
batting_clustered %>%
  ggplot(., aes(x = batting_avg, y = hr_per_hit, color = cluster)) +
  geom_point()
```

Even though we're only viewing two dimensions at a time, these plots still give useful insights into how the clusters differ. Keep in mind that the clustering happened in a 4D space, so the groups may not look perfectly separated in every 2D plot. 

f. Visualizing clusters in 2D using PCA
Since we're clustering using four variables, it's challenging to see the full picture in a single scatterplot. One common method for visualizing high-dimensional data in 2D is called Principal Component Analysis (PCA).

PCA creates new variables (called principal components) that capture the most important variation in the data. While we won't dive into the full details of how PCA works, we can use it create a 2D plot that shows all four-variables clusters in one view.

The code below displays the K-means clustering results based on the PCA transformed data. `k_batting` and `batting_scaled` are placeholders for the object names, use the object names you created earlier in section 7.
```{r}
fviz_cluster(k_batting, data = batting_scaled,
             geom = "point", ellipse.type = "convex")
```

This gives a simplified picture of how well the clusters are separated overall. It's not a perfect view, but it's a helpful summary when working with more than two variables. 

# 8. Interpreting the Batting Clusters
Based on the scatterplots using batting average, slugging percentage, home runs per hit and strikeout rate, each cluster represents a different type of batter. Here's one way to interpret the four clusters. 
  
- Cluster 1 (Red): Struggling Hitters
  - below average `batting_avg`, low `slg_percent`, below average `hr_per_hit`, average `k_rate`
  - These batters tend to struggle at the plate and offer limited offensive value. They're often young players adjusting to MLB pitching, defensive specialists, or bench players.
  
- Cluster 2 (Green): Contact Hitters
  - high `batting_avg`, average `slg_percent`, below average `hr_per_hit`, low `k_rate`
  - These hitters focus on putting the ball in play and rarely strike out. They may not hit many home runs, but they're consistent and tough outs. They may be leadoff hitters and they help generate offense by getting on base.
  
- Cluster 3 (Blue): Elite Hitters
  - high `batting_avg`, high `slg_percent`, high `hr_per_hit`, average `k_rate`
  - These players are the total package at the plate, great at making contact and generating power. They're often All-Stars, typically bat second or third in the lineup, and frequently appear on MVP ballots.
  
- Cluster 4 (Purple): Power Hitters
  - average `batting_avg`, average `slg_percent`, high `hr_per_hit`, high `k_rate`
  - These players swing hard and aim for home runs, even if it means striking out more. They're dangerous at the plate, especially with runners on base. They can change a game with one swing, often batting cleanup. 
  
# 9. Conclusion
In this activity, you were introduced to K-means clustering, an unsupervised learning method used to find groups within data based on similarity. Rather than labeling players manually or relying on traditional categories, K-means allowed us to explore how pitchers and batters naturally cluster based on their performance.

By applying K-means to the MLB pitching and batting dataset, we identified distinct player archetypes such as power hitters, contact hitter, elite pitchers, and control specialists, using only a few carefully chosen variables. While baseball was the context, the same clustering method can be used in many fields to reveal hidden structure and uncover meaningful patterns.
