---
title: "2025 Women's Track Running: Fixing Non-Linear Data Transformation"
format: html
    embed-resources: true
---

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)

track <- read.csv("WomensTrack2025.csv")
```

1. Make a scatterplot using distance (Event) to predict speed

```{r}
ggplot(track, aes(x = Event, y = Speed)) +
  geom_point()
```

1.1. What do you notice (think of model conditions)?

The data does not match a linear trend. There is clear curvature in the data. Speed seems to at first, dramatically decrease as distance increases, but the speed at which it decreases slows as distance continues to grow.

One way we can adapt to this is by applying a transformation to the data.

2. Based on the model's issue, apply a transformation to the data to make a linear model more appropriate.

```{r}
track <- track |> mutate(ln_distance = log(Event))
```

3. Make a scatterplot showing the transformed relationship.

```{r}
ggplot(track, aes(x = ln_distance, y = Speed)) +
  geom_point()
```

4. Define a model using the transformed data and print a summary.

```{r}
mod_track <- lm(Speed ~ ln_distance, track)
summary(mod_track)
```

5. Interpret in context, the coefficient on the transformed distance term.

As the natural log of distance increases by 1, Speed will, on average, decrease by 0.92 meters/second.