---
title: "Climbing: Data Wrangling and Data Visualization Worksheet Solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
  docx:  default
  pdf:  default
---


Read in dataset and necessary packages: 
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
Climbing <- read_csv("frhinehart/Climbing/Module/Climbing.csv")
```

1) Calculate the mean total points scored across the 10 countries. What is the median of the points?

```{r}
mean_total <- mean(Climbing$Points)
median_total <- median(Climbing$Points)

mean_total
median_total
```

`r round(mean_total, 3)`
`r round(median_total, 3)`

2) Find the standard deviation of the total points. What does this tell you about the spread of performance across the countries?

```{r}
sd_total <- sd(Climbing$Points)
sd_total
```

`r round(sd_total, 3)`

There is a very high spread in total points across the countries. They vary a lot. 

3) Which location appears to have the most competitive scores (i.e., the smallest range across countries)?

```{r}
location_ranges <- sapply(Climbing[ , 4:9], function(Points) max(Points) - min(Points))
most_competitive_location <- names(location_ranges)[which.min(location_ranges)]

location_ranges
most_competitive_location
```

Bern was the most competitive location

4) Which country had the most balanced performance (i.e., the smallest standard deviation in their six location scores)? What might a low standard deviation indicate about a country’s consistency?

```{r}
row_sd <- apply(Climbing[ , 4:9], 1, sd)
most_balanced_country <- Climbing$Name[which.min(row_sd)]

row_sd
most_balanced_country
```

Austria had the most balanced performance. This indicates they preformed relatively the same across each location. 

5) Create a new dataframe that has all locations for each country in one column and score for that location in another column. 

```{r}
long_data <- Climbing |>
  pivot_longer(cols = Keqiao:Innsbruck, 
               names_to = "Location", 
               values_to = "Score")
View(long_data)
```

6) Create a bar graph showing the total points for each country and color by location. What are the top three countries with the most points overall?

```{r}
ggplot(long_data, aes(x = Name, y = Points, fill = Location)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Points by Country", x = "Country", y = "Points") +
  theme_minimal()
```

Japan, France, and United Kingdom have the most points overall. 

7) Which locations contributed the most to each of Japan, France, and United Kingdom's scores? Report how much contribution these locations had on the overall score in percent form. 

```{r}
TopThree <- long_data |> slice(1:18)

TopThreePercent <- TopThree |>
  mutate(Contribution_Percent = round((Score / Points) * 100, 1))
View(TopThreePercent)
```

Japan - Salt Lake City - 19.5%
France - Prague - 22.7%
United Kingdom - Innsbruck - 25.9%

