---
title: "NFL Point Spreads - Solutions"
author: "R. Lock"
format: pdf
---

Solutions to 20 questions from the NFL Point Spreads and Game Scores module.

```{r}
#| message: false
#| warning: false
library(Hmisc)
```

Get the data set.

```{r}
NFLPoints<-read.csv("NFLPoints.csv")
```

**1. What proportion of the time does the favored team actually win the game outright?**

Procedure: CI for a proportion (one sample)

Via formula

```{r}
x=sum(NFLPoints$FavWin)
n=length(NFLPoints$FavWin)
phat=x/n
SE=sqrt(phat*(1-phat)/n)
zstar=qnorm(0.975)
ME=zstar*SE
lower=phat-ME
upper=phat+ME
```

$\hat{p} =$ `r x`/`r n`=`r round(phat,3)` with SE=`r round(SE,4)`

CI for p is `r round(phat,3)` $\pm$ `r round(zstar,3)` $\cdot$ `r  round(SE,4)` = (`r round(lower,3)`,`r round(upper,3)`)

We are 95% sure that the proportion of times the favored team wins the game is between 0.641 and 0.719.

Using an internal R function 
```{r}
prop.test(x,n,correct=FALSE)
```

Note that the method used for this CI is slightly different than the formula-based method above. 

**2. Is the chance the favorite covers the spread discernibly different from 0.50?**

Procedure: Test for proportion (one sample)

$H_0:p=0.5$ vs. $H_a:p\ne 0.5$ where $p$ is the proportion of games the favored team covers the spread

Via formula

```{r}
x=sum(NFLPoints$FavCover)
n=length(NFLPoints$FavCover)
phat=x/n
SE=sqrt(phat*(1-phat)/n)
z=(phat-0.5)/SE
pvalue=2*pnorm(-abs(z))
```

$\hat{p} =$ `r x`/`r n`=`r round(phat,3)` with SE=`r round(SE,4)`

z=(`r round(phat,3)` - 0.5)/`r round(SE,4)`=`r round(z,2)`, p-value=`r round(pvalue,3)`

At a 5% significance level we do not have convincing evidence that the proportion of times the favorite covers the spread is different from 0.50.


Using an internal R function 
```{r}
prop.test(x,n, correct=FALSE)
```

Note that the R prop.test function uses a chi-square statistic, rather than a z-statistic, but the square root of that chi-square statistic should match the z-statistic up to the +/- sign. 


**3.	How different is the average point spread when the favored team is playing at home as compared to the favorite playing on the road? **

Procedure: CI for a difference in means (two samples)

```{r}
(HomePts<-t.test(Pts~HomeDog, data=NFLPoints))
```

The difference in means is $\overline{x}_H-\overline{x}_A= 5.119-4.176=0.943$ and the 95% confidence interval for the difference in means is (`r round(HomePts$conf.int[1],2)`,`r  round(HomePts$conf.int[2],2)`).

We are 95% sure that the average point spread when the  home team is favored is between 0.44 and 1.44 more than when the road team is favored.

**4.	Many football fans say that the home field advantage is about a field goal (three points). Is the average home margin (*HomeDiff*) discernibly different from three points?**

Procedure: Test for a mean (one sample using *HomeDiff*) OR
Test for two means (paired data, using *HomeScore* and *AwayScore*)


$H_0:\mu=3$ vs. $H_a:\mu \ne 3$ where $\mu$ is the mean home field margin

```{r}
t.test(NFLPoints$HomeDiff,mu=3)
```

The mean home field advantage is $\overline{x}_d=$

