---
title: "Powerlifting Competition Data Analysis"
---

0. Packages

```{r}
library(dplyr)
library(readr)
library(fread)
```


1. Read in data

Source: This page uses data from the OpenPowerlifting project, https://www.openpowerlifting.org.

You may download a copy of the data at https://data.openpowerlifting.org.


```{r}
url_source <- "https://openpowerlifting.gitlab.io/opl-csv/files/openpowerlifting-latest.zip"

# manually download the file and load it (it is very large)

# do not store the full data on GitHub!

opl_data <- read_csv("local path")
```

2. Reduce size of data. 

a. Keep only SBD: Squat-Bench-Deadlift, also commonly called "Full Power" events

b. Keep only "Raw" equipment

c. Drop the following columns: Event, Equipment, BirthYearClass, Squat1Kg, Squat2Kg, Squat3Kg, Squat4Kg, Bench1Kg,  Bench2Kg, Bench3Kg, Bench4Kg, Deadlift1Kg, Deadlift2Kg, Deadlift3Kg, Deadlift4Kg, Tested, State, Country, Federation, ParentFederation, MeetState, MeetState, MeetTown, MeetName

d. Rename Best3BenchKg, Best3SquatKg, and Best3DeadliftKg so that the Best3 part is not in the name


# Questions

1. Chi-Square Testing: Is there an association between male and female participants and age category?

2. Correlations/Regressions - subset to just a specific age group for females and follow the Ironman style module

3. Stat 213 style regression: May require pivoting the data and subsetting to have just M and F lifters from the same age group (pick one of the groups in the 24 to 49 categoricies  as they all have decent balance between M & F)

+ Goal is to understand/measure differences between male and female powerlifters.

+ Build a model of 

Full model:

WeightLifted ~ Type_of_Lift * Sex * Bodyweight

(i.e., all the interactions!)

WeightLifted (numerical) and Type_of_Lift (categorical) are new variables constructed after pivoting

Will likely need to work through a series of nested models to measure the difference between males and females. (My suspicion is that, after accounting for bodyweight, there shouldn't be much, if any, difference in weight lifted for Males or Females)

