---
title: "Untitled"
format: html
editor: visual
---

```{r}
library(readr)
men_and_women_2022_2023 <- read_csv("~/stat_289/stat_289_score/aj_modules/college_lacrosse/lacrosse_indicator_module/data/men_and_women_2022_2023.csv")
```
```{r}
mod_faceoffs = lm(win_loss_pctg~faceoff_win_pct, data = men_and_women_2022_2023 )
plot(win_loss_pctg~faceoff_win_pct, data = men_and_women_2022_2023)
abline(mod_faceoffs)
```

```{r}
summary(mod_faceoffs)
```



```{r}
library(dplyr)
women <- men_and_women_2022_2023 %>%
  filter(Division == "Womens")

men <- men_and_women_2022_2023 %>%
  filter(Division == "Mens")
  
```
```{r}
mod_women = lm(win_loss_pctg~faceoff_win_pct, data = women )
mod_men = lm(win_loss_pctg~faceoff_win_pct, data = men )
plot(win_loss_pctg~faceoff_win_pct, data = men_and_women_2022_2023)
abline(mod_women, col = "red")
abline(mod_men, col = "blue")

```

```{r}
lax <- men_and_women_2022_2023%>%
  mutate(genderIND = ifelse(Division == "Mens", 1, 0))
```


```{r}
summary(lm(win_loss_pctg~faceoff_win_pct+genderIND, data = lax))
```
Indicator coefficient is essentially zero. This makes sense because someone HAS to win the faceoff no matter the gender. The R^2 value also didn't change.


```{r}
summary(lm(win_loss_pctg~faceoff_win_pct+genderIND+faceoff_win_pct*genderIND, data = lax))
```
womens: win_loss_pctg = -0.7291 + 2.4478*faceoff_win_pct
faceoff win pctg of 0 has a win/loss pctg of -0.7291. For every 1 pctg increase in faceoffs wins, win loss pctg goes up by 2.4478


mens: win_loss_pctg = -0.257 + 1.5024*faceoff_win_pct
faceoff win pctg of 0 has a win/loss pctg of -0.257. For every 1 pctg increase in faceoffs wins, win loss pctg goes up by 1.5024

Yes this model is better. R^2 goes up

The lines cross at about where faceoff pctg is 50%.

```{r}
cor(women$shots_per_game, women$faceoff_win_pct)
cor(men$shots_per_game, men$faceoff_win_pct)
# there's a higher correlation between faceoff win pctg and winning the game for women
```

```{r}
mean(women$shot_pctg)
mean(men$shot_pctg)
# women score more often
```






