---
title: "rowing_start"
format: html
---

I am playing around with this Olympics data set I found on [Kaggle](https://www.kaggle.com/code/kalilurrahman/olympic-games-eda){}. I want to look specifically at rowing and medals. The data set contains data from the 1896 Olympics to the 2022 Olympics. 

# Loading the packages and reading in the data

```{r}
library(tidyverse)
library(readr)
```

```{r}
athletes_df <- read_csv("/Users/abigailsmith/Desktop/stat289/awsmit22/rowing_module/athlete_events.csv")
```

# Playing with the data

This table has the total number of medals per nation in rowing with na values ommitted

```{r}
rowing_df <- athletes_df |> filter(Sport == "Rowing")
```

```{r}
rowingsum <- rowing_df |> mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(NOC) |> summarise(total_medals = sum(n_medal))
```

```{r}
rowing_sum <- rowing_df |> filter(NOC == "USA" & !is.na(Medal)) |> mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(NOC, Medal) |> summarise(total_medals = sum(n_medal))
```

```{r}
ggplot(data = rowing_sum, aes(x = Medal, y = total_medals)) + geom_boxplot()
```

List of all the rowing events that the US competed in between 1896 and 2022. 

```{r}
rowing_df |> filter(NOC == "USA") |> group_by(Event) |> summarise()
```

Something strange is that after properly dividing up the medals, it seems that the is an uneven number i figured out it was because of this https://www.olympedia.org/athletes/40630, I found Durbrow's ID number and filtered the dataset so that he would not be included, making the number even.

I looked it up and in the olympics a team sport medal counts as one medal, not the number of medals of the athletes on the team so I divided the medals by the number of rowers in each event so that the medals would be more accurate.

```{r}
events_df <- rowing_df |> filter(NOC == "USA" & ID != 30817) |> mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(Event) |> summarise(total_medals = sum(n_medal)) |> mutate(total_medals = case_when(Event == "Rowing Women's Coxed Eights" ~ total_medals/9, Event == "Rowing Men's Coxed Eights" ~ total_medals/9, Event == "Rowing Men's Coxed Fours" ~ total_medals/5, Event == "Rowing Men's Coxed Pairs" ~ total_medals/3, Event == "Rowing Men's Coxed Fours" ~ total_medals/5, Event == "Rowing Men's Coxless Fours" ~ total_medals/4, Event == "Rowing Men's Coxless Pairs" ~ total_medals/2, Event == "Rowing Men's Double Sculls" ~ total_medals/2, Event == "Rowing Men's Lightweight Coxless Fours" ~ total_medals/4, Event == "Rowing Men's Lightweight Double Sculls" ~ total_medals/2, Event == "Rowing Men's Quadruple Sculls" ~ total_medals/4, Event == "Rowing Men's Single Sculls" ~ total_medals/1, Event == "Rowing Women's Coxed Fours" ~ total_medals/5, Event == "Rowing Women's Coxed Quadruple Sculls" ~ total_medals/5, Event == "Rowing Women's Coxless Fours" ~ total_medals/4, Event == "Rowing Women's Coxless Pairs" ~ total_medals/2, Event == "Rowing Women's Double Sculls" ~ total_medals/2, Event == "Rowing Women's Lightweight Double Sculls" ~ total_medals/2, Event == "Rowing Women's Quadruple Sculls" ~ total_medals/4, Event == "Rowing Women's Single Sculls" ~ total_medals/1))
```

```{r}
rowing_df |> mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(Event, NOC, Year) |> summarise(total_medals = sum(n_medal)) |> mutate(total_medals = case_when(Event == "Rowing Women's Coxed Eights" ~ total_medals/9, Event == "Rowing Men's Coxed Eights" ~ total_medals/9, Event == "Rowing Men's Coxed Fours" ~ total_medals/5, Event == "Rowing Men's Coxed Pairs" ~ total_medals/3, Event == "Rowing Men's Coxed Fours" ~ total_medals/5, Event == "Rowing Men's Coxless Fours" ~ total_medals/4, Event == "Rowing Men's Coxless Pairs" ~ total_medals/2, Event == "Rowing Men's Double Sculls" ~ total_medals/2, Event == "Rowing Men's Lightweight Coxless Fours" ~ total_medals/4, Event == "Rowing Men's Lightweight Double Sculls" ~ total_medals/2, Event == "Rowing Men's Quadruple Sculls" ~ total_medals/4, Event == "Rowing Men's Single Sculls" ~ total_medals/1, Event == "Rowing Women's Coxed Fours" ~ total_medals/5, Event == "Rowing Women's Coxed Quadruple Sculls" ~ total_medals/5, Event == "Rowing Women's Coxless Fours" ~ total_medals/4, Event == "Rowing Women's Coxless Pairs" ~ total_medals/2, Event == "Rowing Women's Double Sculls" ~ total_medals/2, Event == "Rowing Women's Lightweight Double Sculls" ~ total_medals/2, Event == "Rowing Women's Quadruple Sculls" ~ total_medals/4, Event == "Rowing Women's Single Sculls" ~ total_medals/1, Event == "Rowing Men's 17-Man Naval Rowing Boats" ~ total_medals/17, Event == "Rowing Men's 6-Man Naval Rowing Boats" ~ total_medals/6, Event == "Rowing Men's Coxed Fours, Inriggers" ~ total_medals/5,  Event == "Rowing Men's Coxed Fours, Outriggers" ~ total_medals/5, Event == "Rowing Men's Coxed Pairs (1 kilometres)" ~ total_medals/3, Event == "Rowing Men's Coxed Pairs (1 mile)" ~ total_medals/3))
```

```{r}
nozeros <-rowingsum |> filter(total_medals !=0)
```

I am thinking to ignore the Olympics rules on the medals and just look at the medals data as is. 

# Working on coming up with some simple questions

Sumamry statistics

```{r}
summary(nozeros$total_medals)
```

```{r}
rowingyear <- rowing_df |> mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(Year) |> summarise(total_medals = sum(n_medal))
```

```{r}
ggplot(data = rowingyear, aes(x = Year, y = total_medals)) + geom_line()
```

```{r}
usaonly <- rowing_df |> filter(NOC == "USA") |>mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(Year) |> summarise(total_medals = sum(n_medal))
```

```{r}
ggplot(data = usaonly, aes(x = Year, y = total_medals)) + geom_line()
```

```{r}
year_noc <- rowing_df |> mutate(n_medal= case_when(Medal == "Gold" ~ 1, Medal == "Silver" ~ 1, Medal == "Bronze" ~ 1, is.na(Medal) ~ 0)) |> group_by(Year, NOC) |> summarise(total_medals = sum(n_medal))
```

```{r}
ggplot(data = year_noc, aes(x = Year, y = total_medals, colour = NOC)) + geom_line()
```

making progress
