---
title: "NCAA 2022-2023 Lacrosse"
output: html_document
date: "2023-07-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
lax <- read_csv("data/lax_2022_2023.csv")

ggcorr(lax, # update data here
       
       #suggestion: keep these options the same
       label = TRUE, label_size = 3, 
       label_round = 2, label_alpha = TRUE,
       hjust = 0.75, layout.exp = 1
       )
```



# proportion model for shot/ FO win percentages - sum of whether or not the team wins the faceoff

## What is the variable of interest? 
The proportion of faceoffs won

## The true population proportion for the 2022-2023 season
```{r}
mean(lax$FO_win_pctg)
```
The true population proportion for faceoff wins in the 2022-20223 season is .501 (meaning they are won roughly 50% of the time).

## In the 2022-2023 season, Duke won 330 out of 548 total faceoffs
# Hypothesis 

Ho: p = .501
Ha: p > .501 

## Prepare: What is the sample size and sample proportion 
The sample size is 548 and the sample proportion is 0.602

# Check the success-failure condition 
nPo = 548(.501) = 274.548 > 10
n(1-Po) = 548(1-.501) = 273.452 > 10 




# Calculate the Z test statistic  
z = .602 - .501/ sqrt(.501*.499/548) = 4.728708


# Calculate the p-value at a significance level of alpha = .05 

Does this sample proportion provide significant evidence that Duke has a greater faceoff 
percentage than what is normal for all of Division 1 lacrosse in the 2022-2023 season?
Provide all necessary details and a conclusion in context.


```{r}
cor(lax$faceoff_win_pct, lax$win_loss_pctg)
```



