---
title: "racewalk_question_prep"
format: html
---

# Load in the libraries

```{r}
library(tidyverse)
library(readr)
library(broom)
```

# Load in the prepped data

```{r}
racewalkingdf <- read_csv("~/Desktop/git/stat_289_score/awsmit22/racewalking/data_repo_submission/racewalking.csv")

racewalking_df <- racewalkingdf |> mutate(DISTANCE = as.factor(DISTANCE))
```

# Question 1

Compare the distribution of SPEED for the 10k races with the distribution for the 20k races.

```{r}
ggplot(racewalking_df, aes(x = SPEED)) + geom_histogram(bins = 15, colour = "black", fill = "lightblue") + facet_wrap(~DISTANCE) + theme_minimal()
```

# Question 2

Perform a t.test to determine if there is a significant difference in the mean speed for each distance.

```{r}
t.test(SPEED~DISTANCE, var.equal=TRUE, data=racewalking_df)
```

# Question 3

Based on the below boxplot would you say that there is a significant difference in the mean speed for each distance and for each gender?

```{r}
ggplot(racewalking_df, aes(y = SPEED, x = DISTANCE, fill = DISTANCE)) + geom_boxplot(show.legend = FALSE) + facet_wrap(~GENDER) + theme_minimal() + scale_fill_brewer() 
```

# Question 4

Use a two-way ANOVA test to determine if there is a significant difference in the mean speed for each distance and for each gender?

```{r}
amodA <- aov(SPEED~DISTANCE,data=racewalking_df)
summary(amodA)

amodB <- aov(SPEED~GENDER,data=racewalking_df)
summary(amodB)

amodC <- aov(SPEED~DISTANCE+GENDER,data=racewalking_df)
summary(amodC)
```

# Question 5

Fit the model: REC = DISTANCE + PODIUM + SPEED + GENDER. 

```{r}
# a)
mod_rec <- glm(REC ~ DISTANCE + PODIUM + SPEED + GENDER, family = "binomial", data = racewalking_df)
mod_rec |> tidy()

# b)
# i)
log_odds <- -0.255-(1.27*0)-(1.98*1)+(0.00729*221)-(0.123*1)
log_odds

# ii)
odds <- exp(-0.74691)
odds

# iii)
pihat <- 0.4738284/(1+0.4738284)
pihat
```


A female athlete walked in the 10k with a podium finish and a speed of 221 meters per minute. Calculate the log(odds), odds, and predicted odds of her making a record. 

# Question 6

Perform a g-test to assess the overall fit of the multiple logistic regression model: REC = DISTANCE + PODIUM + SPEED + GENDER 

```{r}
record_mod <- glm(REC ~ DISTANCE + PODIUM + SPEED + GENDER, family = "binomial", data = racewalking_df)
record_mod |> summary()

# g-stat)
g_stat <-319.17 - 291.97
g_stat

# p-value)
1-pchisq(g_stat,4)
```
