---
title: "Cleaning"
output: html_document
date: "2023-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(GGally)
```


# There is a significantly higher correlation between faceoff win percentage and overall win percentage for women than there is for men

# Shot percetage on goal also has a strong correlation with win percentage 

```{r}
draw <- read_csv("draw_pctg.csv", col_select = c(1,4))
turnovers <- read_csv("turnovers.csv", col_select = 1:2)
sog <- read_csv("sog.csv", col_select = 1:2)
assists <- read_csv("assists.csv", col_select = 1:2)
caused_turnovers <- read_csv("caused_turnovers.csv", col_select = 1:2)
clearing_pctg <- read_csv("clearning_pctg.csv", col_select = 1:2)
draws_per_game <- read_csv("draws_per_game.csv", col_select = 1:2)
fouls <- read_csv("fouls.csv", col_select = 1:2)
free_position_pctg <- read_csv("free_position_pctg.csv", col_select = 1:2)
ground_balls <- read_csv("ground_balls.csv", col_select = 1:2)
points <- read_csv("points.csv", col_select = 1:2)
save_pctg <- read_csv("save_pctg.csv", col_select = 1:2)
goals_allowed <- read_csv("goals_allowed.csv", col_select = 1:2)
goals_per_game <- read_csv("goals_per_game.csv", col_select = 1:2)
shots <- read_csv("shots.csv", col_select = 1:2)
wins <- read_csv("win_pctg.csv", col_select = 1:2)
shot_pctg <- read_csv("shot_pctg.csv", col_select = 1:2)

w_lax <- left_join(draw, turnovers, by = "Team")
w_lax <- left_join(w_lax, sog, by = "Team")
w_lax <- left_join(w_lax, assists, by = "Team")
w_lax <- left_join(w_lax, caused_turnovers, by = "Team")
w_lax <- left_join(w_lax, clearing_pctg, by = "Team")
w_lax <- left_join(w_lax, draws_per_game, by = "Team")
w_lax <- left_join(w_lax, fouls, by = "Team")
w_lax <- left_join(w_lax, free_position_pctg, by = "Team")
w_lax <- left_join(w_lax, ground_balls, by = "Team")
w_lax <- left_join(w_lax, points, by = "Team")
w_lax <- left_join(w_lax, save_pctg, by = "Team")
w_lax <- left_join(w_lax, goals_allowed, by = "Team")
w_lax <- left_join(w_lax, goals_per_game, by = "Team")
w_lax <- left_join(w_lax, shots, by = "Team")
w_lax <- left_join(w_lax, shot_pctg, by = "Team")
w_lax <- left_join(w_lax, wins, by = "Team")

cor(w_lax$draw_pctg, w_lax$win_pctg)
cor(w_lax$sog_per_game, w_lax$win_pctg)
```

```{r}
mod <- lm(win_pctg ~ sog_per_game + draw_pctg, data = w_lax)
summary(mod)
```


# Idea: Have students compare the correlation of of faceoff_win_pctg to win_pctg for men and women
# Its higher for women being the faceoffs are more meaningful.

# Then, why could this be?
# look at histograms of turnovers per game for men and women 

```{r}
write_csv(w_lax, file = "womens_lax_2022-2023.csv")

w_lax <- read_csv("womens_lax_2022-2023.csv")


# Correlation is higher for women than men 
cor(w_lax$draw_pctg, w_lax$win_pctg)
cor(lax$faceoff_win_pct, lax$win_loss_pctg)
```
```{r}
# 
hist(lax$avg_turnovers)
hist(w_lax$turnovers_per_game)
```
```{r}
mean(lax$avg_turnovers)
mean(w_lax$turnovers_per_game)

# not crazy
```



```{r}
ggcorr(w_lax, # update data here
       
       #suggestion: keep these options the same
       label = TRUE, label_size = 3, 
       label_round = 2, label_alpha = TRUE,
       hjust = 0.8, layout.exp = 1
       )
```

