
```{r}
library(tidyverse)
```

```{r}
ncaa_gymnastics_data <- read.csv("/Users/vivianjohnson/Desktop/stat_289_score_new/vgjohn/gymnastics_module/Exploration/ncaa_gymnastics.csv")
```

```{r}
ncaa_gymnastics_data %>%
  ggplot(mapping = aes(x = Total)) +
  geom_histogram()
```
```{r}
avg_by_year <- 
  ncaa_gymnastics_data %>%
  group_by(Year) %>%
  summarise(
    meanVault = mean(Vault),
    meanBar = mean(Bar),
    meanBeam = mean(Beam),
    meanFloor = mean(Floor),
    meanTotal = mean(Total)
    )

avg_by_year
```
```{r}
# line graph 
ggplot(
  data = avg_by_year,
  mapping = aes(x = Year,
                y = meanTotal)
) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = avg_by_year$Year)

# smooth 
ggplot(
  data = avg_by_year,
  mapping = aes(x = Year,
                y = meanTotal)
) +
  geom_point() +
  geom_smooth() +
  scale_x_continuous(breaks = avg_by_year$Year)
```
confidence intervals and random samples 
proportion? difference in proportions? 

```{r}
ncaa_gymnastics_data %>% group_by(Team) %>% count(Rank)

```
look at random sampling with the scores for different events , standard deviation 

scatterplot of 
```{r}
ggplot(data = ncaa_gymnastics_data,
       mapping = aes(x = Beam, y = Bar)) +
  geom_point()
```
each event correlation matrix with the other events 
```{r}
cor(x = ncaa_gymnastics_data[, c(4:7)])
```
not very strong correlations between events

regression analysis to see how much each event contributes to the total 
```{r}
reg_total <- lm(Total ~ Vault + Bar + Beam + Floor,
                data = ncaa_gymnastics_data)

summary(reg_total)
```
can see that bar and beam probably contribute more to the total score than floor and vault because the coefficients are higher

interesting because bar and beam had the highest correlation between events. 

bootstrap samples and confidence intervals for the mean scores on each event 

```{r}
# mean scores for each first place team 
ncaa_gymnastics_data %>% group_by(Rank) %>%
  summarise(
    mean_Vault = mean(Vault),
    sd_Vault = sd(Vault),
    mean_Bar = mean(Bar),
    sd_Bar = sd(Bar),
    mean_Beam = mean(Beam),
    sd_Beam = sd(Beam),
    mean_Floor = mean(Floor),
    sd_Floor = sd(Floor),
    mean_Total = mean(Total),
    sd_Total = sd(Total)
  )
```

want to create a bootstrap samples for the mean using data from all the number one ranke teams 
```{r}
ncaa_gymnastics_data %>% filter(Rank == 1) %>% select(Total)
```

