---
title: "cleaning"
format: html
---

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(lubridate)
```

We will begin by reading the data in. To save us from having to do it later, as you read in the data, add a line to read the `Date` column in as a day-month-year date (from the `lubridate` package) and the `Discipline` and `Sex` columns as factors as they both have limited options (this data only contains F for female and M for male).

```{r}
diving <- read_csv("continentalDiving.csv") |> mutate(Sex = as.factor(Sex), Discipline = as.factor(Discipline), Date = dmy(Date))
```


This data contains results from the time-based and distance-based freediving disciplines. This means that the `Result` and `Announced` columns contain both distance measurements and time measurements. Because of this, the easiest way to work with this data is to split it into two data sets, one with distance disciplines and one with the static discipline. Write code to do that.

```{r}
static <- diving |> filter (Discipline == "STA")

distance <- anti_join(diving, static)
```

For the static data, we will use the `lubridate` package to put the `Result` and `Announced` columns into formats easier to work with. Ultimately, we want to have this data in the `duration` form, but it is currently character and cannot be converted right away. First, we must use a function from the `lubridate` package to convert the data into minute-second time data.

```{r}

```

Now, the data can be converted into a `duration` class.

For the distance data, the `Result` and `Announced` columns contain 'm's to indicate the unit, but this forces the data to read in as characters instead of integers like we want. We must parse the numbers.


------------------------

All data: separate country and name, date as dmy, Gender and Discipline as factors, add prop_over(?) and perc_best(?), Ranks (??)

Distance: parse numbers from Result and Announced

Static: read as ms, as duration (note unit)


