---
title: "Centering Predictors to Improve Interpretations of the Intercept"
author: Stat 213 (Chapman)
format: docx
editor: visual
---

# Load packages

```{r}
library(readr)  # to read in data
library(dplyr)  # to make a new variable
library(ggplot2) # for graphs
library(broom)   # for residuals
```

# Example 1: Velvet Ants Revisited

### 1. Read in the Velvet Ant data

```{r}
stings = read_csv("VelvetAntStings.csv")
```

### 2. Recall our fitted model to predict sting length from mesosomal length, and the interpretation of the intercept.

In the above model, the intercept does not have a practically meaninful interpretation because "x=0" is outside the range of our dataset. We can make the interpretation meaningful by refitting the model with a "centered" predictor - that is, shifting the values of the explanatory variable so that "x=0" is within the range of our data.

### 3. What is the average mesosomal length?

```{r}
mean(stings$Mesosoma)
```

### 4. Create a new variable in our dataset named meso_centered by subtracting the average mesosomal length from each observation.

```{r}
stings = stings %>% 
  mutate(meso_centered = Mesosoma - mean(Mesosoma))
```

### 5. What is the average value of this new variable?

```{r}
mean(stings$meso_centered)
```

### 6. How would we interpret the mesosomal length of the first and second species in the dataset?

### 7. Fit a model that predicts sting length from centered mesosomal length. Write the fitted model below and interpret the slope and intercept of the model. What has changed and what has stayed the same?

```{r}
stings_mod2 = lm(Sting ~ meso_centered, data=stings)

summary(stings_mod2)
```

### 8. Use this model to predict the sting length for the first species in the dataset. What needs to be done differently in this situation?

# Example 2: Olympic Men's Figure Skating

### 1. Read in the Men's Figure Skating data

```{r}
skating = read_csv("MensFigureSkating.csv")
```

### 2. Fit a model that uses year to predict total points (TP). Write the fitted model below. Provide an interpretation of the intercept of the model.

```{r}

skating_mod1 = lm(TP ~ Year, data=skating)
summary(skating_mod1)


ggplot(skating, aes(x=Year,y=TP)) + 
  geom_point()+
  geom_smooth(method="loess",se=FALSE)
```

### 3. Consider the new variable being made in the code below. What is special about the number used to create the new variable? What would be an interpretation of the new variable? Would you rather create the new variable by subtracting mean(Year)? Why or why not? (Before we move on, we will give the new variable a better name.)

```{r}
skating = skating %>%
  mutate(YearsSince2006 = Year - 2006)
```

### 4. Fit a model that uses the new variable created above to predict TP. Provide an interpretation of the intercept of the model.

```{r}
skating_mod2 = lm(TP ~ YearsSince2006, data=skating)
summary(skating_mod2)
```

**Moral of this handout:**
