---
title: "powerlift_class"
author: "S. Weaver"
date: "2023-06-27"
output: html_document
---

## TODO: Change the labels and add titles to graphs
## TODO: Calculate summary stats of the histograms (ex. median and mean)
## TODO: Create worksheet and key (save graphs after they are completed)


 Correlations/Regressions - subset to just a specific age group for females and follow the Ironman style module

```{r}
library(readr)
library(here)
library(dplyr)
library(ggplot2)
library(ggthemes)

opl_class <- read_csv(here("powerlifting/data/powerliftingregression.csv"))
View(opl_class)
```

## new dataset that is just females who are 25 (original dataset was ages 25-49)
```{r}
opl_female <- opl_class |> filter(Sex == "F") |> filter(Age == 25.0)

write_csv(opl_female, "opl_female.csv")
```

## distribution of benchKg
```{r}
ggplot(data = opl_female, aes(x = BenchKg, )) +
  geom_histogram(colour="black", fill="pink") +
  theme_minimal() +
  labs(title = "Bench Weight for Female Power Lifting",
       x = "Bench Weight (in kilograms)",
       y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
summary(opl_female$BenchKg)
```


## distribution of SquatKg
```{r}
ggplot(data = opl_female, aes(x = SquatKg, )) +
  geom_histogram(colour="black", fill="pink") +
  theme_minimal() +
  labs(title = "Squat Weight for Female Power Lifting",
       x = "Squat Weight (in kilograms)",
       y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
summary(opl_female$SquatKg)
```


## distribution of DealiftKg
```{r}
ggplot(data = opl_female, aes(x = DeadliftKg, )) +
  geom_histogram(colour="black", fill="pink") +
  theme_minimal() +
  labs(title = "Deadlift Weight for Female Power Lifting",
       x = "Deadlift Weight (in kilograms)",
       y = "Count") +
  theme(plot.title = element_text(hjust = 0.5))
```
```{r}
summary(opl_female$DeadliftKg)
```

## BenchKg vs SquatKg
```{r}
ggplot(data = opl_female, aes(x = BenchKg, y = SquatKg)) +
  geom_point(color = "darkgreen") +
  theme_minimal() +
  labs(title = "Scatterplot of BenchKg vs SquatKg") +
  theme(plot.title = element_text(hjust = 0.5))
```
## Correlation Bench vs Squat
```{r}
test_1 <- cor.test(opl_female$BenchKg, opl_female$SquatKg)
test_1

## correlation = 0.8207528
```



## BenchKg vs DeadliftKg
```{r}
ggplot(data = opl_female, aes(x = BenchKg, y = DeadliftKg)) +
  geom_point(color = "darkgreen") +
  theme_minimal() +
  labs(title = "Scatterplot of BenchKg vs DeadliftKg") +
  theme(plot.title = element_text(hjust = 0.5))
```
## Correlation Bench vs Deadlift
```{r}
test_2 <- cor.test(opl_female$BenchKg, opl_female$DeadliftKg)
test_2

## correlation = 0.7824763
```


## SquatKg vs DeadliftKg
```{r}
ggplot(data = opl_female, aes(x = SquatKg, y = DeadliftKg)) +
  geom_point(color = "darkgreen") +
  theme_minimal() +
  labs(title = "Scatterplot of SquatKg vs DeadliftKg") +
  theme(plot.title = element_text(hjust = 0.5))
```
## Correlation Deadlift vs Squat
```{r}
test_3 <- cor.test(opl_female$DeadliftKg, opl_female$SquatKg)
test_3

## correlation = 0.8630972
```

