---
title: "Batting Worksheet"
author: "Brendan Karadenes"
format: html
---

                Multicollinearity Using Negro League Baseball Data
The Negro Leagues were made up of American professional baseball teams of African Americans during the late 19th century and into the 1940s. The leagues contained some of the best talent in baseball history and many of which had success in Major League Baseball when the league became integrated in 1947. In 2024, Negro League baseball data was officially incorporated into the record book for Major League Baseball giving credit to some of the game’s best players.
We will be investigating Negro League career batting statistics from 1896 to 1947. The data contains 502 position players (non-pitchers) that competed during this period. Our variables of interest include the number of runs scored, weighted on-base average, walk percentage, and isolated power. A player is credited with a run scored when they successfully make it around the bases and cross Homeplate. Weighted on-base average is a metric that takes into account each way a batter can get on base and gives higher weights to more productive outcomes (i.e. homeruns and doubles have more weight than singles and walks). Walk rate is simply the number of walks a batter has divided by their plate appearances. Isolated power is a metric that extra-base hits (doubles, triples, and homeruns) by the number of at-bats a player has, giving weights to the more productive outcomes. 
The goal of this worksheet is to see if there is multicollinearity between these metrics and if so, how strong it is. Also, we want to see which one is most correlated with runs scored.

Data Description
R – runs scored
wOBA – weighted on-base average
BB_rate – walk percentage
ISO – isolated power

1.	Before looking at any output, which of the variables (besides runs scored) in the data description above do you think will be most correlated with runs scored? Explain.

Sample Answer: I think weighted on-base average would be most correlated with runs scored because it takes into account a combination of getting on base and hitting for power. This makes it a more dynamic stat than walk rate and isolated power because it doesn't just focus on one aspect of a hitter's game.

```{r}
## Load the necessary packages and data for this worksheet
library(tidyverse)
batting <- ("ngl_batting.csv")
```

2. Let's consider three different models to predict runs scored.

a. Model 1: Walk percentage as a predictor of runs scored. Create the model and obtain the summary output.

```{r}
mod1 = lm(R ~ BB_rate, data = batting)
summary(mod1)
```

b. Model 2: Weighted on-base average as a predictor of runs scored. Create the model and obtain the summary output.

```{r}
mod2 = lm(R ~ wOBA, data = batting)
summary(mod2)
```

c. Model 3: Isolated Power as a predictor for runs scored. Create the model and obtain the summary output. 

```{r}
mod3 = lm(R ~ ISO, data = batting)
summary(mod3)
```

d. Now create a model that uses all three (isolated power, weighted on base average, and walk rate) as predictors for runs scored. Look at the summary output. Is there anything surprising about your results? How do the p-values differ from when you did each individually and why?

Model 4: Walk percentage, weighted on-base average, and isolated power as a predictor for runs scored

```{r}
mod4 = lm(R ~ BB_rate + wOBA + ISO, data = batting)
summary(mod4)
```

Sample Answer: The p-values are much higher than when I did each individually, likely due to correlations between the independent variables. Also, adding more variables changes the model itself, so it's likely that the model and the p-values change as well. It's surprising that the t-value for the predictor variables decreased by such a large amount due to adding the other variables in the model, perhaps due to correlation between them.

e. Install and load the package `GGally` and use the `ggpairs` function to construct a correlation matrix. What do you notice?

```{r}
## install.packages("GGally")
library(GGally)
ggpairs(data = batting, columns = c("R", "wOBA", "BB_rate", "ISO"))
```

Sample Answer: There is a high correlation between wOBA and ISO and correlations between all of the variables. This explains why the p-values and standard errors increased when we added them all together in one model. 

3. Compute the vif of the model with all of the predictor variables by hand and verify your answers by using the `vif()` function from the car package.

```{r}
modwOBA = lm(wOBA ~ R + BB_rate + ISO, data=batting)
modBB = lm(BB_rate ~ R + wOBA + ISO, data=batting)
modISO = lm(ISO ~ R + BB_rate + wOBA, data=batting)
```

```{r}
summary(modwOBA)
summary(modBB)
summary(modISO)
```

wOBA: 1/(1-0.79) = 4.762
BB_rate: 1/(1-0.2667) = 1.364
ISO: 1/(1-0.7528) = 4.045

```{r}
library(car)
vif(mod4)
```

4. Are any of these VIFs high? Explain.

Sample Answer: No, all of the VIFs are under 5, so they wouldn't be considered high indicating that the predictors have low correlations with the other predictors. In particular, walk rate has the lowest VIF, so it has the lowest correlation with the other predictor variables. This is good because lower VIFs mean the model is more likely to be more accurate.

5. Looking back at the correlation matrix you constructed in question 2e, why do you think isolated power and weighted on-base average had the highest correlation? Why do you think walk rate and isolated power had the lowest. Refer back the top of this file for detailed descriptions of the variables.

Sample Answer: Since both isolated power and weighted on-base average both give high weights to extra base hits, it makes sense that they would be highly correlated with each other. Walk rate and isolated power had the lowest correlation because they use different metrics to calculate their stats. Walk rate uses walks per plate appearance while isolated power only takes into account doubles triples and homeruns per at-bat. 

6. Of the three predictors, which do you think a player should focus on maximizing if they want to score more runs? Explain.

Sample Answer: They should want to maximize their weighted on-base average because that's the most highly correlated with runs scored. So, they should focus on combining high on base numbers with power not just one or the other.

7. Add a predictor variable that will bring at least one of the VIFs over 5 to the model with walk rate, weighted on-base average, and isolated power as predictors of runs. Explain what the variable was and why you think it made one of the VIFs increase.

Sample answer:

```{r}
modVIF = lm(R ~ wOBA + BB_rate + ISO + OPS, data=batting)
vif(modVIF)
```

The OPS variable had a high VIF and made the isolated power and weighted on-base average VIFs increase because it created a high correlation between the variables and raised the possibility that multicollinearity now exists in the model.

8. Brainstorm why multicollinearity can be a problem.

Multicollinearity can be a problem because it undermines the statistical significance of the predictor variables. Meaning, it reduces the precision of the model which makes the p-values untrustworthy.


