---
title: "Figure_skating_Solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
format: 
  html:
    embed-resources: true
---

Reading in the Dataset:
```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(rvest)
library(here)
library(dplyr)
Skating <- read_csv("frhinehart/Olympic Figure Skating/Module/MensFigureSkating.csv")
here::here
```

Question 1:

a) Fit a model that uses year to predict total points (TP). Write the fitted model below using the model summary output. 
```{r}
skating_mod1 = lm(TP ~ Year, data = Skating)
summary(skating_mod1)
```

$$\hat{TP} = -8594.3305 + 4.3806 \times Year $$

b) Provide an interpretation of the intercept of the model.

The intercept is -8594.3305 which means that when the year is zero the estimated total points is -8594.3305. 

c) Plot the model on a scatterplot and add a smoother. Comment on any trends you see on the graph.
```{r}
ggplot(data = Skating, aes(x = Year, y = TP)) + 
  geom_point()+
  geom_smooth(method ="loess", se = FALSE)
```

According to the smoother, the total points tend to increase as years increase and leveling off in 2022.  

Question 3: 

Consider the new variable being made in the code below. Use the internet to determine what is special about the number used to create the new variable? What would be an interpretation of the new variable? Would you rather create the new variable by subtracting mean(Year)? Why or why not? (Before we move on, we will give the new variable a better name.)
```{r}
Skating2 <- Skating |>
  mutate(YearsSince2006 = Year - 2006)
```

The number is the year when the new international point system was introduced. Scores are calculated based on difficulty and execution of each maneuver instead of the previous 6.0 scale in order to promote more of a fair judging system. 

The interpretation of the new variable is the number of years since 2006/the number of years after the new scoring system was introduced. 

Subtracting mean(Year) would center the data around zero, making the average year equal to zero. This does not provide the same understanding of changes relative to after the shift in point system. Using 2006 as a reference point offers a meaningful baseline for analyzing trends and changes in the sport after the new judging system implementation. 

Question 4: Fit a model that uses the new variable created above to predict TP. Provide an interpretation of the intercept of the model.

```{r}
skating_mod2 = lm(TP ~ YearsSince2006, data = Skating2)
summary(skating_mod2)
```

193.2033 is the intercept which means that in 2006 the estimated total points is 193.2033.

Question 5:

a) Fit a model that uses `SP` (Short program) to predict `FS` (Free Skating). Comment on the 4 model assumptions and if applicable, provide relevant plots to support your answer.

```{r}
model <- lm(SP ~ FS, data = Skating2)
plot(model)
```

1) Normal - No because there is significant amount of points that deviate from the fitted line on the qq plot (specifically the tails)
2) Independent - Yes. We can assume the data was collected randomly and independent from each other. 
3) Constant Variance - No because The residuals seem to be fanning outward towards the middle on the residual vs fitted plot
4) Linearity - No, there is a curve in the data points on the residual vs fitted

b) Based on the answers in part a what transformation should you preform on what variable?

A log transformation on the y variable

b) Preform a the transformation on and write down the fitted model below using the model summary output.

```{r}
model_log <- lm(log(SP) ~ FS, data = Skating2)
summary(model_log)
```

$$\hat{FS} = e^{3.591561} \times e^{.0052495SP} $$
c) Interpret $$\hat{\beta{_0}}$$ and $$\hat{\beta{_1}}$$. 

$$\hat{\beta{_0}}$$ - The predicted FS when SP = 0 is e^3.591561

$$\hat{\beta{_1}}$$ - For each one-unit increase in SP, the predicted FS increases by a factor of e^.0052495









a) Using the combined data set, create a scatter plot that uses SP to predict FS colored by sex. Comment on what you notice about the male and female scores. 
```{r}
library(dplyr)
url1 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2022_Winter_Olympics_%E2%80%93_Women%27s_singles"
tab1 <- read_html(url1) |>
  html_nodes("table") |> html_table()

womenSkate <- tab1[[8]]

url2 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2022_Winter_Olympics_%E2%80%93_Men%27s_singles"
tab2 <- read_html(url2) |>
  html_nodes("table") |> html_table()

menSkate <- tab2[[9]]

womenSkateTidy <- womenSkate[, c(2:4, 6, 8)] |> slice(1:24, 30) |> mutate(Sex = "Female") |> mutate(SP = as.numeric(SP)) |> mutate(FS = as.numeric(FS)) |> mutate(Total = as.numeric(Total))

menSkateTidy <- menSkate[, c(2:4, 6, 8)] |> slice(1:24) |> mutate(Sex = "Male") |> mutate(SP = as.numeric(SP)) |> mutate(FS = as.numeric(FS)) |> mutate(Total = as.numeric(Total))

combined_data <- bind_rows(womenSkateTidy, menSkateTidy)
```

```{r}
ggplot(combined_data, aes(x = SP, y = FS, color = Sex)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Free Skating vs. Short Program",
       x = "Short Program Score",
       y = "Free Skating Score") +
  theme_minimal()
```

Male short program scores tend to be higher than the female short program scores which lead to a higher free skating score than the females. 


```{r}
#2006 women data 
url3 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2006_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab3 <- read_html(url3) |>
  html_nodes("table") |> html_table()
womenSkate2006 <- tab3[[6]] 
anyDuplicated(names(womenSkate2006))
colnames(womenSkate2006) <- make.names(colnames(womenSkate2006), unique = TRUE)
womenSkate2006 <- womenSkate2006 |> mutate(Year = "2006")
#2010 women data 
url4 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2010_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab4 <- read_html(url4) |>
  html_nodes("table") |> html_table()
womenSkate2010 <- tab4[[7]] 
anyDuplicated(names(womenSkate2010))
colnames(womenSkate2010) <- make.names(colnames(womenSkate2010), unique = TRUE)
womenSkate2010 <- womenSkate2010 |> mutate(Year = "2010")
#2014 women data 
url5 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2014_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab5 <- read_html(url5) |>
  html_nodes("table") |> html_table()
womenSkate2014 <- tab5[[8]] 
anyDuplicated(names(womenSkate2014))
colnames(womenSkate2014) <- make.names(colnames(womenSkate2014), unique = TRUE)
womenSkate2014 <- womenSkate2014 |> mutate(Year = "2014")
#2018 women data 
url6 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2018_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab6 <- read_html(url6) |>
  html_nodes("table") |> html_table()
womenSkate2018 <- tab6[[8]] 
anyDuplicated(names(womenSkate2018))
colnames(womenSkate2018) <- make.names(colnames(womenSkate2018), unique = TRUE)
womenSkate2018 <- womenSkate2018 |> mutate(Year = "2018") |>
  rename(FS = "FS.1", FS.1 = "FS") |>
  rename(SP = "SP.1", SP.1 = "SP")
#2022 women data 
url7 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2022_Winter_Olympics_%E2%80%93_Women%27s_singles"
tab7 <- read_html(url7) |>
  html_nodes("table") |> html_table()
womenSkate2022 <- tab7[[8]] 
anyDuplicated(names(womenSkate2022))
colnames(womenSkate2022) <- make.names(colnames(womenSkate2022), unique = TRUE)
womenSkate2022 <- womenSkate2022 |> mutate(Year = "2022")

womenSkate2006$SP.1 <- as.character(womenSkate2006$SP.1)
womenSkate2010$SP.1 <- as.character(womenSkate2010$SP.1)
womenSkate2014$SP.1 <- as.character(womenSkate2014$SP.1)
womenSkate2018$SP.1 <- as.character(womenSkate2018$SP.1)
womenSkate2022$SP.1 <- as.character(womenSkate2022$SP.1)

allwomen_data <- bind_rows(womenSkate2006, womenSkate2010, womenSkate2014, womenSkate2018, womenSkate2022) |>
  mutate(`Total Points` = coalesce(TP, Total, `Total.points`, `Total.Points`)) |>
  select(-TP, -Total, -`Total.points`, -`Total.Points`) |>
  select(1:3, 5, 7, 8, 15) |>
  slice(1:24, 31:54, 65:85, 93:116, 124:147)
```

```{r}
ggplot(data = allwomen_data, aes(x = as.numeric(SP.1), y = as.numeric(FS.1), color = as.factor(Year))) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(data = Skating, aes(x = SP, y = FS, color = as.factor(Year))) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(data = allwomen_data, aes(x = as.numeric(SP.1), y = as.numeric(FS.1), color = as.factor(Year))) +
  geom_point() +
  geom_point(data = Skating, aes(x = SP, y = FS, color = as.factor(Year)))
  geom_smooth(method = "lm")
  
ggplot(data = Skating, aes(x = as.factor(Year), y = SP)) +
  geom_boxplot()

ggplot(data = Skating, aes(x = as.factor(Year), y = FS)) +
  geom_boxplot()

ggplot(data = allwomen_data, aes(x = as.factor(Year), y = as.numeric(SP.1))) +
  geom_boxplot()

ggplot(data = allwomen_data, aes(x = as.factor(Year), y = as.numeric(FS.1))) +
  geom_boxplot()
```

Did changing the scoring equalize the scoring

what happened in scoring between 2014 and 2018/why scoring higher
- Advancements in technical difficulty that led to higher scoring (athleticism over artistry) - Nathan chen landing 5 quads in FS
- evolution of skating styles to include more complex choreo and innovative elements
- The ISU continued to refine the criteria for PCS, which assess the artistic aspects of a performance, including choreography, interpretation, and presentation. Judges became more adept at distinguishing subtle differences in these areas, leading to more precise and often higher PCS, thereby increasing overall scores.
- The introduction of technical panels, consisting of specialists who identify and evaluate each element in a program, added a layer of precision to the scoring process. This system reduced human error and inconsistencies, leading to more accurate assessments and, in some cases, higher scores for well-executed elements. This system was in place during the 2014 Winter Olympics and continued through the 2018 Winter Olympics.

lowered in 2022 because
- After the 2018 Winter olympics the updated GOE scoring was introduced that ranged from -5 to 5 to allow a broader range of scores that better reflected the naunces in the execution of elements
- After the 2018 olympics the Zagitova Rule was introduced that limited the number of jump elements eligible for the 10% bonus in the second half of a program to encourage a more balanced distribution of technical elements throughout the program


what happened in nathen chen short program in 2018
 - faltered performance on every jump