---
title: "FigureSkatingOutlier_Solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
---

Read in the data set:
```{r}
library(ggplot2)
library(tidyverse)
library(here)
library(dplyr)
menSkating <- read_csv(here::here("frhinehart/Olympic Figure Skating/Module/MensFigureSkating.csv")) |>
  select(2:7) 
```

Question 1:

a) Create a scatter plot using short program score to predict free skate score and color by year. Add a smoother. What year has a distinct outlier?

```{r}
ggplot(data = menSkating, aes(x = SP, y = FS, color = as.factor(Year))) +
  geom_point() +
  geom_smooth(method = "lm")
```

Year 2018 has a distinct outlier

b) Filter the to only include that year you answered in part a. Rename the data set to a new name and recreate the scatter plot from above.

```{r}
menSkating2018 <- menSkating |> filter(Year == 2018)

ggplot(data = menSkating2018, aes(x = SP, y = FS, color = as.factor(Year))) +
  geom_point() +
  geom_smooth(method = "lm")
```

The outlier on the graph is Nathan Chen (which is observation 5 in the data set). Even with faltering every jump in his short program skate, his free skate score was record high because he preformed 6 quad jumps, landing 5 perfectly. The 2018 scores had a steeper slope than other years (shown in part a graphs) suggesting that it was due do Nathan Chen's performance. We can investigate if his performance was influenctial by looking at different outlier detectors.

We will first start will Standardized Residuals

## Standardized residuals:

The difference between an observed value and a predicted value in a regression model. It is calculated by Residual = Observed value – Predicted value. A residual that is larger than 2 or smaller than -2 is often flagged as an outlier.

Question 2:

a) Fit the linear model of the 2018 men skating data setand call it residualModel. Write down the model. 

```{r}
residualModel <- lm(FS ~ SP, data = menSkating2018)
residualModel
```

$$\hat{FS} = 1.508SP + 38.921$$
b) Plot residualModel to get diagnostic plots using the plot() function
```{r}
plot(residualModel)
```

We want to look at the Residuals vs Fitted plot. This gives us an idea of which observations have the highest residuals.

c) Which observation has the highest residual when looking at the graph?

Observation 5

d) Calculate the residuals of all the points the rstandard function. State the residual of the observation stated in part c.
```{r}
std_residuals <- rstandard(residualModel)
std_residuals
```

Residual of observation 5 is 3.247580568

e) Is the observation considered an outlier?

Yes because it is bigger than 2

This determines that Nathan Chen's performance is considered an outlier but we don't know yet if it is influential in the overall 2018 score trend. We will use Leverage and Cook distance to determine if Chen's performance was influential. 

## Leverage:

The extent to which the coefficients in the regression model would change if a particular observation was removed from the dataset. A point is considered to have high leverage if the leverage is greater than 2(p+1) divided by n. Where p is the number of predictors and n is the number of observations.

$$ >(2(p+1))/n $$

Question 3:

a) Fit the same model from question 2 but call it leverageModel.
```{r}
leverageModel <- lm(FS ~ SP, data = menSkating2018)
leverageModel
```

b) Calculate leverage for each observation using the hatvalues function.
```{r}
leverage_values <- hatvalues(leverageModel)
leverage_values
```

c) Plot leverageModel to get diagnostic plots using the plot() function
```{r}
plot(leverageModel)
```

We want to look at the Residuals vs Leverage plot. On this plot, it labels the three observations. If a observation lies outside the dashed lines (which is cook's distance) than that point is considered to be an influential point.

d) Based on the graph, is observation 5 a influential point?

No becuase it lies within the dashed lines therefore not influential.

e) Report observation's 5 leverage value from part b

The leverage value is 0.05206357.

f) Using the formula provided, determine if the calculation supports your answer in part e

(2(1+1))/ 4 = .16666

0.05206357 is not greater than .16666 so the calculation does support part e that observation 5 is not influential. 

We will now look at cook's distance as another way to determine a point's influence. 

## Cook's Distance:

Identifies influential observations that can significantly impact the model's coefficients and predictions. A point is considered moderatly influential if the cook distance is .5. Strongly influential is greater than 1. 

Question 4:

a) Fit a linear model and call it cookModel. 
```{r}
cookModel <- lm(FS ~ SP, data = menSkating2018)
cookModel
```

b) Calculate Cooks Distance using the cooks.distance function
```{r}
cooksD <- cooks.distance(cookModel)
cooksD
```

c) Plot the cooks distance to identify which point is has the highest cook's distance. State it's value.
```{r}
plot(cooksD, type = "h", main = "Cook's Distance", ylab = "Distance", xlab = "Index")
abline(h = 4 / length(cooksD), col = "red", lty = 2)
```

2.896307e-01

d) Does the value found in part c indicate a influential observation? Does it support the Residual vs. Leverage plot

No, observation 5 is not a influential point as it is less than .5. This does support the Residual vs. Leverage plot

You can also calculate cook's distance using the formala:

$$Dᵢ = (rᵢ² / (p * MSE)) * (hᵢᵢ / (1 - hᵢᵢ)²)$$
r: residual of the ith observation
p: number of predictors in model
hii: Levergae of the ith observation
MSE: Mean Squared Error of the model

e) Use the anova function to run anova on the cookModel to get the MSE value. 
```{r}
anova(cookModel)
```

f) Calculate by hand cook's distance

.66????

Overall, we have determine that Nathan Chen's performance was unusual compared to the other skater's however it was not influenital in the overall 2018 winter Olympic score trend. 
