---
title: "Update1"
author: Faith Rhinehart
format: revealjs
---

##Dataset

The two main datasets are of men and women's figure skating from the 2006 Winter Olympics to the 2022 winter Olympics. The data was scraped from Wikipedia and tidied into two separate datasets. 

The variables for each include:
1) Rank: The athletes overall rank compared to other competitors
2) Name: Athlete
3) Nation: Athlete's representing country
4) TP/Total Point: Total Score of Athlete (SP + FS)
5) SP/SP.1: Short program score
6) FS/FS.1: Free skate score
7) Year: Olympic Year

```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(rvest)
library(here)
library(dplyr)
#2006 women data 
url3 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2006_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab3 <- read_html(url3) |>
  html_nodes("table") |> html_table()
womenSkate2006 <- tab3[[6]] 
anyDuplicated(names(womenSkate2006))
colnames(womenSkate2006) <- make.names(colnames(womenSkate2006), unique = TRUE)
womenSkate2006 <- womenSkate2006 |> mutate(Year = "2006")
#2010 women data 
url4 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2010_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab4 <- read_html(url4) |>
  html_nodes("table") |> html_table()
womenSkate2010 <- tab4[[7]] 
anyDuplicated(names(womenSkate2010))
colnames(womenSkate2010) <- make.names(colnames(womenSkate2010), unique = TRUE)
womenSkate2010 <- womenSkate2010 |> mutate(Year = "2010")
#2014 women data 
url5 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2014_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab5 <- read_html(url5) |>
  html_nodes("table") |> html_table()
womenSkate2014 <- tab5[[8]] 
anyDuplicated(names(womenSkate2014))
colnames(womenSkate2014) <- make.names(colnames(womenSkate2014), unique = TRUE)
womenSkate2014 <- womenSkate2014 |> mutate(Year = "2014")
#2018 women data 
url6 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2018_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab6 <- read_html(url6) |>
  html_nodes("table") |> html_table()
womenSkate2018 <- tab6[[8]] 
anyDuplicated(names(womenSkate2018))
colnames(womenSkate2018) <- make.names(colnames(womenSkate2018), unique = TRUE)
womenSkate2018 <- womenSkate2018 |> mutate(Year = "2018") |>
  rename(FS = "FS.1", FS.1 = "FS") |>
  rename(SP = "SP.1", SP.1 = "SP")
#2022 women data 
url7 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2022_Winter_Olympics_%E2%80%93_Women%27s_singles"
tab7 <- read_html(url7) |>
  html_nodes("table") |> html_table()
womenSkate2022 <- tab7[[8]] 
anyDuplicated(names(womenSkate2022))
colnames(womenSkate2022) <- make.names(colnames(womenSkate2022), unique = TRUE)
womenSkate2022 <- womenSkate2022 |> mutate(Year = "2022")

womenSkate2006$SP.1 <- as.character(womenSkate2006$SP.1)
womenSkate2010$SP.1 <- as.character(womenSkate2010$SP.1)
womenSkate2014$SP.1 <- as.character(womenSkate2014$SP.1)
womenSkate2018$SP.1 <- as.character(womenSkate2018$SP.1)
womenSkate2022$SP.1 <- as.character(womenSkate2022$SP.1)

allwomen_data <- bind_rows(womenSkate2006, womenSkate2010, womenSkate2014, womenSkate2018, womenSkate2022) |>
  mutate(`Total Points` = coalesce(TP, Total, `Total.points`, `Total.Points`)) |>
  select(-TP, -Total, -`Total.points`, -`Total.Points`) |>
  select(1:3, 5, 7, 8, 15) |>
  slice(1:24, 31:54, 65:85, 93:116, 124:147)

Skating <- read_csv("frhinehart/Olympic Figure Skating/Module/MensFigureSkating.csv")
```


##Interesting Graphs part 1:
```{r}
library(ggplot2)
ggplot(data = allwomen_data, aes(x = as.numeric(Year), y = as.numeric(`Total Points`))) + 
  geom_point() +
  geom_smooth(method ="loess", se = FALSE) +
  labs(title = "Women's Total Points")
```

```{r}
ggplot(data = Skating, aes(x = Year, y = TP)) + 
  geom_point()+
  geom_smooth(method ="loess", se = FALSE) +
  labs(title = "Men's Total Points")
```

- we can notice that there seems to be an upward trend in athletes scoring higher throughout the years

##Interesting Graphs part 2:
```{r}
ggplot(data = allwomen_data, aes(x = as.numeric(SP.1), y = as.numeric(FS.1), color = as.factor(Year))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Women's SP versus FS Points")
```

```{r}
ggplot(data = Skating, aes(x = SP, y = FS, color = as.factor(Year))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Men's SP versus FS Points")
```

- Looking at the Men's graph we see the there is a spike in points between 2014 and 2018 than levels back down in 2022
- For the women, the spike is not as drastic but still there is an increase
- Why this sudden spike in the men?

##Athleticism over Artistry
- Athletes began to notice that the "risk is worth the reward" (even attempting a jump can lead to more points than artistry)
- Advancements in technical difficulty that led to a higher scoring 
  - Nathan Chen landed 5 Quad jumps in his FS which brought up overall scores (outlier on graph) even though his SP score was low due      to him faltering on every jump
- Evolution of skating styles to include more complex choreo and innovative elements
- Refined the criteria -> Judges became more adept at distinguishing subtle differences, leading to more precise and often higher overall scores
- The introduction of technical panels reduced human error and inconsistencies, leading to more accurate assessments and higher scores for well-executed elements

##After 2018
- Scores began to lower a bit in 2022 after policies were placed to "close the gap" between athleticism and artistry scoring
- Updated GOE scoring was introduced that ranged from -5 to 5 to allow a broader range of scores that better reflected the naunces in the execution of elements
- The Zagitova Rule was introduced that limited the number of jump elements eligible for the 10% bonus in the second half of a program to encourage a more balanced distribution of technical elements throughout the program

##What's Next
- Develop a question/storyline for the module to explore relating to these ideas
- Module will most likely be Stat 213 level
- ANOVA at the end to identify if the new regulations did in fact equalize the scoring

##Thank You
