---
title: "Update1"
author: Faith Rhinehart
format: revealjs
---

##Dataset

The two main datasets are of men and women's figure skating from the 2006 Winter Olympics to the 2022 winter Olympics. The data was scraped from Wikipedia and tidied into two separate datasets. 

The variables for each include:
1) Rank: The athletes overall rank compared to other competitors
2) Name: Athlete
3) Nation: Athlete's representing country
4) TP/Total Point: Total Score of Athlete (SP + FS)
5) SP/SP.1: Short program score
6) FS/FS.1: Free skate score
7) Year: Olympic Year

```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(rvest)
library(here)
library(dplyr)
#2006 women data 
url3 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2006_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab3 <- read_html(url3) |>
  html_nodes("table") |> html_table()
womenSkate2006 <- tab3[[6]] 
anyDuplicated(names(womenSkate2006))
colnames(womenSkate2006) <- make.names(colnames(womenSkate2006), unique = TRUE)
womenSkate2006 <- womenSkate2006 |> mutate(Year = "2006")
#2010 women data 
url4 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2010_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab4 <- read_html(url4) |>
  html_nodes("table") |> html_table()
womenSkate2010 <- tab4[[7]] 
anyDuplicated(names(womenSkate2010))
colnames(womenSkate2010) <- make.names(colnames(womenSkate2010), unique = TRUE)
womenSkate2010 <- womenSkate2010 |> mutate(Year = "2010")
#2014 women data 
url5 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2014_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab5 <- read_html(url5) |>
  html_nodes("table") |> html_table()
womenSkate2014 <- tab5[[8]] 
anyDuplicated(names(womenSkate2014))
colnames(womenSkate2014) <- make.names(colnames(womenSkate2014), unique = TRUE)
womenSkate2014 <- womenSkate2014 |> mutate(Year = "2014")
#2018 women data 
url6 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2018_Winter_Olympics_%E2%80%93_Ladies%27_singles"
tab6 <- read_html(url6) |>
  html_nodes("table") |> html_table()
womenSkate2018 <- tab6[[8]] 
anyDuplicated(names(womenSkate2018))
colnames(womenSkate2018) <- make.names(colnames(womenSkate2018), unique = TRUE)
womenSkate2018 <- womenSkate2018 |> mutate(Year = "2018") |>
  rename(FS = "FS.1", FS.1 = "FS") |>
  rename(SP = "SP.1", SP.1 = "SP")
#2022 women data 
url7 <- "https://en.wikipedia.org/wiki/Figure_skating_at_the_2022_Winter_Olympics_%E2%80%93_Women%27s_singles"
tab7 <- read_html(url7) |>
  html_nodes("table") |> html_table()
womenSkate2022 <- tab7[[8]] 
anyDuplicated(names(womenSkate2022))
colnames(womenSkate2022) <- make.names(colnames(womenSkate2022), unique = TRUE)
womenSkate2022 <- womenSkate2022 |> mutate(Year = "2022")

womenSkate2006$SP.1 <- as.character(womenSkate2006$SP.1)
womenSkate2010$SP.1 <- as.character(womenSkate2010$SP.1)
womenSkate2014$SP.1 <- as.character(womenSkate2014$SP.1)
womenSkate2018$SP.1 <- as.character(womenSkate2018$SP.1)
womenSkate2022$SP.1 <- as.character(womenSkate2022$SP.1)

allwomen_data <- bind_rows(womenSkate2006, womenSkate2010, womenSkate2014, womenSkate2018, womenSkate2022) |>
  mutate(`Total Points` = coalesce(TP, Total, `Total.points`, `Total.Points`)) |>
  select(-TP, -Total, -`Total.points`, -`Total.Points`) |>
  select(1:3, 5, 7, 8, 15) |>
  slice(1:24, 31:54, 65:85, 93:116, 124:147)

Skating <- read_csv("frhinehart/Olympic Figure Skating/Module/MensFigureSkating.csv")
```


##Interesting Graphs part 1:
```{r}
library(ggplot2)
ggplot(data = allwomen_data, aes(x = as.numeric(Year), y = as.numeric(`Total Points`))) + 
  geom_point() +
  geom_smooth(method ="loess", se = FALSE) +
  labs(title = "Women's Total Points")
```

```{r}
ggplot(data = Skating, aes(x = Year, y = TP)) + 
  geom_point()+
  geom_smooth(method ="loess", se = FALSE) +
  labs(title = "Men's Total Points")
```

- we can notice that there seems to be an upward trend in athletes scoring higher throughout the years
