---
title: "Horse Racing Correlation Worksheet Solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
  docx:  default
  pdf:  default
---

Read in the dataset and necessary packages:
```{r}
library(readr)
library(ggplot2)
horseRacing <- read_csv("horseRacing.csv", 
    col_types = cols(Finish = col_number(), 
        `Qualifying Points` = col_number()))
```

1) Make a scatter plot that has `Qualifying Points` on the x-axis and `Finish` on the y-axis. Add a smoother, give your graph a title, and comment on what you notice. Remember lower numbered finishes are better.

```{r}
ggplot(data = horseRacing, aes(x = `Qualifying Points`, y = Finish)) +
  geom_point() +
  geom_smooth(method = lm) +
  labs(title = "Qualifying Points vs Finished Position")
```

As qualifying points increase, finish position tends to decrease. Since lower finish numbers are better (e.g., 1st place), this suggests that horses with more qualifying points generally finish in better positions. 

2) Calculate the correlation coefficient (r) of the scatterplot. (If using the `cor` function in R studio remember to change qualifying points and finish to numeric)
```{r}
cor_coef <- cor(horseRacing$`Qualifying Points`, horseRacing$Finish)
cor_coef
```

`r round(cor_coef, 3)`

a. What does the sign of r tell you?

Since it is negative it means that as qualifying points increase the finished position decreases

b. Does the sign make sense with the research question? Explain.

Yes because it is going in the same direction. 

c. Is the strength of the relationship weak, moderate, or strong?

It is a moderately negative correlation that suggests some general trend but not a perfect one. 

3) Check for outliers by fitting qualifying points and finish onto a model. Using the Residual vs Fitted plot comment on which observations appear to have a high residual and could be an outlier. 
```{r}
model <- lm(Finish ~ `Qualifying Points`, data = horseRacing)
plot(model)
```

Observation 19 and 3 appear to have high residuals and could be an outlier. 

a. How might these outliers affect your interpretation of r?

A horse with very high qualifying points but a poor finish (or vice versa) can pull the line of best fit away from the true trend, weakening or distorting the correlation.

4) Using the Q-Q residual plot, Do the points in the scatterplot suggest a linear trend? If not, what kind of relationship do you see?

The residuals mostly follow the Q-Q line, suggesting that the residuals are approximately normally distributed and the relationship between qualifying points and finish is likely linear.

5) Would this correlation suggest causation? Why or why not?

No, correlation does not imply causation. Although qualifying points are associated with better finishes, we cannot conclude that earning more qualifying points directly causes a horse to finish better, especially with other factors at play.

6) How can confounding variables might influence the result?

Variables like Final Odds, Jockey, and Trainer can all influence a horse’s finish independently of qualifying points. This means the correlation between qualifying points and finish might be partly explained by these other factors and not purely by qualifying points.