---
title: "SLR Module"
output: html_document
date: "2023-07-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://www.racing-reference.info/active-drivers-stats/

```{r}
nascar<- read_csv("bignessalyssa_LATE_1937_225990_nascar_df-1.csv")

nascar_2022 <-nascar%>%
  filter(Year == 2022)%>%
  select(-1, -21)
```

The data set contains each drivers' statistics from the 2022 season. 

```{r}
ggplot(data = nascar_2022, 
       aes(x = AvgStart, y = DriverRating ))+
  geom_point()+
  geom_smooth(se = FALSE)+
  labs(x = "Average Start", y = "Driver Rating", title = "2022 NASCAR Season")
```

Above is scatterplot displaying the relationship between Average Start and Driver Rating. 


1. In the scatterplot above what is the explanatory variable? What is the response? What are the units of both variables respectively?

2. What does each point on the scatter plot represent?

3. Assess and decribe the trend displayed in the scatterplot. (strong, negative, and linear)

4. Given the trend of the smoother, does a linear trend seem reasonable?

5. Write the population equation for the linear model predicting Driver Rating from Average Start.

6.. Fit the least squares regression equation for modeling Driver Rating from Average Start.
```{r}
mod1 <- lm(DriverRating ~ AvgStart, data = nascar_2022)
summary(mod1)
```

7. Interpret the slope coefficient in the context of the data.

8. Provide a literal interpretation of the intercept coefficient. Does it make sense in context of the data? (no, extrapolation)

9. Although it is already known, use the model to predict the Driver Rating of Joey Logano who has an AvgStart of 10.9.

10. Given Joey Logano's actual Driver Rating in 2022 was 89.5, calculate the residual of the model prediction. 

11. What is the size of a typical error for this model? (residual standard error)

12. What percent of variation in Driver Rating is explained by the model using Avg Start? What is the sample correlation?

13. What are the model assumptions for simple linear regression? Examine the plots below to check the assumptions for the linear model. Which assumptions are met? Are there any issues?

```{r}
plot(mod1)
```




```{r}
cor(nascar_2022$AvgStart, nascar_2022$DriverRating)
```

