---
title: "nba_tidying_worksheet"
format: html
---


# Background/context



# 0. Load the following packages
```{r}
#| message: false
library(tidyverse)
library(rvest)
library(janitor)
```


# 1. Load in nba_player_profiles_2025.csv from the data folder.
```{r}

```


# 2. Clean the dataset using the following packages from tidyverse: `dplyr`, `tidyr`, `readr` and `stringr`.
Once you are finished cleaning this dataset, these should be the following variables.

Variable | Type | Description
--------|-------|--------------------------------------------------
`name` | chr | Full name of the NBA player
`team` | chr | Three-letter abbreviation of the team the player is on
`position` | chr | Player's primary on-court position in abbreviated form
`height_inches` | num | Player's height in inches
`wingspan_inches` | num | Player's wingspan in inches
`wingspan_advantage` | num | Difference between wingspan and height in inches (wingspan - height)


a. Convert the `height` variable (which is currently a character string like "6'4") into a new numeric variable called `height_inches` that represents each player's height in total inches. **Hint** using helper variables (such as separate variables for feet and inches) can make this process easier.
```{r}

```


b. Do the same thing as above, but with the `wingspan` variable. Create a new numeric variable called `wingspan_inches` that represents each player’s wingspan in total inches.
```{r}

```


c. Extract the player's position (e.g. "SG" or "C") from the `name` variable and store it in a new variable called `position`, so that the position is no longer part of the name column.
```{r}

```


d. Remove players who are not currently on a team. In the `name` column, these players do not have a three-letter team abbreviation at the end of their name (e.g. "LeBron JamesLAL" vs. "LeBron James"). Then, split the name and team abbreviation into two separate variables: `name` and `team`. **Hint:** Use stringr functions to complete this step.
```{r}

```



# 3. Scrape data from Basketball Reference
- Go to this link, [basketball-reference.com](https://www.basketball-reference.com/leagues/NBA_2025_totals.html)
- In the middle of the page, you'll see boxes with links to different data tables. Choose one of the following tables to scrape: **Totals, Per Game, Per 36 Min, Per 100 Poss** or **Advanced**.

From the janitor library, use the help section to get more details on the functions.

Function | Description
--------|---------------------------------------------------------
`clean_names()` | Standardizes column names to be lowercase, snake_case, and R-friendly.
`row_to_names()` | Promotes a row (usually the first) to become the column names of the table.


a. Scrape the data from the chosen data table
```{r}

```


b. To tidy the chosen dataset, complete the following steps:

**First**, standardize all variable names using the `clean_names()` function.

**Second**, remove duplicate player entries by keeping only the row that represents
a player’s full season total. (For example, Luka Dončić was traded mid-season, so he might appear three times: once for his stats with DAL, once with LAL, and once for his total 2024–25 season stats.) **Hint:** Use `group_by()` and `ungroup()` to help identify duplicates. Also, games played will be the highest for rows that represent a full season total.

**Optional:** Use `select()` to keep only the variables you're interested in.
```{r}

```


c. Do the same steps as in parts **a** and **b**, but this time choose one of the following tables: **Play-by-Play**, **Shooting**, or **Adjusted Shooting**.

**Note:** These tables contain **two** rows of column headers. To fix this issue, use the `row_to_names()` function.
```{r}

```


d. Combine the two cleaned datasets you scraped from Basketball Reference into a single dataset.
```{r}

```



# 4. Combine the cleaned player profiles dataset from part 2, with the combined Basketball Reference dataset you created in part 3.d.
In the 'nba_player_profiles_2025.csv' file, two names were intentionally spelled differently from how Basketball Reference spells the player's names. Keep this issue in the back of your mind. 

a. Use `left_join()` to combine the datasets. Do you notice the misspelled names, what occurred?


b. Use `stringdist_join` from `fuzzyjoin` to combine the datasets. Compare the differences using `stringdist_join` and using `left_join()`.


b. fix names manually.












