---
title: "Exploring NCAA Men's Water Polo: Data Wrangling and Visualization"
---

In this activity, we will be using data from the 2023 NCAA Men's Water Polo season to explore data wrangling and data visualizations in R.

## Part 1 - Introduction

Water polo is a fast paced, competitive game that combines different elements of swimming, soccer, and basketball. The National Collegiate Athletic Association (NCAA) is the governing body that oversees collegiate water polo among many of the top US colleges.

##### Understanding Water Polo

-   The objective in water polo is to score more goals than the opposing team by throwing a ball into the opponent's net.

-   Each team consists of six players plus a goal keeper.

-   Each game consists of four quarters, each eight minutes long, as well as a half-time break.

##### The Field

-   Water polo is played in a pool with a depth of at least six feet. Though exact dimensions vary, the standard size for a NCAA Men's Water Polo game is about 25 by 20 meters. There are two, three meter wide goals are located at each end of the pool.

##### Basic Rules

-   Players are not allowed to touch the bottom of the pool and must tread water or swim throughout the duration of the game.

-   Players also must only use one hand to hold or pass the ball (except for the goalkeeper).

-   There are some minor fours that result in free throws for the opposing team. For major fouls, a player can be temporarily sent out of the game.

Overall, water polo is an exciting sport that requires a lot of physical fitness and athleticism. The NCAA divides competing schools into different divisions based on school size and budgets for athletics.

## Part 2 - Loading the Data

For this activity, you'll be using data from the data sets `ncaa_waterpolo_goals.csv` and `ncaa_waterpolo_assists.csv`. The data sets contain information about the conference leaders in goals and assists for NCAA Men's Water Polo. Each row represents one of the top ten leaders and information about what they lead in, how many goals or assists they have, their conference, and what school they attend.

**1. Load in the data files `ncaa_waterpolo_goals.csv` and `ncaa_waterpolo_assists.csv`.**

```{r}

```

## Part 3 - The Data Sets Individually

**2. The goals data set and the assists data set both have a `Conference` variable designating what conference the leader plays in. For both data sets, make `Conference` a factor with each level being a unique value.**

```{r}

```

**3. Create a boxplot showing the number of goals scored for each conference.**

```{r}

```

**4. Based on the box plots you created above, answer the following questions:**

a.  **Which conference shows the highest median number of goals scored, and which shows the lowest?**


b.  **Are the distributions of goals scored in each conference symmetric, or is there obvious skewness? If so, what direction is the data skewed? What does the skewness imply about the competitive dynamics of the conference.**


## Part 4 - The Data Sets Together

**5. What schools have leaders in both the assists category and the goals category, and how many of each does the school have in total? Create a table that contains only the conference, team, number of goals, and number of assists. (Hint: Use a join() function)**

```{r}

```

**6. What does one row in this data set represent?**



**7. What proportion of teams have leaders in both goals and assists?**


**8. For the teams that have leaders in both categories, what are the average number of goals scored and assists made? Find the teams whose leaders scored higher than average in both categories.**

```{r}

```

**9. Does this information provide evidence to defend your answer to question 4? Explain why or why not.**

