---
title: "Exploring NCAA Men's Water Polo: Data Wrangling and Visualization"
---

In this activity, we will be using data from the 2023 NCAA Men's Water Polo season to explore data wrangling and data visualizations in R.

## Part 1 - Introduction

Water polo is a fast paced, competitive game that combines different elements of swimming, soccer, and basketball. The National Collegiate Athletic Association (NCAA) is the governing body that oversees collegiate water polo among many of the top US colleges.

##### Understanding Water Polo

-   The objective in water polo is to score more goals than the opposing team by throwing a ball into the opponent's net.

-   Each team consists of six players plus a goal keeper.

-   Each game consists of four quarters, each eight minutes long, as well as a half-time break.

##### The Field

-   Water polo is played in a pool with a depth of at least six feet. Though exact dimensions vary, the standard size for a NCAA Men's Water Polo game is about 25 by 20 meters. There are two, three meter wide goals are located at each end of the pool.

##### Basic Rules

-   Players are not allowed to touch the bottom of the pool and must tread water or swim throughout the duration of the game.

-   Players also must only use one hand to hold or pass the ball (except for the goalkeeper).

-   There are some minor fours that result in free throws for the opposing team. For major fouls, a player can be temporarily sent out of the game.

Overall, water polo is an exciting sport that requires a lot of physical fitness and athleticism. The NCAA divides competing schools into different divisions based on school size and budgets for athletics.

## Part 2 - Loading the Data

For this activity, you'll be using data from the data sets `ncaa_waterpolo_goals.csv` and `ncaa_waterpolo_assists.csv`. The data sets contain information about the conference leaders in goals and assists for NCAA Men's Water Polo. Each row represents one of the top ten leaders and information about what they lead in, how many goals or assists they have, their conference, and what school they attend.

**1. Load in the data files `ncaa_waterpolo_goals.csv` and `ncaa_waterpolo_assists.csv`.**

```{r}
library(tidyverse, quietly = TRUE)

ncaa_waterpolo_goals <- read_csv("ncaa_waterpolo_goals.csv",show_col_types = FALSE)
ncaa_waterpolo_assists <- read_csv("ncaa_waterpolo_assists.csv",show_col_types = FALSE)
```

## Part 3 - The Data Sets Individually

**1. The goals data set and the assists data set both have a `Conference` variable designating what conference the leader plays in. For both data sets, make `Conference` a factor with each level being a unique value.**

```{r}
ncaa_waterpolo_goals$Conference <- factor(ncaa_waterpolo_goals$Conference)
ncaa_waterpolo_assists$Conference <- factor(ncaa_waterpolo_assists$Conference)
```

**2. Create a boxplot showing the number of goals scored for each conference.**

```{r}
ncaa_waterpolo_goals %>% group_by(Conference) %>%
  ggplot(.,
         mapping = aes(x = Goals,
                       fill = Conference
                       )) + geom_boxplot()
```

**3. Based on the box plots you created above, answer the following questions:**

a.  **Which conference shows the highest median number of goals scored, and which shows the lowest?**

-   The Western Water Polo Association has the highest median number of goals scored. The Mid Atlantic Water Polo Conference has the least amount of goals scored.

b.  **Are the distributions of goals scored in each conference symmetric, or is there obvious skewness? If so, what direction is the data skewed? What does the skewness imply about the competitive dynamics of the conference.**

* The two conferences with the most obvious skewness is the Western Water Polo Association and the Northeast Water Polo Conference. All of the other conferences are relatively equal in distribution. Both the Western Water Polo Association and the Northeast Water Polo Conference are skewed in the positive direction this implies that there could be one or a few players from that conference that are scoring a considerably higher number of goals than other players in that conference.

**4. Out of the top ten leaders, find the total number of goals and assists scored for each conference. Make one table which contains the conference, number of goals, and number of assists. Which has the most and which has the least in each category? (Hint: Use a join() function)**

```{r}
ncaa_waterpolo_goals %>% group_by(Conference) %>%
  summarize(
    sum = sum(Goals)
    )%>% 
  arrange(desc(sum))
```

-   The Western Water Polo Association has the most amount of goals scored, 814. The Mid Atlantic Water Polo Association has the least amount of goals scored, 442.

## Part 4 - The Data Sets Together

**1. What schools have leaders in both the assists category and the goals category, and how many of each does the school have in total? Create a table that contains only the conference, team, number of goals, and number of assists. (Hint: Use a join() function)**

```{r}
# total goals for each team 
team_summary_goals <-
  ncaa_waterpolo_goals %>% group_by(Team, Conference) %>%
  summarize(
    n_goals = sum(Goals)
  )

# assists by team and conference 
team_summary_assists <-
  ncaa_waterpolo_assists %>% group_by(Team, Conference) %>%
  summarize(
    n_assists = sum(Assists)
  )

# joining together by team so get the matching teams 
full_summary_table <- inner_join(team_summary_assists, team_summary_goals,by = "Team") %>%
  arrange(desc(n_goals)) %>%
  select(-Conference.y)

# editing format 
full_summary_table <- full_summary_table %>%
  rename(
    Conference = Conference.x,
    `Total Assists` = n_assists,
    `Total Goals` = n_goals
  )

full_summary_table <- full_summary_table[, c("Conference", "Team", "Total Assists", "Total Goals")]

full_summary_table

```

**2. What does one row in this data set represent?**

* Each row represents a team and has the total number of goals and assists scored by leaders in that conference 

**3. What proportion of teams have leaders in both goals and assists?**

* $27/70 = 0.386 = 38.6$% of teams have leaders in both goals and assists

**4.**

```{r}
cor(full_summary_table$`Total Assists`, full_summary_table$`Total Goals`)
```

```{r}
ggplot(data = full_summary_table,
       mapping = aes(x = `Total Assists`,
                     y = `Total Goals`,
                     color = Conference)) +
  geom_point()
```


what are the average number of goals scored and assists made? find teams whose leaders scored higher than average in both categories? 

```{r}
# finding the average of both columns 
avg_goals <- mean(full_summary_table$`Total Goals`)
avg_assists <- mean(full_summary_table$`Total Assists`)

# filtering
full_summary_table %>%
  filter(`Total Assists` > avg_assists &
           `Total Goals` > avg_goals)
```

Does this information provide evidence to defend your answer to question 3. Explain why or why not. 

This does defend the answer to question three. The Western Water Polo Association and the Northeast Water Polo Conference each have two teams that have above average total goals and total assists. This helps to explain why the data is considerably skewed in the positive direction, especially if the median of those conferences is lower than the average. 

