---
title: "soccer_Solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
---
Loading necessary packages:
```{r}
library(dplyr)
```


Reading in the Women's data set:
```{r}
url <- "https://www.ussoccer.com/uswnt-stats/2024"
tab <- read_html(url) |>
  html_nodes("table")
womenSoccer <- tab[[1]] |> html_table()
```

Reading in the Men's data set
```{r}
url1 <- "https://www.ussoccer.com/usmnt-stats/2024"
tab1 <- read_html(url1) |>
  html_nodes("table")
menSoccer <- tab1[[1]] |> html_table()
```

Notice in both data sets the variable names are being treated as values instead of the column names. 

1) Fix the `womenSoccer` and the `menSoccer` data set so that the first row is the variable names.

```{r}
colnames(womenSoccer) <- as.character(womenSoccer[1, ])
womenSoccer1 <- womenSoccer[-1, ]

colnames(menSoccer) <- as.character(menSoccer[1, ])
menSoccer1 <- menSoccer[-1, ]
```

2) Combine the men and women's data set into one data set. 

```{r}
combined_soccer <- bind_rows(womenSoccer1, menSoccer1)
```

a - What looks wrong in the new data set?

There are a bunch of NA values 

b - Why do you think the data set doesn't look right?

The variable names are labeled differently in each data set and the men's soccer data set has added variables the women's data set does not have.

3) Change the variable names of the women's data set to match the men's and get rid of any variables in the men's that don't appear in the women's data set.

```{r}
womenSoccerMatched <- womenSoccer1 |> rename(Player = Name, Pos = Pos., YC = Y, RC = R, `Career Goals` = Goals, MIN = Min)

menSoccerMatched <- menSoccer1 |> select(1:7, 9:10, 12:13)
```

Before we recombine the data sets, it might be helpful to know if the player is male or female. 

4) Add a new variable to both individual data sets that says "M" or "F' for gender of player. 

```{r}
womenSoccerMatched$Sex <- "F"

menSoccerMatched$Sex <- "M"
View(menSoccerMatched)
```

5) Slice the Men's data set to exclude `OWN GOAL`, `TOTALS`, and `OPPONENT TOTALS` and slice the Women's data set to exclude `Own Goal`, `TOTAL`, and `Opponent`.

```{r}
cleanedwomenSoccer <- womenSoccerMatched |> slice(1:39)
cleanedmenSoccer <- menSoccerMatched |> slice(1:48)
```

6) Combine the the updated data sets.

```{r}
combined_SoccerMatched <- bind_rows(cleanedwomenSoccer, cleanedmenSoccer)
```



