---
title: "2025_Track_NLS_SOLUTIONS"
format: html
---

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)

track <- read.csv("WomensTrack2025.csv")
```

1. Make a scatterplot using distance (Event) to predict speed.

```{r}
ggplot(track, aes(x = Event, y = Speed)) +
  geom_point()
```

The scatterplot above shows that at first, there is very quick decay in speed as distance increases, but then that decay slows and soon levels out. There are a few different ways we could account for this trend in a model. One would be to apply a transformation to the data and continue to use a linear model, another way (the one this worksheet will be exploring) is to switch to a non-linear model. 

There are many different kinds of non-linear models, so the first step in using one is to decide the most useful way to build it. In this case, there are 2 non-linear models that show trends most similar to the data here. One is a logarithmic decay model and the other is a power decay mode. 

We will start by applying an logarithmic decay model to the data. A logarithmic decay model has the following form:

$$ y = a - b\ln(x) $$
Where a is the value of y when x = 1—where the decay curve starts—and b is how fast the value decreases (and for decay must be positive, as the negative is already included in the equation). As b is larger, the drop will be steeper at first. The ln(x) term allows the decay to slow over time.

```{r}
logarithmic <- function(x, a, b) { a-b*log(x)}
```

The above code defines the logarithmic decay function, allowing us to use it within the `nls` function, as in the code below. For `nls` models, you must provide starting values for each of the parameters; these values do not have to be exact, they just have to be reasonable guesses. Here, we will guess 10 for $a$ as we know that it must be higher than 9 (as that is the value when x = 100), but we cannot see how much higher. For $b$ we will guess 1 as the curve is not too extreme.

```{r}
nls_mod <- nls(Speed ~ logarithmic(Event, a, b),
                data = track,
                start = list(a = 10, b = 1))
summary(nls_mod)
```

2. Write out the fitted model.

$$\widehat{\text{Speed}} = 13.12 - 0.92\ln(\text{Event}) $$

3. Report the AIC and BIC for the model.

```{r}
AIC(nls_mod)
BIC(nls_mod)
```

The AIC for the model is 252.70 and the BIC is 264.29.

4. Plug in the correct values for a and b and run the code to make a scatterplot that shows the nls line with the data.

```{r}
a = 
b = 
ggplot(track, aes(x = Event)) +
  geom_point(aes(y = Speed)) +
  geom_function(fun = function(x) { a-b*log(x) }, color="hotpink")
```


Next, we will repeat the process to apply a power decay model to the data. A power curve regression equation has the following form:

$$ y = ax^b $$
```{r}
power_curve <- function(x, a, b) { a*(x^b) }

nls_mod2 <- nls(Speed ~ power_curve(Event, a, b),
                data = track,
                start = list(a = 15, b = -.1))

summary(nls_mod2)
AIC(nls_mod2)
```

Plug in the correct function and use the scatterplot to test different $a$ and $b$ values to find good starting guesses.
```{r}
a = 
b = 
ggplot(track, aes(x = Event)) +
  geom_point(aes(y = Speed)) +
  geom_function(fun = function(x) {  }, color="hotpink")
```

Fit the model

Write out the fitted model.

Report the AIC and BIC for the model.



The easi

--------
Report AIC, BIC, R-squared (for each model).

If you completed the transformation module with this data set, compare the model you created there with the models from this module.

Which model do you prefer? Why?

Answers may vary. A simple answer would be to chose the power curve model because it has the most success (based on the R-squared), but equally as valid would be an argument in favor of the log transformation model due to its simplicity. The R-squared values do not show a significant enough preference for any of the models to override a preference because of ease of use and explanation.