---
title: "Erging Data Wrangling Worksheet"
format: html
---

1.  Copy this code to load the libraries needed for this worksheet.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
```

2.  Copy this code to read in the data set.

```{r message=FALSE, warning=FALSE}
erg_df <- read_csv("erg.csv")
erg_df <- erg_df |> mutate(Verified = as.factor(Verified), Gender = as.factor(Gender), Year = as.factor(Year), Distance = as.factor(Distance), Age_Group = as.factor(Age_Group))
```

3.  Make a side-by-side boxplot of the `Speed` for `Verified` and `Non-Verified` pieces. Give the plot a label.

```{r}

```

4.  Describe the distribution of the plot.



5.  Find the summary statistics for `Speed` and calculate its IQR.

```{r}

```

6.  Make a table of the top 10 speeds for rowers. Including only their `Name`, `Speed` and if their piece was `Verified`.

```{r}

```

7.  Using the table, identify the 4 outliers seen in the boxplot.

```{r}

```

8.  Perform a t.test to evaluate if there is a significant difference in the mean `Speed` for `Verified` and `Non-Verified` pieces.

```{r}

```

9.  Make a new dataset which does not have the 4 outliers in it.

```{r}

```

10.  With the new dataset make another side-by-side boxplot of the `Speed` for `Verified` and `Non-Verified` pieces and give the plot a label.

```{r}

```

11.  Comment on the distribution of this new boxplot.

```{r}

```

12.  Perform a t.test with the new dataset and re-evalute if there is a significant difference in the mean `Speed` for `Verified` and `Non-Verified` pieces.

```{r}

```

13.  What does the number of outliers for `Non-Verified` pieces say about the reliability of `Non-Verified` pieces in the dataset?

14. Is it ethical to remove the outliers from the dataset as done in task 9?
