---
title: "Erging Data Wrangling Worksheet"
format: html
---
### Description
This activity looks at individual player statistics for handball's [Bundesliga 2022-2023 season](https://www.kaggle.com/datasets/nemanjagojkovic/handball-bundesliga-stats-202223){target="_blank"}. 

Handball is a popular sport in many European countries such as Germany
where it is said to have originated. The Bundesliga, for example, is a
German men's professional handball league. Handball is typically played
indoors on a rectangular court (20m x 40m). There are two goals (3m x
2m) on opposite sides of the court, the goal for each team is to score a
goal by getting the ball in the other team's goal. The challenge to this
is that there is a semicircle with a radius of 6m surrounding the goal
which players other than the goal keeper are not allowed in, making it
challenging to score. Players run back and forth down the court passing
the ball to each other and trying to score. 

There are seven positions total in handball: the goalkeeper who defends the team's goal, left and
rights backs are positioned on the left and right side of their half of
the court to provide further defense, the center can move up and down
the court and is usually the one trying to score, left and right wings
can also move up and down the court, serving as offense when the team
pushes for attack and defense when the opposing team tries to score,
lastly, the pivot is considered strictly an offensive player as they are
usually position in the opposing side of the court, they often work
closely with the center. The Bundesliga regular season length is 34
games, with playoffs the maximum number of games a team could play is
41. Players do not play every game in the season and subbing is common,
generally speaking playing time goes to players with experience. This can create a bit of a disparity in which players with less
playing time will not necessarily have statistics that accurately
display their skills due to the smaller sample size.

Handball is considered a contact sport which means aggressive strategies
are often used in games. Aggressiveness can be measured in the penalty
statistics as players who tend to get more penalties are usually more
considered more aggressive players overall. Their success can be
measured with the [handball performance index (HPI)](https://www.liquimoly-hbl.de/en/s/handball-performance-index/2021-22/handball-performance-index--data-based--transparent--fair/){target="_blank"}, a calculated
statistic which essentially ranks how good a player is. This data set could provide insight on if
players or teams that are more aggressive as measured by penalties are more successful than those
who are more passive. 

### Tasks

1.  Copy this code to load the libraries needed for this worksheet.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
```

2.  Copy this code to read in the data set.

```{r message=FALSE, warning=FALSE}
erg_df <- read_csv("erg.csv")
erg_df <- erg_df |> mutate(Verified = as.factor(Verified), Gender = as.factor(Gender), Year = as.factor(Year), Distance = as.factor(Distance), Age_Group = as.factor(Age_Group))
```

3.  Make a side-by-side boxplot of the `Speed` for `Verified` and `Non-Verified` pieces. Give the plot a label.

```{r}

```

4.  Describe the distribution of the plot.



5.  Find the summary statistics for `Speed` and calculate its IQR.

```{r}

```

6.  Make a table of the top 10 speeds for rowers. Including only their `Name`, `Speed` and if their piece was `Verified`.

```{r}

```

7.  Using the table, identify the 4 outliers seen in the boxplot.

```{r}

```

8.  Perform a t.test to evaluate if there is a significant difference in the mean `Speed` for `Verified` and `Non-Verified` pieces.

```{r}

```

9.  Make a new dataset which does not have the 4 outliers in it.

```{r}

```

10.  With the new dataset make another side-by-side boxplot of the `Speed` for `Verified` and `Non-Verified` pieces and give the plot a label.

```{r}

```

11.  Comment on the distribution of this new boxplot.

```{r}

```

12.  Perform a t.test with the new dataset and re-evalute if there is a significant difference in the mean `Speed` for `Verified` and `Non-Verified` pieces.

```{r}

```

13.  What does the number of outliers for `Non-Verified` pieces say about the reliability of `Non-Verified` pieces in the dataset?

14. Is it ethical to remove the outliers from the dataset as done in task 9?
