---
title: "handball_questions"
format: html
---

# Libraries

```{r}
library(tidyverse)
library(readr)
library(modelr)
library(broom)
library(stringr)
```

# Data

```{r}
handball_clean <- read_csv("handball_bundesliga_23.csv")
```

# Question 1

Calculate and interpret a 98% confidence interval for the mean HPI of a player with 30 total_penalties. Check your work with R.

```{r}
mod <- lm(HPI ~ total_penalties, data=handball_clean)

newx <- data.frame(total_penalties=30)

predict(mod, newx,interval="confidence", level = 0.98)
```

# Question 2

Calculate and interpret a 98% prediction interval for the HPI of a player with 30 total_penalties. Check your work with R.

```{r}
mod <- lm(HPI ~ total_penalties, data=handball_clean)

newx <- data.frame(total_penalties=30)

predict(mod, newx,interval="prediction", level = 0.98)
```

# Question 3

The below scatterplots show the mean HPI of each club against clubs' mean penalties and offense. In comparing the two plots, what do you expect from the model: HPI = total_offense + total_penalties?

```{r}
#a&b)
mod_plot <- handball_clean |> group_by(CLUB) |> summarise(club_offense = mean(total_offense), club_pen = mean(total_penalties), club_hpi = mean(HPI))
#b)
ggplot(data = mod_plot, aes(x = club_offense, y = club_hpi)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_minimal() + labs(title = "Scatterplot of club_hpi against club_offense with regression line")
#a)
ggplot(data = mod_plot, aes(x = club_pen, y = club_hpi)) +
  geom_point() +
  geom_smooth(method = "lm") + theme_minimal() + labs(title = "Scatterplot of club_hpi against club_penalties with regression line")
```
         

# Question 4

Below is a table of the coefficients for the model HPI = ÃŸ~total_offense~ + ÃŸ~total_penalties~, interpret ÃŸ~total_offense~ and ÃŸ~total_penalties~ in the context of HPI.

```{r}
hpi_mod <- lm(HPI~total_offense+total_penalties, data=handball_clean)
summary(hpi_mod)
```

# Question 5

Perform an ANOVA test to assess the overall fit of HPI = total_offense + total_penalties. Fill in the ANOVA table below and interpret the results. Check your work with R.


H~o~: ð›½~total_offense~ = ð›½~total_penalties~ = 0

H~a~: ð›½~total_offense~ OR ð›½~total_penalties~ â‰  0

```{r}
anovamod <- lm(HPI~total_offense+total_penalties, data=handball_clean)
summary(anovamod)
anova(anovamod)
```

# Question 6

The below scatterplot shows total_penalties against total_offense with a regression line. Based on this plot what do you expect the correlation between total_penalties and total_offense to be?

```{r}
ggplot(data = handball_clean, aes(x = total_offense, y = total_penalties)) + geom_point(alpha = 0.5, aes(colour = POSITION)) + theme_minimal() + geom_smooth(method = "lm") + labs(title = "Scatterplot of total_penalties against total_offense with a regression line")
```

# Question 7

Find the correlation of total_penalties and total_offense. 

```{r}
cor(handball_clean$total_penalties, handball_clean$total_offense)
```


# Question 8

Test the significance of the correlation between the total_offense and the total_penalties of a player. Provide an interpretation of the results. Check your work with R.

```{r}
cor.test(handball_clean$total_offense, handball_clean$total_penalties)
```

# Question 9

Could it be concluded that having more penalties increases the skill and success of a player in the form of HPI?
