---
title: "NonlinearLeastSquares_Solutions"
format: 
  html:
    embed-resources: true
format: html
---

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(broom)
library(readr)
powerlift <- read_csv(here::here("frhinehart/Powerlifting/module/powerlift_sample_data_updated.csv")) |>
  filter(Event == "SBD") |>
  filter(Equipment %in% c("Raw", "Single-ply")) |>
  filter(Place != "DQ") |>
  filter(Place != "NS") |>
  drop_na(Place) |>
  filter(Place != "G") |>
  filter(Place != "DD") |>
  drop_na(Age) |>
  drop_na(AgeClass) |>
  drop_na(Best3DeadliftKg)
```


1) Graph a scatter plot showing the relationship comparing Age and Best3DeadliftKg. Add a smoother. What do you notice about the curve?

```{r}
ggplot(data = powerlift, aes(x = Age, y = Best3DeadliftKg)) +
  geom_point() +
  geom_smooth(color = "blue", se = FALSE)
```

It is nonlinear.

2) 

a) Fit a nonlinear least squares model using the shifted gamma function provided below:
```{r}
# Defining a shifted gamma function
shifted_gamma <- function(x, a, b, x0) {
  x_start <- pmax(x - x0, 0)
  a * x_start * exp(-b * x_start)}
```

```{r}
nls_fit <- nls(Best3DeadliftKg ~ shifted_gamma(Age, a, b, x0),
 data = powerlift,
 start = list(a = 300, b = 0.05, x0 = 10))
nls_fit
```

This is the model you should get:
$$y = (19.67175)(age - 4.94486)e^{((-.03451)(age - 4.94486))}$$
The model starts using age to model how strength progresses, with the deadlift increasing as age increases. The  rate of increase slows due to the exponential decay factor. As people age, the exponential decay term reduces the growth of the deadlift. The older someone gets, the less their deadlift will increase. This can be modeled when graphed...

b) Add predicted values to your dataset
```{r}
powerlift$pred <- predict(nls_fit)
```

c) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a prediction line. 
```{r}
ggplot(powerlift, aes(x = Age)) +
  geom_point(aes(y = Best3DeadliftKg), 
             alpha = 0.3, color = "gray40") +
  geom_smooth(aes(y = Best3DeadliftKg), color = "blue", fill = "blue") +
  #geom_function(fun = function(x) {(19.67175)*(x - 4.94486)*exp((-.03451)*(x - 4.94486))} ) +
  geom_line(aes(y = pred), color = "red", linewidth = 1.2) +
 labs(title = "Fitted shifted Gamma Curve to Deadlift vs. Age",
   x = "Age",
   y = "Best 3 Deadlift (kg)") +
 theme_minimal()
```

d) For this particular equation the maximum age for peak deadlift is 33.92. If you have taken calculus, derive $$(x-x0)e^{(-b(x-x0))}$$ to model this using the values given in part a

$$z = x-c \text{ which can be rewritten as } ze^{-bz}$$ 
$$z(e^{-bz}) + (e^{-bz})$$
$$e^{-bz} (-zb + 1) = 0$$
What is in the () is what needs to equal 0 so we can drop $$e^{-bz}$$
$$-zb + 1 = 0 \text{ which equals } z = 1/b$$
Since z = x - x0 the updated formula is 
$$x - x0 = 1/b \text{ where } x = x0 + 1/b$$
Plug in values and get:
$$x = 4.94486 +(1/-.3451) = 33.92 $$
e) Use boot strap to calculate a 95% confidence interval for age where the peak occurs (x = 33.92 on the original dataset). Interpret your results and does it contain the value found in part d. 

```{r}
values <- coef(nls_fit)
x_max <- values["x0"] + 1/values["b"]
```

```{r}
# Step 0. Set up objects 
B <- 1000 # number of bootstrap samples

agePeak_boots <- rep(NA, B) # a blank vector to store the bootstrap statistics

# set up loop
for(i in 1:B){
  # 1. Resample Data
  resampled_data <- powerlift |>
    slice_sample(prop = 1, replace = TRUE)
          
  #2. Calculate statistic
  # Step 1: refit model
  nls_fit_boot <- nls(Best3DeadliftKg ~ shifted_gamma(Age, a, b, x0),
  data = resampled_data,
  start = list(a = 300, b = 0.05, x0 = 10))
  values <- coef(nls_fit_boot)
  x_max_boot <- values["x0"] + 1/values["b"]
    
  #3. Store results
  agePeak_boots[i] <- x_max_boot
} # Back to 1. Resample Data, repeat B times
```

```{r}
#4. Calculate CI
qplot(agePeak_boots)

conf.level <- 0.95
alpha <- 1 - conf.level

quantile(agePeak_boots, probs = c(alpha/2, 1-alpha/2)) |>
  round(2)
```

With 95% confidence, the peak age for max deadlift is between 33.46 and 34.42 years old. Which contains the max value found in part d. 

3) 
a) Fit a nonlinear least squares model using the reparameterized double exponential with x_shift provided below
```{r}
double_exp_shift <- function(x, A, k1, k2, x_shift) {
  x_start <- pmax(x - x_shift, 0)
  A * exp(-k1 * x_start) * (1 - exp(-k2 * x_start))}
```

```{r}
nls_fit2 <- nls(Best3DeadliftKg ~ double_exp_shift(Age, A, k1, k2, x_start),
  data = powerlift,
  start = list(A = 300, k1 = 0.02, k2 = 0.1, x_start = 27))
nls_fit2
```

This is the model you should get:
$$y = (285.116)e^{((-.0111)(xstart))} * 1- e^{-.1150 * xstart)}$$

The second term increases significantly, reflecting a growth in deadlift strength over time. This suggests that the individual’s best 3 deadlift increases as they age, reaching their peak in their 20s or 30s. The first starts to lower the deadlift value, reflecting the gradual reduction in strength due to aging where the second term reaches it's maximum values and stabilizes. This can be modeled when graphed...

b) Add predicted values to your dataset
```{r}
powerlift$pred <- predict(nls_fit2)
```

c) Plot a scatterplot of Actual vs Predicted using a smoother for Best3DeadliftKg and with a line of best fit.
```{r}
ggplot(powerlift, aes(x = Age)) +
  geom_point(aes(y = Best3DeadliftKg), alpha = 0.3, color = "gray40") +
  geom_smooth(aes(y = Best3DeadliftKg), color = "blue", fill = "blue") +
  geom_line(aes(y = pred), color = "darkgreen", linewidth = 1.2) +
  labs(title = "Double Exponential with Explicit shift Parameter",
    subtitle = "Estimated shift location is directly modeled",
    x = "Age",
    y = "Best 3 Deadlift (kg)") +
  theme_minimal()
```


d) For this particular equation the maximum age for peak deadlift is 26.05. If you have taken calculus, derive $$Ae^{(-k1(x-xstart)} * (1-e^{k2(x-xstart)})$$ to model this using the values given in part a

we will use z - x-xstart to simplify solving

$$e^{-k1z} = e^{-(k1 + k2)z}$$
$$-k1e^{-k1z} - (k1 + k2) e^{-k2z}= 0$$
$$e^{-k2z} = k1/k1+k2$$
$$-k2z = ln(k1/k1+k2)$$
$$z = -1/k2ln(k1/k1+k2) $$
replace z with x - xstart...
$$x = (-1/k2) ln(k1/(k1+k2) + xstart$$
PLug in the values: 
$$x = (-1/0.11501654) ln(.01109791/(.01109791+0.11501654) + 8.12145801$$

e) Use boot strap to calculate a 95% confidence interval for age where the peak occurs (x = 26.05 on the original dataset). Interpret your results and does it contain the value found in part d.

```{r}
values2 <- coef(nls_fit2)
x_max2 <- (-1 / values2["k2"]) * log(values2["k1"] / (values2["k1"] + values2["k2"]))
```

```{r}
# Step 0. Set up objects 
B <- 1000 # number of bootstrap samples

agePeak_boots2 <- rep(NA, B) # a blank vector to store the bootstrap statistics

# set up loop
for(i in 1:B){
  # 1. Resample Data
  resampled_data <- powerlift |>
    slice_sample(prop = 1, replace = TRUE)

          
  #2. Calculate statistic
  # Step 1: refit model
  nls_fit_boot2 <- nls(Best3DeadliftKg ~ double_exp_shift(Age, A, k1, k2, x_shift),
  data = resampled_data,
  start = list(A = 300, k1 = 0.02, k2 = 0.1, x_shift = 27))
  values2 <- coef(nls_fit_boot2)
  x_max_boot2 <-  (-1 / values2["k2"]) * log(values2["k1"] / (values2["k1"] + values2["k2"]))
    
  #3. Store results
  agePeak_boots2[i] <- x_max_boot2
} # Back to 1. Resample Data, repeat B times
```

```{r}
#4. Calculate CI
qplot(agePeak_boots2)

conf.level <- 0.95
alpha <- 1 - conf.level

quantile(agePeak_boots2, probs = c(alpha/2, 1-alpha/2)) |>
  round(2)
```

With 95% confidence, the peak age for max deadlift is between 20.14 and 22.49. Which contains the max value found in part d. 

4) Compare both models you created. Which one is more useful in predicting dead lift using age? Report the residual sum-of-squares for both models

Gamma residual sum-of-squares: 9838215

Reparameterized residual sum-of-squares: 9448921

reparameterized double exponential with x_shift is more useful since it has a lower residual sum-of-squares

5) If you completed the other powerlift module, use the indicator variable code provided below to recreate a 5th order polynomial regression model and find the AIC
```{r}
powerlift_new$SexIND = ifelse(powerlift_new$Sex == "M", 1, 0)
``` 

```{r}
dl_finish <- lm(data = powerlift_new, Best3DeadliftKg ~ poly(Age, 5)*SexIND)
summary(dl_finish)

AIC(dl_finish)
```

a) Find the AIC for the model you answered in question 4
```{r}
AIC(nls_fit2)
```

b) Compare the polynomial model' AIC to your answer in question 4's AIC. Which one is better in predicting dead lift using age? Report each of the model's AIC. 

The polynomial model is better as it has a lower AIC

dl_finish AIC: 38328
nls_fit2 AIC: 40775.1



