---
title: "Horse Racing: Pearson Correlation Hypothesis Test Worksheet Solutions"
Authors: Faith Rhinehart
execute: 
  echo: true
  warning: false
  error: true
format: 
  html:
    embed-resources: true
  docx:  default
  pdf:  default
---

Read in the dataset and necessary packages: 
```{r}
library(readr)
library(ggplot2)
horseRacing <- read_csv("frhinehart/Horse Racing/Simple&MulipleLinearRegressionModule/Module/horseRacing.csv")
```

Research Question of Interest: We are testing whether Morning Line Odds is associated with Finish Position.

1. Visualization

	a) Create a scatterplot comparing Morning Line Odds and Finish. Add a smoother.

```{r}
ggplot(data = horseRacing, aes(x = `Morning Line Odds`, y = Finish)) +
  geom_point() +
  geom_smooth(method = lm)
```
	
	b) What do you notice about the relationship of the plot?

As morning line odds increase the finish position increases. Suggesting that the lower odds place better than the higher odds.

Let’s test this with a simple linear regression (or Pearson correlation).

2. Model Building

  a) Write the population model for the question of interest.

$$Finish = \beta_0 + \beta_1MorningLineOdds + \epsilon $$

  b) Fit the model using the model summary output. Report the fitted model.
  
```{r}
model = lm(Finish ~ `Morning Line Odds`, data = horseRacing)
summary(model)
```

$$Finish = 1.9966 + 0.3638MorningLineOdds$$

  c) Interpret the slope of the model in context.
  
For every 1-unit increase in Morning Line Odds, the average finish position increases by about 0.36 places, meaning the horse tends to place slightly worse.

3. Defining Hypothesis

  a) Define the null and alternative hypothesis.
$$H_o: \beta_1 = 0 \space (\text{no relationship})$$
$$H_a: \beta_1 \neq 0 \space (\text{there is a relationship})$$

  b) Based on the model summary output, do you expect to reject to fail to reject the null.
  
I expect to reject the null hypothesis.

4. Run the two-tailed hypothesis test. Hint: your answers should match the model summary output

Show your work for the following: 

  a) Report the standard error
  
0.1060

  b) Calculate the degrees of freedom

$$df = n - 2 = 19-2 = 17$$
  
  c) Calculate the t-statistic. Show your work of how you calculated it. 
  
$$t = \frac{\hat{\beta_1}}{SE} = \frac{0.3638} {0.1060} = 3.432$$
  
  d) Report the p-value using a two-tailed t-distribution online software or by hand.
  
0.0032
  
5. Conclusion

  a) Conclude your results
  
There is strong evidence (df = 17, t-statistic = 3.432, p-value = .0032) to conclude that there is a relationship between morning line odds and finished place. 
  
6. Reflection/Opinion

  a) If you were betting in the Kentucky Derby would you trust Morning Line Odds, why?