---
title: "racewalk_dataprep"
format: html
---
# Load in libraries

```{r}
library(tidyverse)
library(readr)
```

# Read in data

Data was copied and pasted from [this link](https://www.watchathletics.com/page/5352/results-world-athletics-race-walking-team-championships-2024#google_vignette){target="_blank"} and then modified in an excel document to include  distance and sex.

That excel document is read in below.

```{r}
worldchamps <- read_csv("~/Desktop/git/stat_289_score/awsmit22/racewalking/module/worldchamps.csv")
```

# Data Prep

Create a data set with only the 10k racers and separate the `TIME` variable so that it is `TIME`  and `RECORD` seperately. minutes and seconds, using these two new variables, the `TIME` variable can be recalculated as total minutes. 

```{r}
ten <- worldchamps |> filter(DISTANCE == 10000) |> separate(TIME, into = c("TIME", "RECORD"), sep = " ") |> separate(TIME, into = c("MIN", "SEC"), sep = ":") |> mutate(MIN = as.double(MIN), SEC = as.double(SEC)) |> mutate(SEC = SEC/60) |> mutate(TIME = SEC + MIN) |> mutate(RECORD = replace_na(RECORD, replace = "NONE")) 
```

Create a data set with only the 20k racers and separate the `TIME` variable so that it is minutes and seconds, using these two new variables, the `TIME` variable can be recalculated as total minutes.

```{r}
twenty <- worldchamps |> filter(DISTANCE == 20000) |> separate(TIME, into = c("TIME", "RECORD"), sep = " ") |> separate(TIME, into = c("HRS", "MIN", "SEC"), sep = ":") |> mutate(HRS = as.double(HRS), MIN = as.double(MIN), SEC = as.double(SEC)) |> mutate(HRS = HRS*60, SEC = SEC/60) |> mutate(TIME = HRS + SEC + MIN) |> mutate(RECORD = replace_na(RECORD, replace = "NONE")) 
```

```{r}
complete <- full_join(ten, twenty)
```

```{r}
racewalking <- complete |> select(POS, COUNTRY, ATHLETE, RECORD, DISTANCE, GENDER, TIME)
```

```{r}
racewalking_df <- racewalking |> mutate(SPEED= DISTANCE/TIME) 
```

```{r}
racewalking_df <- racewalking_df |> mutate(GENDER = if_else(GENDER == 1, true = "Woman", false = "Man"))
```

```{r}
racewalking_df <- racewalking_df |> mutate(DISTANCE = as.factor(DISTANCE))
```