---
title: "fencing_213_question_prep"
format: html
---
# Libraries

```{r}
library(tidyverse)
library(readr)
```

# Data

```{r}
fencing_df <- read_csv("fencing.csv")
fencing_df
```

# Question 1

Create a subset for Ivies and another for Non-Ivies. Write down the number of Ivy students and the number of Non-Ivy students.

```{r}
ivies <- subset(fencing_df, Ivy == "Ivy")
non_ivies <- subset(fencing_df, Ivy == "Non-Ivy")
```

**Ivy = 54**
**Non-Ivy = 90**

# Question 2

Find the mean Victories for Ivies and Non-Ivies.

```{r}
mean(ivies$Victories)
mean(non_ivies$Victories)
```

**Ivies = 12.54**
**Non-Ivies = 10.88**

# Question 3

Find the standard deviation of Victories for Ivies and Non-Ivies.

```{r}
sd(ivies$Victories)
sd(non_ivies$Victories)
```

**Ivies = 3.74**
**Non-Ivies = 3.71**

# Question 4

Create a side-by-side boxplot of Victories for Ivies and Non-Ivies, compare the distributions.

```{r}
boxplot(Victories ~ Ivy, data = fencing_df)
```

# Question 5

Test for a discernible difference in the mean number of victories for ivy league fencers and non-ivy league fencers. 

```{r}
t.test(Victories ~ Ivy, data = fencing_df)
```

**H0: mean victories for ivy = mean victories for non-ivy**
**Ha: mean victories for ivy != mean victories for non-ivy**
**t_value: 2.59**
**p_value: 0.011**
**Conclusion: Reject H0; 0.011 < 0.05**
**Interpretation: We have significant evidence that the mean number of victories for ivy league fencers is higher than the mean number of victories for non-ivy league fencers.**






