---
title: "Canoeing_Outliers"
format: html
---

```{r}
library(tidyverse)
library(dplyr)
library(broom)

RR_dnf_data <- read_csv("phoebechar/raisin_river/module/raisin_river_DNF.csv")
```

## Graph a scatterplot that shows flow rate as a predictor for rate of DNF

```{r}
ggplot(RR_dnf_data, aes(x = flow, y = ratio_DNF)) +
  geom_point() 
```

### What do you notice?

There is one point that is inconsistent with the others.

## Fit a model with flow rate as a predictor for number of DNFs

```{r}
dnf_mod <- lm(ratio_DNF ~ flow, RR_dnf_data)
summary(dnf_mod)
```

## Comment on the usefulness of the model

It is not useful

## Use the augment function to add the residuals into the dataset

```{r}
dnf_resid = left_join(select(rr_DNF, year, ratio_DNF, flow), augment(dnf_mod))
```

## Do any observations have unusual leverage? standardized residual? Influence?

```{r}
dnf_resid |> filter(.hat > 2*(2/9)) ## 2017 has unusual leverage
dnf_resid |> filter( abs(.std.resid) >= 2) ## none with usual standardized residual
dnf_resid |> filter(.cooksd > 2) ## 2017 is highly influential
```

## Remove the influential observation from the data set.

```{r}
RR_dnf_data <- RR_dnf_data |> filter(year != 2017)
```

## regraph the scatterplot without the influential observation

```{r}
ggplot(RR_dnf_data, aes(x = flow, y = ratio_DNF)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

## refit the model

```{r}
dnf_mod <- lm(ratio_DNF ~ flow, RR_dnf_data)
summary(dnf_mod)
```
