---
title: "Erging Data Wrangling Answers"
format: html
---
1.  Copy this code to load the libraries needed for this worksheet.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
```

2.  Copy this code to read in the data set.

```{r message=FALSE, warning=FALSE}
erg_df <- read_csv("erg.csv")
erg_df <- erg_df |> mutate(Verified = as.factor(Verified), Gender = as.factor(Gender), Year = as.factor(Year), Distance = as.factor(Distance), Age_Group = as.factor(Age_Group))
```

3.  Make a side-by-side boxplot of the `Speed` for `Verified` and
`Non-Verified` pieces. Give the plot a label.

```{r}
ggplot(data = erg_df, aes(x = Verified, y = Speed)) + geom_boxplot() + theme_minimal() + labs(title = "Side-by-Side Boxplots of Speed for Verified and Non-Verified Pieces") 
```

4.  Describe the distribution of the plot.

**The plots look clustered towards the lower end of the y-axis. This is likely due to the 4 outliers on the non-verified plot which are heavily skewing the data.**



5.  Find the summary statistics for `Speed` and calculate its IQR.

```{r}
summary(erg_df$Speed)
301-256.5
```

**The IQR for `Speed` is 44.5.**

6.  Make a table of the top 10 speeds for rowers. Including only their `Name`, `Speed` and if their piece was `Verified`.

```{r}
erg_df |> arrange(desc(Speed)) |> select(Name, Speed, Verified)
```



7.  Using the table, identify the 4 outliers seen in the boxplot.

```{r}
1.5*44.5
301 + 66.75
```

**Natasshia Lee, Travis Willcox, Lynn Hansford, and Toño Martí Ayala are all outliers as their speeds are greater than `1.5(IQR) + Q3`.**



8.  Perform a t.test to evaluate if there is a significant difference in the
mean `Speed` for `Verified` and `Non-Verified` pieces.

```{r}
t.test(Speed ~ Verified, data = erg_df)
```

**H0: mean speed for verified pieces = mean speed for non-verified pieces** 
**Ha: mean speed for verified pieces != mean speed for non-verified pieces** 
**P-value = 0.877** **0.877 \> 0.05**
**Conclusion: Do not Reject H0** 
**Interpretation: We do not have significant evidence that the mean speed for non-verified pieces is higher than for verified pieces.**

9.  Make a new dataset which does not have the 4 outliers in it.

```{r}
no_outliers <- erg_df |> filter(Name != "Toño Martí ayala" & Name != "Lynn Hansford" & Name != "Travis Willcox" & Name != "Natasshia Lee")
```


10. With the new dataset make another side-by-side boxplot of the `Speed` for `Verified` and
`Non-Verified` pieces and give the plot a label.

```{r}
ggplot(data = no_outliers, aes(x = Verified, y = Speed)) + geom_boxplot() + theme_minimal() + labs(title = "Side-by-Side Boxplots of Speed for Verified and Non-Verified Pieces") + theme(plot.title = element_text(hjust = 0.5))
```

11.  Comment on the distribution of this new boxplot.

**The distribution is much more spread out without the outliers.**

12.  Perform a t.test with the new dataset and re-evalute if there is a
significant difference in the mean `Speed` for `Verified` and
`Non-Verified` pieces.

```{r}
t.test(Speed ~ Verified, data = no_outliers)
```

**H0: mean speed for verified pieces = mean speed for non-verified pieces** 
**Ha: mean speed for verified pieces != mean speed for non-verified pieces** 
**P-value = 3.493e-07** **3.493e-07 \< 0.05**
**Conclusion: Reject H0** 
**Interpretation: We have significant evidence that the mean speed for non-verified pieces is lower than for verified pieces.**

13.  What does the number of outliers for `Non-Verified` pieces say about the reliability of `Non-Verified` pieces in the dataset?

**Since non-verified pieces are manually entered it leaves room for error as people can accidentally enter the wrong values into the logbook. Additionally, given that these values are self reported it is possible for people to lie about their results.**

14.  Is it ethical to remove the outliers from the dataset as done in task 9?

**Given that the outlier entries are likely mistakes, not actual values and that they have such a big impact on the data, it is better to remove them. However, it should be stated before hand that they have been removed from the dataset.**
