---
title: "fencing_question_prep"
format: html
---

# Libraries

```{r}
library(tidyverse)
library(readr)
```

# Data

```{r}
fencing_df <- read_csv("~/Desktop/git/score_new/awsmit22/fencing_module/exploration/fencing.csv")
fencing_df
```

# Question 1

Which of these histograms of `Victories` for each `Weapon` and `Gender` combination looks normally distributed?

```{r}
ggplot(data = fencing_df, aes(x = Victories)) + geom_histogram(bins = 10, colour = "black", fill = "lightblue") + facet_grid( Gender ~ Weapon) + theme_minimal() + theme(axis.title = element_text(size = 12, face = "bold"),strip.text = element_text(size = 12, face = "bold"), plot.title = element_text(size = 20, hjust = 0.5)) + labs(title = "Histograms of Victories for Each Weapon and Both Genders") 
```

*The distribution of `Victories` for womens' foil looks to have a normal distribution, due to its symmetrical appearance.*

# Question 2

The mean number of `Victories` for womens' foil is 11.5 and the standard deviation is 3.97. Can it be concluded that the distribution of `Victories` for womens' foil is normal?

```{r}
wf <- fencing_df |> filter(Gender == "Women" & Weapon == "Foil")
mean(wf$Victories)
sd(wf$Victories)
```

*No, normal distribution has to have a mean of 0 and a standard deviation of 1, just because a variable has a symmetrical distribution does not mean it is normally distributed.*

# Question 3

The 5 number summary of `Victories` for womens' foil is provided below. Calculate the IQR of `Victories` for womens' foil.

```{r}
summary(wf$Victories)
```

```{r}
# Answer
iqr <- 14-9.75
iqr
```

*4.25*

# Question 4

This table shows the 5 fencers with the least `Victories` in women's foil. Identify any outliers using the IQR method. 

```{r}
wf |> arrange(Victories) |> head()
```

```{r}
# Answer
1.5 * 4.25
9.75 - 6.375
```

2 < 3.375

*Alina Lee is an outlier since her number of Victories is less than the lower quantile minus 1.5(IQR).*

# Question 5

Draw a boxplot of `Victories` for womens' foil. 

```{r}
boxplot(wf$Victories)
```

# Question 6

Describe the trend of this scatterplot of `Victories` and `Ind` for the entire dataset.

```{r}
ggplot(data = fencing_df, aes(x=Ind, y = Victories)) + geom_point(alpha = 0.5, colour = "blue") + theme_minimal() + labs(title = "Scatterplot of the Indicator Variable Against Victories") + theme(plot.title = element_text(size = 15, hjust = 0.5))
```

*Victories increase with a higher indicator variable.*

# Question 7

The least squares line for `Victories` predicted by `Ind` is Victories = 11.5 + 0.16Ind. Interpret the slope of the line. 

```{r}
lm(Victories ~ Ind, data = fencing_df)
```

*For every unit increase in the indicator variable, a fencer's victories increase by 0.16.*

# Question 8

A fencer has an indicator score of 16, calculate their predicted `Victories`.

```{r}
victories <- 11.5 + (0.16*16)
victories
```

*A fencer with an indicator score of 16 should have 14.06 or roughly 14 victories.*

# Question 9

The actual value of `Victories` for a fencer with an indicator score of 16 is 15, calculate the residual.

```{r}
fencing_df |> filter(Ind == 16)
```

```{r}
# Answer
resid <- 15-14
resid
```

*1*

# Question 10

The model Victories = 11.5 + 0.16Ind has an r-squared value of 92%. Explain the significance of this value. 

```{r}
mod_9 <- lm(Victories ~ Ind, data = fencing_df)
mod_9 |> summary()
```

*R-squared refers to the proportion of total variability in Y explained by X. In terms of this model, this means that 92% of variablity in victories is explained by the indicator variable. This means that the model does a good job predicting the number of victories.*

