---
title: "Exploration"
author: "Brendan Karadenes"
format: html
---

## Data Exploration

# 213/234 Example Ideas

1. Load in the necessary packages
2. Load in the datasets
3. Reconstruct Plots
4. Clean Data


```{r}
# Load in the necessary packages and data
library(tidyverse)
library(here)
PWHL_Final <- read.csv(here("bkaradenes/PWHL_Module/data/PWHL_Final.csv"))
```

```{r}
ggplot(data = PWHL_Final) + 
  geom_density(aes(G_Per_GP, color = Pos, fill = Pos), alpha = 0.25) +
  theme_minimal()
```

## Question 1

Describe the distributions for each position in the density plot above. Make sure to mention shape and skew.

## Question 2

What would you need to add to the R code below to add a title and change the x-axis label to "Goals Per Games Played". 

```{r}
ggplot(data = PWHL_Graph, aes(x = Age_Group, y = P, color = Pos)) + 
  geom_boxplot() +
  coord_flip() +
  labs(x = "Age Group", y = "Number of Points", title = "Points between different Age Groups and Positions") +
  theme_minimal()
```


## Question 3

First create three different age groups Youngest (22-25), Middle (26-30), and Oldest (31-36). Then, create side-by-side boxplots that display the range of points for each group. Make sure to add a theme, flip the axis, and change the x and y axis labels.

```{r}
PWHL_Graph <- PWHL_Final %>%
  mutate(Age_Group = case_when(
    Age >= 22 & Age <= 26 ~ "youngest",
    Age >= 27 & Age <= 30  ~ "middle",
    Age >= 31 & Age <= 36 ~ "oldest"
  )) %>%
  filter(Pos != "G")
PWHL_Graph <- PWHL_Graph %>%
  mutate(Age_Group = factor(Age_Group, levels = c("youngest", "middle", "oldest")))
```

```{r}
ggplot(data = PWHL_Graph, aes(x = Age_Group, y = P, color = Pos)) + 
  geom_boxplot() +
  coord_flip() +
  labs(x = "Age Group", y = "Number of Points", title = "Points between different Age Groups and Positions") +
  theme_minimal()
```

1. Use the dplyr package to change the variable name "P" to "Points".

```{r}
PWHL_Scoring <-
  PWHL_Scoring %>%
    rename(Points = P)
```

2. Filter for rows with players over the age of 25.

```{r}
PWHL_Scoring %>%
  filter(Age > 25)
```

3. Filter the rows for players that are either Forwards (F) or Defense (D).

```{r}
PWHL_Scoring %>%
  filter(Pos %in% c("F", "D"))
```


